### 2023-08-13 code/230510_Competition.Rmd
first thing

---
title: "220410_Competition"
author: "Justin"
date: "4/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Library 
```{r}
library(tidygeocoder)
library(tidyverse)
library(did)
library(tidycensus)
library(gtsummary)
library(readxl)
library(sandwich)
library(lmtest)
library(ggpubr)
library(kableExtra)
library(tigris)

library(gridExtra)
library(cowplot)
library(Matching)
library(fixest)
library(directlabels)

`%notin%` <- Negate(`%in%`)

select <- dplyr::select

#Sys.setenv('R_MAX_VSIZE'=32000000000)

```

## Define new Tufte data theme
```{r, echo = FALSE, results = "asis", message=FALSE, warning = FALSE, fig.align="center"}

is.extrafont.installed <- function(){
  if(is.element("extrafont", installed.packages()[,1])){
    library(extrafont)
    # probably need something here to run font_import()
    return(T)
  }else{
    warning("Library extrafont installed; using system sans/serif libraries as fallback fonts. 
    To enable full font support, run: 
      install.packages('extrafont') 
      font_import()")
    return(F)
  }
}

# set font
base_font_family_tufte <- function(){
  if(is.extrafont.installed()){
    library(extrafont)
    tuftefont <- choose_font(c("Gill Sans MT", "Gill Sans", "GillSans", "Verdana", "serif"), quiet = FALSE)  
  }else{
    tuftefont <- "serif"
  }
  return(tuftefont)
}

theme_tufte_revised <- function(base_size = 11, base_family = base_font_family_tufte(), ticks = TRUE) {

  ret <- theme_bw(base_family = base_family, base_size = base_size) + 
    theme(
      axis.line = element_line(color = 'black'),
      axis.title.x = element_text(vjust = -0.3), 
      axis.title.y = element_text(vjust = 0.8),
      legend.background = element_blank(), 
      legend.key = element_blank(), 
      legend.title = element_text(face="plain"),
      panel.background = element_blank(), 
      panel.border = element_blank(),
      panel.grid = element_blank(),
      plot.background = element_blank(),
      strip.background = element_blank()
    )

  if (!ticks) {
    ret <- ret + theme(axis.ticks = element_blank())
  }

  ret
}

```

# Data

## Read in UDS

```{r}

## Zips
locs <- paste("UDS/UDS-", c(2014:2021), "-Full-Dataset.xlsx", sep = "")
uds.zip <- read_excel(locs[1], sheet = "HealthCenterZipCodes")

new.zip.names <- c("BHCMIS ID", "Grant Number", "Reporting Year" , "Zip Code", "Zip Code Type",  "None_Uninsured Patients" ,  "Medicaid_CHIP_Other Public Patients", "Medicare Patients", "Private Patients", "Total Number of Patients" ) 
colnames(uds.zip) <-    new.zip.names

for(i in 2:length(locs)){
  temp.zip <- read_excel(locs[i], sheet = "HealthCenterZipCodes")
  colnames(temp.zip) <- new.zip.names
  uds.zip <- rbind(uds.zip, temp.zip %>% mutate(`Zip Code` = str_pad(`Zip Code`, 5, pad = "0")))
}

## Rename
new.zip.names <- c("BHCMIS ID", "Grant Number", "Reporting Year" , "Zip Code", "Zip Code Type",  "None_Uninsured Patients" ,  "Medicaid_CHIP_Other Public Patients", "Medicare Patients", "Private Patients", "Total Number of Patients" )
colnames(uds.zip) <-    new.zip.names

## Clean

uds.zip[uds.zip$`Total Number of Patients`== '-' | uds.zip$`Total Number of Patients`== '--',]$`Total Number of Patients` <- NA
uds.zip$`Total Number of Patients` <- as.numeric(uds.zip$`Total Number of Patients`)
uds.zip <- uds.zip[!is.na(uds.zip$`Total Number of Patients`),] %>%
    mutate(`Reporting Year` = as.numeric(`Reporting Year`))

uds.zip$`None_Uninsured Patients` <- as.numeric(uds.zip$`None_Uninsured Patients`)
uds.zip$`Medicaid_CHIP_Other Public Patients` <- as.numeric(uds.zip$`Medicaid_CHIP_Other Public Patients`)
uds.zip$`Medicare Patients` <- as.numeric(uds.zip$`Medicare Patients`)
uds.zip$`Private Patients` <- as.numeric(uds.zip$`Private Patients`)

uds.zip <- cbind(uds.zip, broken.total = rowSums(uds.zip[,c(6:9)], na.rm = TRUE), missings = rowSums(is.na(uds.zip[,c(6:9)]))) %>%
  mutate(`Total Number of Patients` = as.numeric(`Total Number of Patients`)) %>%
  mutate(`Total Number of Patients` = ifelse(is.na(`Total Number of Patients`), round(missings*16/4), `Total Number of Patients`)) %>%
  mutate(`None_Uninsured Patients` = ifelse(is.na(`None_Uninsured Patients`), (`Total Number of Patients` - broken.total) * 1/missings, `None_Uninsured Patients`),
         `Medicaid_CHIP_Other Public Patients` = ifelse(is.na(`Medicaid_CHIP_Other Public Patients`), (`Total Number of Patients` - broken.total) * 1/missings, `Medicaid_CHIP_Other Public Patients`),
         `Medicare Patients` = ifelse(is.na(`Medicare Patients`), (`Total Number of Patients` - broken.total) * 1/missings, `Medicare Patients`),
         `Private Patients` = ifelse(is.na(`Private Patients`), (`Total Number of Patients` - broken.total) * 1/missings, `Private Patients`))

# ## Old version with missingness still in there  
# uds.zip$`Reporting Year` <- as.numeric(uds.zip$`Reporting Year`)
# uds.zip$`Total Number of Patients` <- as.numeric(uds.zip$`Total Number of Patients`)
# uds.zip$`None_Uninsured Patients` <- as.numeric(uds.zip$`None_Uninsured Patients`)
# uds.zip$`Medicaid_CHIP_Other Public Patients` <- as.numeric(uds.zip$`Medicaid_CHIP_Other Public Patients`)
# uds.zip$`Medicare Patients` <- as.numeric(uds.zip$`Medicare Patients`)
# uds.zip$`Private Patients` <- as.numeric(uds.zip$`Private Patients`)
# 
# 
# uds.zip[is.na(uds.zip$`Total Number of Patients`),]$`Total Number of Patients` <- 0
# uds.zip[grep("-", uds.zip$`None_Uninsured Patients`),]$`None_Uninsured Patients` <- 0
# uds.zip[grep("-", uds.zip$`Medicaid_CHIP_Other Public Patients`),]$`Medicaid_CHIP_Other Public Patients` <- 0
# uds.zip[grep("-", uds.zip$`Medicare Patients`),]$`Medicare Patients` <- 0
# uds.zip[grep("-", uds.zip$`Private Patients`),]$`Private Patients` <- 0
# 
uds.zip.totals <- uds.zip %>%
  select(`Grant Number`, `Reporting Year`, `None_Uninsured Patients`:`Total Number of Patients`) %>%
  group_by(`Grant Number`, `Reporting Year`) %>% summarise(None = sum(`None_Uninsured Patients`, na.rm = T),
                                                           MCR = sum(`Medicare Patients`, na.rm = T),
                                                           COM = sum(`Private Patients`, na.rm = T),
                                                           MCD = sum(`Medicaid_CHIP_Other Public Patients`, na.rm = T),
                                                           Total = sum(`Total Number of Patients`, na.rm = T))

uds.zip <- uds.zip[uds.zip$`Zip Code Type` == "ZipCode",]

## Sites
locs <- paste("UDS/UDS-", c(2014:2021), "-Full-Dataset.xlsx", sep = "")

uds.sites <- read_excel(locs[1], sheet = "HealthCenterSiteInfo") %>% 
  select(GrantNumber, SiteName, SiteType, LocationType, LocationSetting, OperationalSchedule, CalendarSchedule, TotalWeeklyHoursOfOperation, SiteStreetAddress, SiteCity, SiteState, SiteZIPCode) %>% 
  mutate(Year = 2014)

colnames(uds.sites) <- c("Grant Number", "Site Name", "Site Type", "Location Type", "Location Setting", "Operational Schedule", "Calendar Schedule", "Total Weekly Hours Of Operation", "Site Street Address", "Site City", "Site State", "Site ZIP Code", "Year")

for(i in 2:(length(locs)-1)){
  uds.sites <- rbind(uds.sites, read_excel(locs[i], sheet = "HealthCenterSiteInfo") %>% 
                       select("Grant Number", "Site Name", "Site Type", "Location Type", "Location Setting", "Operational Schedule", "Calendar Schedule", "Total Weekly Hours Of Operation", "Site Street Address", "Site City", "Site State", "Site ZIP Code") %>% 
  mutate(Year = 2013 + i))
}

temp.sites <- read_excel(locs[i+1], sheet = "HealthCenterSiteInfo" )%>% 
  select(GrantNumber, SiteName, SiteType, LocationType, LocationSetting, OperationalSchedule, CalendarSchedule, TotalWeeklyHoursOfOperation, SiteStreetAddress, SiteCity, SiteState, SiteZIPCode) %>% 
  mutate(Year = 2021)
colnames(temp.sites) <- c("Grant Number", "Site Name", "Site Type", "Location Type", "Location Setting", "Operational Schedule", "Calendar Schedule", "Total Weekly Hours Of Operation", "Site Street Address", "Site City", "Site State", "Site ZIP Code", "Year")

uds.sites <- rbind(uds.sites, temp.sites)

uds.sites$Zip <- substr(uds.sites$`Site ZIP Code`, 1, 5)

uds.sites <- uds.sites %>% 
  filter(`Site State` %notin% c('AK','HI','GU','VI','PW','MH','AS','FM','PR','MP'))

uds.sites[grep(pattern = "-", uds.sites$Zip),]$Zip <- NA

```

## Clean up some of the missing zip codes

```{r}

tbcoded <- uds.sites[is.na(uds.sites$Zip),]
coded <- geo(street = tbcoded[c(1:1000),]$`Site Street Address`, 
             city = tbcoded[c(1:1000),]$`Site City`, 
             state = tbcoded[c(1:1000),]$`Site State`, 
             method = "census", 
             full_results = TRUE)

for (g in 1:9){
  g
coded <- rbind(coded, 
               geo(street = tbcoded[c((1 + g*1000):((g+1)*1000)),]$`Site Street Address`, 
             city = tbcoded[c((1 + g*1000):((g+1)*1000)),]$`Site City`, 
             state = tbcoded[c((1 + g*1000):((g+1)*1000)),]$`Site State`, 
             method = "census", 
             full_results = TRUE) )
}

coded <- rbind(coded, 
               geo(street = tbcoded[c(10001:10975),]$`Site Street Address`, 
             city = tbcoded[c(10001:10975),]$`Site City`, 
             state = tbcoded[c(10001:10975),]$`Site State`, 
             method = "census", 
             full_results = TRUE) )

coded[grep(pattern = "[[:digit:]]{5}$", coded$matched_address),]

uds.sites[is.na(uds.sites$Zip),]$Zip <- str_extract(coded$matched_address, pattern = "[[:digit:]]{5}$")
uds.sites[is.na(uds.sites$Zip),]

uds.sites.temp <- uds.sites

```

```{r}

uds.sites <- uds.sites.temp

uds.sites.temp %>% filter(`Grant Number` %in% c("H80CS00881") & Year == 2016)

chc <- read.csv("chc.csv")

chc %>% filter(Health.Center.Number %in% c("H80CS00881"))

uds.sites[is.na(uds.sites$Zip),] <- uds.sites[is.na(uds.sites$Zip),] %>%
  left_join(chc, by = c("Grant Number" = "Health.Center.Number", "Site Street Address" = "Site.Address")) %>%
  mutate(Zip = ifelse(is.na(Zip), substr(Health.Center.Organization.ZIP.Code, 1, 5), Zip)) %>%
  select(`Grant Number`:Zip) %>%
  unique()

uds.sites[is.na(uds.sites$Zip),] <- uds.sites[is.na(uds.sites$Zip),] %>%
  left_join(chc, by = c("Grant Number" = "Health.Center.Number", "Site Name" = "Site.Name")) %>%
  mutate(Zip = ifelse(is.na(Zip), substr(Health.Center.Organization.ZIP.Code, 1, 5), Zip)) %>%
  select(`Grant Number`:Zip) %>%
  unique()

uds.sites[is.na(uds.sites$Zip),] 
```

```{r}

table(uds.sites.temp$`Location Setting`)

uds.sites <- uds.sites %>% #filter(`Location Type` == "Permanent" & `Location Setting` %in% c("All Other Clinic Types")) %>% 
  mutate(Access = as.numeric(ifelse(`Total Weekly Hours Of Operation` == "-" & `Operational Schedule` == "Full-Time", 40, 
                  ifelse(`Total Weekly Hours Of Operation` == "-" & `Operational Schedule` == "Part-Time", 10, `Total Weekly Hours Of Operation`))))

uds.sites[is.na(uds.sites$Zip),]

uds.sites %>% 
  arrange(`Total Weekly Hours Of Operation`, `Operational Schedule` )

# z <- geocode(chc.term$SitePhysicalAddress)
# 
# 
# y <- geo(address = chc.term$SitePhysicalAddress, method = "census")
# 
# write.csv(cbind(y[,1], y[,3], y[,2], z), file = "geocodedClosures.csv")

```

## Read in Quality data
```{r}

uds.6b.2012 <- read_excel("UDS/UDS-2014-Full-Dataset.xlsx", sheet = "Table6BClinicalmeasures", skip = 0) 
uds.6b.2013 <- read_excel("UDS/UDS-2014-Full-Dataset.xlsx", sheet = "Table6BClinicalmeasures", skip = 0)

uds.6b.2014 <- read_excel("UDS/UDS-2014-Full-Dataset.xlsx", sheet = "Table6BClinicalmeasures", skip = 0) 
uds.6b.2015 <- read_excel("UDS/UDS-2015-Full-Dataset.xlsx", sheet = "Table6BClinicalmeasures", skip = 1) 
uds.6b.2016 <- read_excel("UDS/UDS-2016-Full-Dataset.xlsx", sheet = "Table6BClinicalmeasures", skip = 0)

uds.6b.2017 <- read_excel("UDS/UDS-2017-Full-Dataset.xlsx", sheet = "Table6BClinicalmeasures", skip = 0) 
uds.6b.2018 <- read_excel("UDS/UDS-2018-Full-Dataset.xlsx", sheet = "Table6BClinicalmeasures", skip = 0) 
uds.6b.2019 <- read_excel("UDS/UDS-2019-Full-Dataset.xlsx", sheet = "Table6BClinicalmeasures", skip = 0)

uds.6b.2020 <- read_excel("UDS/UDS-2020-Full-Dataset.xlsx", sheet = "Table6BClinicalmeasures", skip = 0) 
uds.6b.2021 <- read_excel("UDS/UDS-2021-Full-Dataset.xlsx", sheet = "Table6BClinicalmeasures", skip = 0)

# rbind(
# colnames(uds.6b.2014[,c(2, grep(pattern = "^%", colnames(uds.6b.2014))[-6])]),
# colnames(uds.6b.2015[,c(2, grep(pattern = "^%", colnames(uds.6b.2015))[-6])]),
# colnames(uds.6b.2016[,c(2, grep(pattern = "^%", colnames(uds.6b.2016))[-6])]),
# colnames(uds.6b.2017[,c(2, grep(pattern = "^%", colnames(uds.6b.2017))[-6])]),
# colnames(uds.6b.2018[,c(2, grep(pattern = "^%", colnames(uds.6b.2018))[-6])]),
# colnames(uds.6b.2019[,c(2, grep(pattern = "^%", colnames(uds.6b.2019))[-6])]),
# colnames(uds.6b.2020[,c(2, grep(pattern = "^%", colnames(uds.6b.2020))[-c(3, 11, 13)])]))
# 
# 
# rbind(
# colnames(uds.6b.2014[,c(2, grep(pattern = "^Total", colnames(uds.6b.2014))[-6])]),
# colnames(uds.6b.2015[,c(2, grep(pattern = "^Total", colnames(uds.6b.2015))[-6])]),
# colnames(uds.6b.2016[,c(2, grep(pattern = "^Total", colnames(uds.6b.2016))[-6])]),
# colnames(uds.6b.2017[,c(2, grep(pattern = "^Total", colnames(uds.6b.2017))[-6])]),
# colnames(uds.6b.2018[,c(2, grep(pattern = "^Total", colnames(uds.6b.2018))[-6])]),
# colnames(uds.6b.2019[,c(2, grep(pattern = "^Total", colnames(uds.6b.2019))[-6])]),
# colnames(uds.6b.2020[,c(2, grep(pattern = "^Total", colnames(uds.6b.2020))[-c(3, 11, 13)])]))

namingshit <- c("GrantNumber", "IMM2yo", "Pap", "ChildBMI","AdultBMI","Tobac","CRC","DepScrn")
namingshit.denom <- c("GrantNumber", "IMM2yo.denom", "Pap.denom", "ChildBMI.denom","AdultBMI.denom","Tobac.denom","CRC.denom","DepScrn.denom")

uds.6b.rates <- rbind(
  uds.6b.2014[,c(2, grep(pattern = "^%", colnames(uds.6b.2014))[-6])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit), function(x) namingshit) %>% mutate(Year = 2014), 
  uds.6b.2015[,c(2, grep(pattern = "^%", colnames(uds.6b.2015))[-6])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit), function(x) namingshit) %>% mutate(Year = 2015),
  uds.6b.2016[,c(2, grep(pattern = "^%", colnames(uds.6b.2016))[-6])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit), function(x) namingshit) %>% mutate(Year = 2016),

  uds.6b.2017[,c(2, grep(pattern = "^%", colnames(uds.6b.2017))[-6])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit), function(x) namingshit) %>% mutate(Year = 2017),
  uds.6b.2018[,c(2, grep(pattern = "^%", colnames(uds.6b.2018))[-6])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit), function(x) namingshit) %>% mutate(Year = 2018),

  uds.6b.2019[,c(2, grep(pattern = "^%", colnames(uds.6b.2019))[-6])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit), function(x) namingshit) %>% mutate(Year = 2019),
  uds.6b.2020[,c(2, grep(pattern = "^%", colnames(uds.6b.2020))[-c(3, 11, 13)])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit), function(x) namingshit) %>% mutate(Year = 2020), 
  uds.6b.2021[,c(2, grep(pattern = "^%", colnames(uds.6b.2021))[-c(3, 11, 13)])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit), function(x) namingshit) %>% mutate(Year = 2021))

uds.6b.rates$IMM2yo <- gsub(pattern = "%$", replacement = "", uds.6b.rates$IMM2yo)
uds.6b.rates$Pap <- gsub(pattern = "%$", replacement = "", uds.6b.rates$Pap)
uds.6b.rates$ChildBMI <- gsub(pattern = "%$", replacement = "", uds.6b.rates$ChildBMI)
uds.6b.rates$AdultBMI <- gsub(pattern = "%$", replacement = "", uds.6b.rates$AdultBMI)

uds.6b.rates$Tobac <- gsub(pattern = "%$", replacement = "", uds.6b.rates$Tobac)
uds.6b.rates$CRC <- gsub(pattern = "%$", replacement = "", uds.6b.rates$CRC)
uds.6b.rates$DepScrn <- gsub(pattern = "%$", replacement = "", uds.6b.rates$DepScrn)

uds.6b.denom <- rbind(
  uds.6b.2014[,c(2, grep(pattern = "^Total", colnames(uds.6b.2014))[-6])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit.denom), function(x) namingshit.denom) %>% mutate(Year = 2014), 
  uds.6b.2015[,c(2, grep(pattern = "^Total", colnames(uds.6b.2015))[-6])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit.denom), function(x) namingshit.denom) %>% mutate(Year = 2015),
  uds.6b.2016[,c(2, grep(pattern = "^Total", colnames(uds.6b.2016))[-6])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit.denom), function(x) namingshit.denom) %>% mutate(Year = 2016),

  uds.6b.2017[,c(2, grep(pattern = "^Total", colnames(uds.6b.2017))[-6])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit.denom), function(x) namingshit.denom) %>% mutate(Year = 2017),
  uds.6b.2018[,c(2, grep(pattern = "^Total", colnames(uds.6b.2018))[-6])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit.denom), function(x) namingshit.denom) %>% mutate(Year = 2018),

  uds.6b.2019[,c(2, grep(pattern = "^Total", colnames(uds.6b.2019))[-6])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit.denom), function(x) namingshit.denom) %>% mutate(Year = 2019),
  uds.6b.2020[,c(2, grep(pattern = "^Total", colnames(uds.6b.2020))[-c(3, 11, 13)])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit.denom), function(x) namingshit.denom) %>% mutate(Year = 2020),
  uds.6b.2021[,c(2, grep(pattern = "^Total", colnames(uds.6b.2021))[-c(3, 11, 13)])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit.denom), function(x) namingshit.denom) %>% mutate(Year = 2021))

for (k in 2:8) uds.6b.rates[grep("-", unlist(uds.6b.rates[,k])),k] <- NA
for (k in 2:8) uds.6b.denom[grep("-", unlist(uds.6b.denom[,k])),k] <- NA

uds.6b.rates <-  data.frame(GrantNumber = uds.6b.rates$GrantNumber, Year = uds.6b.rates$Year, lapply(uds.6b.rates[,c(2:8)], as.numeric)) 
uds.6b.denom <-  data.frame(GrantNumber = uds.6b.denom$GrantNumber, Year = uds.6b.denom$Year, lapply(uds.6b.denom[,c(2:8)], as.numeric))

uds.6b.rates[uds.6b.rates$Year == 2014,c(3:9)] <- uds.6b.rates[uds.6b.rates$Year == 2014,c(3:9)] * 100
uds.6b.rates[,c(3:9)] <- uds.6b.rates[,c(3:9)]

## Finding highest-scoring quality FQHC
# uds.6b.rates %>% 
#   mutate(Qcomp = (IMM2yo + Pap + ChildBMI + AdultBMI + Tobac + CRC + DepScrn) / 7) %>% 
#   arrange(desc(Qcomp)) %>% 
#   filter(Year == 2021) %>% 
#   select(GrantNumber, Year, Qcomp) %>% 
#   left_join(chc %>% select(Health.Center.Number, Health.Center.Name, Health.Center.Organization.City, Health.Center.Organization.State) %>% unique(), by = c("GrantNumber" = "Health.Center.Number")) %>% 
#   left_join(uds.6b.denom, by = c("GrantNumber","Year"))

```

## Read in Technology Data
```{r}

uds.HIT.2014 <- read_excel("UDS/UDS-2014-Full-Dataset.xlsx", sheet = "EHRInformation", skip = 0) %>% 
  select(GrantNumber, Tehr_L1b_Ca, Tehr_L2_Ca,	Tehr_L3_Ca, Tehr_L4_Ca, Tehr_L5_Ca, Tehr_L7_Ca)
uds.HIT.2015 <- read_excel("UDS/UDS-2015-Full-Dataset.xlsx", sheet = "EHRInformation", skip = 2) %>% 
  select(GrantNumber, Tehr_L1b_Ca, Tehr_L2_Ca,	Tehr_L3_Ca, Tehr_L4_Ca, Tehr_L5_Ca, Tehr_L7_Ca)
uds.HIT.2016 <- read_excel("UDS/UDS-2016-Full-Dataset.xlsx", sheet = "HITInformation", skip = 1) %>% 
  select(GrantNumber, Tehr_L1b_Ca, Tehr_L2_Ca,	Tehr_L3_Ca, Tehr_L4_Ca, Tehr_L5_Ca, Tehr_L7_Ca)

uds.HIT.2017 <- read_excel("UDS/UDS-2017-Full-Dataset.xlsx", sheet = "HITInformation", skip = 1) %>% 
  select(GrantNumber, Tehr_L1b_Ca, Tehr_L2_Ca,	Tehr_L3_Ca, Tehr_L4_Ca, Tehr_L5_Ca, Tehr_L7_Ca)
uds.HIT.2018 <- read_excel("UDS/UDS-2018-Full-Dataset.xlsx", sheet = "HITInformation", skip = 1) %>% 
  select(GrantNumber, Tehr_L1b_Ca, Tehr_L2_Ca,	Tehr_L3_Ca, Tehr_L4_Ca, Tehr_L5_Ca, Tehr_L7_Ca)
uds.HIT.2019 <- read_excel("UDS/UDS-2019-Full-Dataset.xlsx", sheet = "HITInformation", skip = 1) %>% 
  select(GrantNumber, Tehr_L1b_Ca, Tehr_L2_Ca,	Tehr_L3_Ca, Tehr_L4_Ca, Tehr_L5_Ca, Tehr_L7_Ca)

uds.HIT.2020 <- read_excel("UDS/UDS-2020-Full-Dataset.xlsx", sheet = "HITInformation", skip = 1) %>%
  select(GrantNumber, Tehr_L1b_Ca, Tehr_L4_Ca, Tehr_L7_Ca)
uds.HIT.2021 <- read_excel("UDS/UDS-2021-Full-Dataset.xlsx", sheet = "HITInformation", skip = 0) %>%
  select(GrantNumber, Tehr_L1b_Ca, Tehr_L4_Ca, Tehr_L7_Ca)

uds.HIT.2020 <- data.frame(GrantNumber = uds.HIT.2020[,1], Tehr_L1b_Ca = uds.HIT.2020[,2], Tehr_L2_Ca = NA,	Tehr_L3_Ca = NA, Tehr_L4_Ca = uds.HIT.2020[,3], Tehr_L5_Ca = NA, Tehr_L7_Ca = uds.HIT.2020[,4])
uds.HIT.2021 <- data.frame(GrantNumber = uds.HIT.2021[,1], Tehr_L1b_Ca = uds.HIT.2021[,2], Tehr_L2_Ca = NA,	Tehr_L3_Ca = NA, Tehr_L4_Ca = uds.HIT.2021[,3], Tehr_L5_Ca = NA, Tehr_L7_Ca = uds.HIT.2021[,4])

namevec <- c("Year", "Grant Number", "SwitchEMR","eScripts","eDecisionSupport","InterOrgHITexchange","PatientPortals","UDSQualRep")
uds.HIT <- rbind(cbind(Year = 2014, uds.HIT.2014),
                  cbind(Year = 2015, uds.HIT.2015),
                  cbind(Year = 2016, uds.HIT.2016),
                  cbind(Year = 2017, uds.HIT.2017),

                  cbind(Year = 2018, uds.HIT.2018),
                  cbind(Year = 2019, uds.HIT.2019), 
                  cbind(Year = 2020, uds.HIT.2020),
                  cbind(Year = 2021, uds.HIT.2021)) %>%

  rename_all(vars(namevec), function(x) namevec) %>% 
  mutate(UDSQualRep.levels = ifelse(UDSQualRep == "-", NA, 
                             ifelse(UDSQualRep == "We do not use the EHR", 0, 
                             ifelse(UDSQualRep == "We use the EHR but only to access individual patient charts", 1, 
                             ifelse(UDSQualRep == "We use the EHR in combination with another data analytic system", 3, 2))))) %>% 
  mutate(UDSQualRep.highlowuse = ifelse(UDSQualRep == "-", NA, 
                             ifelse(UDSQualRep == "We do not use the EHR" | UDSQualRep == "We use the EHR but only to access individual patient charts", 0, 1))) %>%  
  mutate(SwitchEMR = ifelse(SwitchEMR == "-", NA, ifelse(SwitchEMR == "Yes", 1, 0)),
         eScripts = ifelse(eScripts == "-" | eScripts == "Not Sure", NA, ifelse(eScripts == "Yes", 1, 0)),
         eDecisionSupport = ifelse(eDecisionSupport == "-" | eDecisionSupport == "Not Sure", NA, ifelse(eDecisionSupport == "Yes", 1, 0)),
         InterOrgHITexchange = ifelse(InterOrgHITexchange == "-" | InterOrgHITexchange == "Not Sure", NA, ifelse(InterOrgHITexchange %in% c("No","None of the above"), 0, 1)),
         PatientPortals = ifelse(PatientPortals == "-" | PatientPortals == "Not Sure", NA, ifelse(PatientPortals %in% c("No","No we do not engage patients using HIT"), 0, 1)))

uds.HIT[uds.HIT$UDSQualRep.levels == 0 & !is.na(uds.HIT$UDSQualRep.levels) & is.na(uds.HIT$SwitchEMR),]$SwitchEMR <- 0
uds.HIT[uds.HIT$UDSQualRep.levels == 0 & !is.na(uds.HIT$UDSQualRep.levels) & is.na(uds.HIT$eScripts),]$eScripts <- 0
uds.HIT[uds.HIT$UDSQualRep.levels == 0 & !is.na(uds.HIT$UDSQualRep.levels) & is.na(uds.HIT$eDecisionSupport),]$eDecisionSupport <- 0
uds.HIT[uds.HIT$UDSQualRep.levels == 0 & !is.na(uds.HIT$UDSQualRep.levels) & is.na(uds.HIT$InterOrgHITexchange),]$InterOrgHITexchange <- 0
uds.HIT[uds.HIT$UDSQualRep.levels == 0 & !is.na(uds.HIT$UDSQualRep.levels) & is.na(uds.HIT$PatientPortals),]$PatientPortals <- 0

# namevec <- c("Year", "Grant Number", "SwitchEMR","InterOrgHITexchange","UDSQualRep")
# uds.HIT <- rbind(cbind(Year = 2014, uds.HIT.2014 %>% select(GrantNumber, Tehr_L1b_Ca, Tehr_L4_Ca, Tehr_L7_Ca)),
#                   cbind(Year = 2015, uds.HIT.2015 %>% select(GrantNumber, Tehr_L1b_Ca, Tehr_L4_Ca, Tehr_L7_Ca)),
#                   cbind(Year = 2016, uds.HIT.2016 %>% select(GrantNumber, Tehr_L1b_Ca, Tehr_L4_Ca, Tehr_L7_Ca)),
#                   cbind(Year = 2017, uds.HIT.2017 %>% select(GrantNumber, Tehr_L1b_Ca, Tehr_L4_Ca, Tehr_L7_Ca)),
#                   
#                   cbind(Year = 2018, uds.HIT.2018 %>% select(GrantNumber, Tehr_L1b_Ca, Tehr_L4_Ca, Tehr_L7_Ca)),
#                   cbind(Year = 2019, uds.HIT.2019 %>% select(GrantNumber, Tehr_L1b_Ca, Tehr_L4_Ca, Tehr_L7_Ca)),          
#                   cbind(Year = 2020, uds.HIT.2020)) %>% 
#   
#   rename_all(vars(namevec), function(x) namevec) %>% 
#   mutate(UDSQualRep.levels = ifelse(UDSQualRep == "-", NA, 
#                              ifelse(UDSQualRep == "We do not use the EHR", 0, 
#                              ifelse(UDSQualRep == "We use the EHR but only to access individual patient charts", 1, 
#                              ifelse(UDSQualRep == "We use the EHR in combination with another data analytic system", 3, 2))))) %>% 
#   mutate(UDSQualRep.highlowuse = ifelse(UDSQualRep == "-", NA, 
#                              ifelse(UDSQualRep == "We do not use the EHR" | UDSQualRep == "We use the EHR but only to access individual patient charts", 0, 1))) %>%  
#   mutate(SwitchEMR = ifelse(SwitchEMR == "-", NA, ifelse(SwitchEMR == "Yes", 1, 0)),
#          InterOrgHITexchange = ifelse(InterOrgHITexchange == "-" | InterOrgHITexchange == "Not Sure", NA, ifelse(InterOrgHITexchange %in% c("No","None of the above"), 0, 1)))

```

## Read in Covariates
```{r}
# Table 3B
uds.3B.2014 <- read_excel("UDS/UDS-2014-Full-Dataset.xlsx", sheet = "Table3B", skip = 0) %>% select(GrantNumber, T3b_L1_Cb, T3b_L3_Cb, T3b_L8_Ca, T3b_L5_Cb, T3b_L8_Cd, T3b_L12_Ca) %>% rename(T3b_L26_Ca = T3b_L8_Cd)
uds.3B.2015 <- read_excel("UDS/UDS-2015-Full-Dataset.xlsx", sheet = "Table3B", skip = 2) %>% select(GrantNumber, T3b_L1_Cb, T3b_L3_Cb, T3b_L8_Ca, T3b_L5_Cb, T3b_L8_Cd, T3b_L12_Ca) %>% rename(T3b_L26_Ca = T3b_L8_Cd)
uds.3B.2016 <- read_excel("UDS/UDS-2016-Full-Dataset.xlsx", sheet = "Table3B", skip = 1) %>% select(GrantNumber, T3b_L1_Cb, T3b_L3_Cb, T3b_L8_Ca, T3b_L5_Cb, T3b_L26_Ca, T3b_L12_Ca)

uds.3B.2017 <- read_excel("UDS/UDS-2017-Full-Dataset.xlsx", sheet = "Table3B", skip = 1) %>% select(GrantNumber, T3b_L1_Cb, T3b_L3_Cb, T3b_L8_Ca, T3b_L5_Cb, T3b_L26_Ca, T3b_L12_Ca)
uds.3B.2018 <- read_excel("UDS/UDS-2018-Full-Dataset.xlsx", sheet = "Table3B", skip = 1) %>% select(GrantNumber, T3b_L1_Cb, T3b_L3_Cb, T3b_L8_Ca, T3b_L5_Cb, T3b_L26_Ca, T3b_L12_Ca)
uds.3B.2019 <- read_excel("UDS/UDS-2019-Full-Dataset.xlsx", sheet = "Table3B", skip = 1) %>% select(GrantNumber, T3b_L1_Cb, T3b_L3_Cb, T3b_L8_Ca, T3b_L5_Cb, T3b_L26_Ca, T3b_L12_Ca)

uds.3B.2020 <- read_excel("UDS/UDS-2020-Full-Dataset.xlsx", sheet = "Table3B", skip = 1) %>% select(GrantNumber, T3b_L1_Cb, T3b_L3_Cb, T3b_L8_Ca, T3b_L5_Cb, T3b_L26_Ca, T3b_L12_Ca)
uds.3B.2021 <- read_excel("UDS/UDS-2021-Full-Dataset.xlsx", sheet = "Table3B", skip = 0) %>% select(GrantNumber, T3b_L1_Cb, T3b_L3_Cb, T3b_L8_Ca, T3b_L5_Cb, T3b_L26_Ca, T3b_L12_Ca)

namevec <- c("Year", "Grant Number", "NHA", "NHB", "HL", "NHW", "Total", "Eng2ndLang")
uds.3B <- rbind(cbind(Year = 2014, uds.3B.2014),
                  cbind(Year = 2015, uds.3B.2015),
                  cbind(Year = 2016, uds.3B.2016),
                  cbind(Year = 2017, uds.3B.2017),

                  cbind(Year = 2018, uds.3B.2018),
                  cbind(Year = 2019, uds.3B.2019), 
                  cbind(Year = 2020, uds.3B.2020),
                  cbind(Year = 2021, uds.3B.2021)) %>%

  rename_all(vars(namevec), function(x) namevec)

# Table 4
uds.4.2014 <- read_excel("UDS/UDS-2014-Full-Dataset.xlsx", sheet = "Table4", skip = 0) %>% select(GrantNumber, T4_L1_Ca, T4_L2_Ca, T4_L3_Ca, T4_L4_Ca, T4_L5_Ca, T4_L16_Ca, T4_L23_Ca, T4_L12_Ca, T4_L12_Cb) 
uds.4.2015 <- read_excel("UDS/UDS-2015-Full-Dataset.xlsx", sheet = "Table4", skip = 2) %>% select(GrantNumber, T4_L1_Ca, T4_L2_Ca, T4_L3_Ca, T4_L4_Ca, T4_L5_Ca, T4_L16_Ca, T4_L23_Ca, T4_L12_Ca, T4_L12_Cb) 
uds.4.2016 <- read_excel("UDS/UDS-2016-Full-Dataset.xlsx", sheet = "Table4", skip = 1) %>% select(GrantNumber, T4_L1_Ca, T4_L2_Ca, T4_L3_Ca, T4_L4_Ca, T4_L5_Ca, T4_L16_Ca, T4_L23_Ca, T4_L12_Ca, T4_L12_Cb)

uds.4.2017 <- read_excel("UDS/UDS-2017-Full-Dataset.xlsx", sheet = "Table4", skip = 1) %>% select(GrantNumber, T4_L1_Ca, T4_L2_Ca, T4_L3_Ca, T4_L4_Ca, T4_L5_Ca, T4_L16_Ca, T4_L23_Ca, T4_L12_Ca, T4_L12_Cb) 
uds.4.2018 <- read_excel("UDS/UDS-2018-Full-Dataset.xlsx", sheet = "Table4", skip = 1) %>% select(GrantNumber, T4_L1_Ca, T4_L2_Ca, T4_L3_Ca, T4_L4_Ca, T4_L5_Ca, T4_L16_Ca, T4_L23_Ca, T4_L12_Ca, T4_L12_Cb) 
uds.4.2019 <- read_excel("UDS/UDS-2019-Full-Dataset.xlsx", sheet = "Table4", skip = 1) %>% select(GrantNumber, T4_L1_Ca, T4_L2_Ca, T4_L3_Ca, T4_L4_Ca, T4_L5_Ca, T4_L16_Ca, T4_L23_Ca, T4_L12_Ca, T4_L12_Cb)

uds.4.2020 <- read_excel("UDS/UDS-2020-Full-Dataset.xlsx", sheet = "Table4", skip = 1) %>% select(GrantNumber, T4_L1_Ca, T4_L2_Ca, T4_L3_Ca, T4_L4_Ca, T4_L5_Ca, T4_L16_Ca, T4_L23_Ca, T4_L12_Ca, T4_L12_Cb) 
uds.4.2021 <- read_excel("UDS/UDS-2021-Full-Dataset.xlsx", sheet = "Table4", skip = 0) %>% select(GrantNumber, T4_L1_Ca, T4_L2_Ca, T4_L3_Ca, T4_L4_Ca, T4_L5_Ca, T4_L16_Ca, T4_L23_Ca, T4_L12_Ca, T4_L12_Cb)

namevec <- c("Year", "Grant Number", "<100%", "101-150%", "151-200%", ">200%", "Unknown", "FarmWorker", "Homeless", "0-17yo", "18+yo")
uds.4 <- rbind(cbind(Year = 2014, uds.4.2014),
                  cbind(Year = 2015, uds.4.2015),
                  cbind(Year = 2016, uds.4.2016),
                  cbind(Year = 2017, uds.4.2017),

                  cbind(Year = 2018, uds.4.2018),
                  cbind(Year = 2019, uds.4.2019), 
                  cbind(Year = 2020, uds.4.2020),
                  cbind(Year = 2021, uds.4.2021)) %>%

  rename_all(vars(namevec), function(x) namevec)

# Put covariate tables together

uds.covariates <- inner_join(uds.3B, uds.4, by = c("Year","Grant Number"))

for (i in 3:ncol(uds.covariates)){
  uds.covariates[,i] <- as.numeric(gsub("-", "", uds.covariates[,i]))
}

#uds.covariates %>% filter(`Grant Number` == "H80CS24200")

uds.covariates <-  uds.covariates %>% 
    mutate(Pct.NHA = NHA / Total * 100,
         Pct.NHB = NHB / Total * 100,
         Pct.HL = HL / Total * 100,
         Pct.NHW = NHW / Total * 100,
         Pct.ESL = Eng2ndLang / Total * 100,
         Pct.LessThan100pctFPL = ifelse(Total == Unknown, 100, `<100%` / (Total- Unknown) * 100),
         Pct.LessThan200pctFPL = ifelse(Total == Unknown, 100, (`<100%` + `101-150%` + `151-200%`) / (Total - Unknown) * 100),
         Pct.FarmWorker = FarmWorker / Total * 100,
         Pct.Homeless = Homeless / Total * 100,
         Pct.Children = `0-17yo` / Total * 100) %>% 
    select(Year, `Grant Number`, Pct.NHA:Pct.Children)

```

## Read in Clinical Outcomes
```{r}
uds.7.2014 <- read_excel("UDS/UDS-2014-Full-Dataset.xlsx", sheet = "Table7_1", skip = 0) %>% select(GrantNumber, T7_Li_C1a, T7_Li_C1b,	T7_Li_C1c,	T7_Li_C1d, T7_Li_C2a,	T7_Li_C2b,	T7_Li_C2c) 
uds.7.2015 <- read_excel("UDS/UDS-2015-Full-Dataset.xlsx", sheet = "Table7_1", skip = 2) %>% select(GrantNumber, T7_Li_C1a, T7_Li_C1b,	T7_Li_C1c,	T7_Li_C1d, T7_Li_C2a,	T7_Li_C2b,	T7_Li_C2c) 
uds.7.2016 <- read_excel("UDS/UDS-2016-Full-Dataset.xlsx", sheet = "Table7_1", skip = 1) %>% select(GrantNumber, T7_Li_C1a, T7_Li_C1b,	T7_Li_C1c,	T7_Li_C1d, T7_Li_C2a,	T7_Li_C2b,	T7_Li_C2c)

uds.7.2017 <- read_excel("UDS/UDS-2017-Full-Dataset.xlsx", sheet = "Table7_1", skip = 1) %>% select(GrantNumber, T7_Li_C1a, T7_Li_C1b,	T7_Li_C1c,	T7_Li_C1d, T7_Li_C2a,	T7_Li_C2b,	T7_Li_C2c) 
uds.7.2018 <- read_excel("UDS/UDS-2018-Full-Dataset.xlsx", sheet = "Table7_1", skip = 1) %>% select(GrantNumber, T7_Li_C1a, T7_Li_C1b,	T7_Li_C1c,	T7_Li_C1d, T7_Li_C2a,	T7_Li_C2b,	T7_Li_C2c) 
uds.7.2019 <- read_excel("UDS/UDS-2019-Full-Dataset.xlsx", sheet = "Table7_1", skip = 1) %>% select(GrantNumber, T7_Li_C1a, T7_Li_C1b,	T7_Li_C1c,	T7_Li_C1d, T7_Li_C2a,	T7_Li_C2b,	T7_Li_C2c)

uds.7.2020 <- read_excel("UDS/UDS-2020-Full-Dataset.xlsx", sheet = "Table7_1", skip = 1) %>% select(GrantNumber, T7_Li_C1a, T7_Li_C1b,	T7_Li_C1c,	T7_Li_C1d, T7_Li_C2a,	T7_Li_C2b,	T7_Li_C2c) 
uds.7.2021 <- read_excel("UDS/UDS-2021-Full-Dataset.xlsx", sheet = "Table7_1", skip = 0) %>% select(GrantNumber, T7_Li_C1a, T7_Li_C1b,	T7_Li_C1c,	T7_Li_C1d, T7_Li_C2a,	T7_Li_C2b,	T7_Li_C2c)

namevec <- c("Year", "Grant Number", "TotalDeliveries", "BirthsLT1500","Births1500to2499","BirthsGTE2500","TotalHTN","Sample","Controlcount")
uds.7 <- rbind(cbind(Year = 2014, uds.7.2014),
                  cbind(Year = 2015, uds.7.2015),
                  cbind(Year = 2016, uds.7.2016),
                  cbind(Year = 2017, uds.7.2017),

                  cbind(Year = 2018, uds.7.2018),
                  cbind(Year = 2019, uds.7.2019), 
                  cbind(Year = 2020, uds.7.2020),
                  cbind(Year = 2021, uds.7.2021)) %>%

  rename_all(vars(namevec), function(x) namevec)

for (i in 3:ncol(uds.7)){
  uds.7[,i] <- as.numeric(gsub("-", "", uds.7[,i]))
}

```

## Read in my financial data
## Prep download
```{r}

cw <- read_excel("/Users/jhm65/Research/CHC Costs/UDS-990 Crosswalk.xlsx")

## Wave 1

EINs <- cw[!is.na(cw$EIN),]$EIN

urls <- paste("https://www.guidestar.org/profile/GetFTASheet?ein=", cw[!is.na(cw$EIN),]$EIN, "&guid=", sep = "")
destfiles <- paste("/Users/jhm65/Research/CHC Costs/990s/", cw[!is.na(cw$EIN),]$EIN, ".xlsx", sep = "") #990s_new or 990s

## Set Binary filter for data to include
z <- rep(NA, length(urls))
for (k in 1:length(EINs)){  #1174

 z[k] <- ifelse(is.data.frame(try(read_excel(destfiles[k], skip = 5) )) == TRUE, 1, 0)

}

## Set first rows in data
chc.990 <- read_excel(destfiles[1], skip = 5)
years <- as.numeric(colnames(chc.990)[-1])
chc.990 <- chc.990[-c(5, 6, 15, 16, 25, 26, 32, 33, 34, 35, 39, 40, 52:58),]

chc.990[,2] <- as.numeric(gsub(unlist(chc.990[,2]), pattern = "N/A", replacement = NA))
chc.990[,3] <- as.numeric(gsub(unlist(chc.990[,3]), pattern = "N/A", replacement = NA))
chc.990[,4] <- as.numeric(gsub(unlist(chc.990[,4]), pattern = "N/A", replacement = NA))
chc.990[,5] <- as.numeric(gsub(unlist(chc.990[,5]), pattern = "N/A", replacement = NA))
chc.990[,6] <- as.numeric(gsub(unlist(chc.990[,6]), pattern = "N/A", replacement = NA))

headers <- unlist(chc.990[,1])
chc.990 <- data.frame(t(chc.990[,-1]))

colnames(chc.990) <- headers
rownames(chc.990) <- NULL

chc.finances <- cbind(ID = EINs[1], year = years, chc.990)

## Repeat for the whole set
for (k in 2:1174){
  if(z[k] == 1 & k %notin% c(388, 511, 670)){ ##remove entries for now, check later if this clinic gets data
chc.990 <- read_excel(destfiles[k], skip = 5)
years <- as.numeric(colnames(chc.990)[-1])
chc.990 <- chc.990[-c(5, 6, 15, 16, 25, 26, 32, 33, 34, 35, 39, 40, 52:58),]

  for (j in 2:ncol(chc.990)){
    chc.990[,j] <- as.numeric(gsub(unlist(chc.990[,j]), pattern = "N/A", replacement = NA))
  }

headers <- unlist(chc.990[,1])
chc.990 <- data.frame(t(chc.990[,-1]))

colnames(chc.990) <- headers
rownames(chc.990) <- NULL

chc.finances <- rbind(chc.finances, cbind(ID = EINs[k], year = years, chc.990) )} else {print(k)}
}

table(chc.finances$year)

quantile(chc.finances$`Unrestricted surplus (deficit) after depreciation`, probs = c(0, 0.025, 0.1, 0.25, 0.5, 0.75, 0.9, 0.975, 1))

colnames(chc.finances) <- gsub("\\%", "pct", colnames(chc.finances))
colnames(chc.finances)[c(4, 6)] <- c("Surplus as a pct of before dep expenses","Surplus as a pct of after dep expenses")

## Wave 2

EINs <- cw[!is.na(cw$EIN),]$EIN

urls <- paste("https://www.guidestar.org/profile/GetFTASheet?ein=", cw[!is.na(cw$EIN),]$EIN, "&guid=", sep = "")
destfiles <- paste("/Users/jhm65/Research/CHC Costs/990s_new/", cw[!is.na(cw$EIN),]$EIN, ".xlsx", sep = "") #990s_new or 990s

## Set Binary filter for data to include
z <- rep(NA, length(urls))
for (k in 1:length(EINs)){  #1174

 z[k] <- ifelse(is.data.frame(try(read_excel(destfiles[k], skip = 5) )) == TRUE, 1, 0)

}

## Set first rows in data
chc.990 <- read_excel(destfiles[1], skip = 5)
years <- as.numeric(colnames(chc.990)[-1])
chc.990 <- chc.990[-c(5, 6, 15, 16, 25, 26, 32, 33, 34, 35, 39, 40, 52:58),]

chc.990[,2] <- as.numeric(gsub(unlist(chc.990[,2]), pattern = "N/A", replacement = NA))
chc.990[,3] <- as.numeric(gsub(unlist(chc.990[,3]), pattern = "N/A", replacement = NA))
chc.990[,4] <- as.numeric(gsub(unlist(chc.990[,4]), pattern = "N/A", replacement = NA))
chc.990[,5] <- as.numeric(gsub(unlist(chc.990[,5]), pattern = "N/A", replacement = NA))
chc.990[,6] <- as.numeric(gsub(unlist(chc.990[,6]), pattern = "N/A", replacement = NA))

headers <- unlist(chc.990[,1])
chc.990 <- data.frame(t(chc.990[,-1]))

colnames(chc.990) <- headers
rownames(chc.990) <- NULL

chc.fin.2 <- cbind(ID = EINs[1], year = years, chc.990)

## Repeat for the whole set
for (k in 2:1174){
  if(z[k] == 1 & k %notin% c(388, 511, 670)){ ##remove entries for now, check later if this clinic gets data
chc.990 <- read_excel(destfiles[k], skip = 5)
years <- as.numeric(colnames(chc.990)[-1])
chc.990 <- chc.990[-c(5, 6, 15, 16, 25, 26, 32, 33, 34, 35, 39, 40, 52:58),]

  for (j in 2:ncol(chc.990)){
    chc.990[,j] <- as.numeric(gsub(unlist(chc.990[,j]), pattern = "N/A", replacement = NA))
  }

headers <- unlist(chc.990[,1])
chc.990 <- data.frame(t(chc.990[,-1]))

colnames(chc.990) <- headers
rownames(chc.990) <- NULL

chc.fin.2 <- rbind(chc.fin.2, cbind(ID = EINs[k], year = years, chc.990) )} else {print(k)}
}

table(chc.fin.2$year)

quantile(chc.fin.2$`Unrestricted surplus (deficit) after depreciation`, probs = c(0, 0.025, 0.1, 0.25, 0.5, 0.75, 0.9, 0.975, 1))

colnames(chc.fin.2) <- gsub("\\%", "pct", colnames(chc.fin.2))
colnames(chc.fin.2)[c(4, 6)] <- c("Surplus as a pct of before dep expenses","Surplus as a pct of after dep expenses")

nrow(chc.finances[chc.finances$`Unrestricted surplus (deficit) after depreciation` <= 250000 & chc.finances$year == 2017,]) / nrow(chc.finances[chc.finances$year == 2017,])

```

## Read in my termination crosswalk to filter out closures
```{r}
chc.term <- read.csv("CHC_termination_crosswalk.csv")
staffing <- read.csv("/Users/jhm65/Research/CHC Costs/FQHC_Staffing.csv")
```

## Read in census data for all zip codes
```{r}

census <- data.frame(ZipCode = 0, NAME = 0, YEAR = 0, 
                     TotalPop = 0,
                     Male = 0, Female = 0, 
                     AgeUnder18 = 0, Age18to64 = 0, Age65andOver = 0, 
                     White = 0, Black = 0, AIAN = 0, Asian = 0, MR = 0, Hispanic = 0, EducLow = 0, EducMid = 0, EducHigh = 0, 
                     PovertyStatusDet = 0, IncomeBelowPoverty = 0, Unemployed = 0, Unemp_denom = 0, 
                     MCD.denom = 0, Medicaid = 0, Unins.denom = 0, Uninsured = 0)

for (i in 1:7){
  census <- rbind(census, get_acs(geography = "zcta", survey = "acs5",
                       variables = c("B01003_001","B17001_001","B17001_002","B01001_002","B01001_026",
                                     "B02001_002","B02001_003","B02001_004","B02001_005","B02001_006","B02001_008",
                                     "B03001_003", 
                                     "B01001_003","B01001_004","B01001_005","B01001_006","B01001_007","B01001_008","B01001_009",
                                     "B01001_010","B01001_011","B01001_013","B01001_014","B01001_015","B01001_016","B01001_017",
                                     "B01001_018","B01001_019","B01001_020","B01001_021","B01001_022","B01001_023","B01001_024","B01001_025",
                                     "B01001_027","B01001_028","B01001_029","B01001_030","B01001_031","B01001_032","B01001_033","B01001_034",
                                     "B01001_035","B01001_036","B01001_037","B01001_038","B01001_039","B01001_040","B01001_041","B01001_042","B01001_043",
                                     "B01001_044","B01001_045","B01001_046","B01001_047","B01001_048","B01001_049",
                                     "B06009_002","B06009_003", "B06009_004","B06009_005","B06009_006",
                                     "B23025_002","B23025_005",
                                     "C27007_001", "C27007_004", "C27007_007", "C27007_010", "C27007_014", "C27007_017", "C27007_020",
                                     "B18135_001","B18135_007","B18135_012","B18135_018","B18135_023","B18135_029","B18135_034"), 
                       year = 2013 + i) %>% 
  pivot_wider(id_cols = c(GEOID, NAME), names_from = variable, values_from = estimate) %>% 
    mutate(YEAR = 2013 + i, TotalPop = B01003_001, PovertyStatusDet = B17001_001, IncomeBelowPoverty = B17001_002,
           AgeUnder18 = B01001_003 + B01001_004 + B01001_005 + B01001_006 + B01001_027 + B01001_028 + B01001_029 + B01001_030,
           Age18to64 = B01001_007 + B01001_008 + B01001_009 + B01001_010 + B01001_011 + B01001_013 + B01001_014 + B01001_015 + 
                       B01001_016 + B01001_017 + B01001_018 + B01001_019 + B01001_031 + B01001_032 + B01001_033 + B01001_034 +
                       B01001_035 + B01001_036 + B01001_037 + B01001_038 + B01001_039 + B01001_040 + B01001_041 + B01001_042 + B01001_043, 
           Age65andOver = B01001_020 + B01001_021 + B01001_022 + B01001_023 + B01001_024 + B01001_025 + 
                          B01001_044 + B01001_045 + B01001_046 + B01001_047 + B01001_048 + B01001_049, 
           Male = B01001_002, Female = B01001_026, 
           White = B02001_002, Black = B02001_003, AIAN = B02001_004, Asian = B02001_005 + B02001_006, MR = B02001_008, 
           Hispanic = B03001_003, 
           EducLow = B06009_002, EducMid = B06009_003 + B06009_004, EducHigh = B06009_005 + B06009_006,
           Unemployed = B23025_005, Unemp_denom = B23025_002,
           MCD.denom = C27007_001	, Medicaid = C27007_004	 + C27007_007 + C27007_010	+ C27007_014 + C27007_017 + C27007_020,
           Unins.denom = B18135_001, Uninsured = B18135_007 + B18135_012 + B18135_018 + B18135_023 + B18135_029 + B18135_034) %>% 
    select(ZipCode = GEOID, NAME, YEAR, TotalPop:Uninsured))

  print(i)
}
census <- census[-1,]

##########
census$ZipCode <- substr(census$NAME, 7, 11)
#census$ZipCode <- str_pad(census$ZipCode, 5, pad = "0")

census %>% 
  filter(YEAR == 2020) %>% 
  select(ZipCode) %>% 
  #unique() %>% 
  table() %>% 
  table()
```

# Analysis
## SEt data again

```{r}

uds.sites[!is.na(uds.sites$Zip),] %>% 
  group_by(`Grant Number`, Zip, Year) %>% 
  summarise(Counts = n()) %>% 
  select(`Grant Number`, Zip, Year, Counts) %>% 
  unique() %>% 
  arrange(Year) %>% 
  pivot_wider(id_cols = c(`Grant Number`, Zip), names_from = Year, values_from = Counts, values_fill = 0) %>% 
  arrange(Zip)

```

## Set Data
```{r}
# Identify number of clinics in each zip code-year
clinics.in.zips <- uds.sites[!is.na(uds.sites$Zip),] %>%
  group_by(Zip, Year) %>%
  summarise(Clinics = n_distinct(`Grant Number`), Sites = n()) %>%
  arrange(desc(Clinics))

# Identify clinics in competitive zip codes
clinics.comp <- uds.sites[!is.na(uds.sites$Zip),] %>% 
  select(`Grant Number`, Zip, Year) %>% 
  unique() %>% 
  left_join(clinics.in.zips, by = c("Zip", "Year")) %>% 
  group_by(`Grant Number`, Year) %>% 
  summarise(Clinics = max(Clinics)) %>% 
  pivot_wider(id_cols = `Grant Number`, names_from = Year, values_from = Clinics) %>% 
  filter(`2014` == 1 & `2015` == 1 & `2016` == 1 ) %>%   
  #filter(`2014` == 1 & `2015` == 1) %>%  
  left_join(chc.term, by = c("Grant Number" = "ClinicID")) %>%
  filter(is.na(NewClinicID)) %>%
  select(`Grant Number`:`2021`) %>% 
  pivot_longer(cols = c(`2014`:`2021`)) %>% 
  group_by(`Grant Number`) %>% 
  filter(min(value) == 1 ) %>% 
  #filter(max(value) == last(value)) %>% 
  mutate(Tx = ifelse(value > 1, 1, 0))

# Classify clinics by introduction of competition
df.comp <- left_join(clinics.comp %>% mutate(Year = as.numeric(name)), uds.zip.totals, by = c("Grant Number", "Year" = "Reporting Year")) %>% 
  left_join(uds.sites[!is.na(uds.sites$Zip),] %>% 
  group_by(`Grant Number`, Year) %>% 
  summarise(Sites = n()), by = c("Grant Number", "Year")) %>% 
  ungroup() %>% 
  group_by(`Grant Number`) %>% 
  mutate(Att = min(ifelse(Tx == 0, 2, 1)*Year, na.rm = TRUE) ) %>% 
  group_by(`Grant Number`) %>% 
  mutate(Att = ifelse(Att < 2022, Att, 0)) %>% 
  ungroup() %>% 
  mutate(ID = as.numeric(factor(`Grant Number`)))

# Identify expandors vs expandees
df.expand <- df.comp %>% 
  left_join(
      left_join(df.comp %>% select(`Grant Number`, Att), uds.sites[!is.na(uds.sites$Zip),] %>% 
        select(`Grant Number`, Zip, Year) %>% 
        group_by(`Grant Number`, Zip) %>% 
        summarise(YearFirstInZip = min(Year)) %>% 
        filter(YearFirstInZip %notin% c(2014)) %>% 
        select(`Grant Number`, YearFirstInZip) %>% 
        unique(), by = c("Grant Number")) %>% #filter(`Grant Number` == "H80CS00101") %>% 
        mutate(Expandor = ifelse(is.na(YearFirstInZip) | Att != YearFirstInZip, 0, 1)) %>%  #  is.na(YearFirstInZip) | Att != YearFirstInZip
        group_by(`Grant Number`) %>% summarise(Expandor = max(Expandor)), by = c("Grant Number")) %>% 
  left_join(uds.HIT, by = c("Grant Number", "Year"))

df.expand <- df.expand %>% 
  left_join(df.expand %>% filter(SwitchEMR == 1) %>%  
            select(`Grant Number`, Year, SwitchEMR) %>% 
            group_by(`Grant Number`) %>% 
            summarise(FirstSwitch = min(Year)), by = c("Grant Number")) #%>% mutate(SwitchEMR = ifelse(is.na(FirstSwitch), 0, ifelse(Year >= FirstSwitch & FirstSwitch >= 2014, 1, 0)))

df.qual.exp <- df.expand %>% 
  left_join(uds.6b.rates, by = c("Grant Number" = "GrantNumber", "Year")) %>% 
  left_join(uds.6b.denom, by = c("Grant Number" = "GrantNumber", "Year")) %>% 
  #filter(Expandor == 0 & Year %in% c(2015, 2016, 2017, 2018, 2019, 2020) & Att %in% c(0, 2017, 2018, 2019, 2020)) %>%  ## Regular go to analysis
  filter(Expandor == 0 & Year %in% c(2015, 2016, 2017, 2018, 2019) & Att %in% c(0, 2017, 2018, 2019)) %>%  ## Really pronounced causal responses, weak first order stuff
  #filter(Expandor == 0 & Year %in% c(2014, 2015, 2016, 2017, 2018, 2019, 2020, 2021) & Att %in% c(0, 2017, 2018, 2019)) %>%
  left_join(uds.sites %>% 
    group_by(`Grant Number`, Year) %>% 
    summarise(TotalAccess = sum(Access, na.rm = T)), by = c("Grant Number","Year")) %>%
inner_join(uds.sites %>% filter(Year == 2015) %>% select(`Grant Number`, Zip) %>% group_by(`Grant Number`) %>% summarise(Zips = n_distinct(Zip)), by = c("Grant Number")) %>%
  mutate(ZipFilter = ifelse(Zips > 2.75, 2, 1)) %>% 
#inner_join(uds.sites %>% filter(Year == 2014) %>% select(`Grant Number`, Zip) %>% group_by(`Grant Number`) %>% summarise(Zips = n_distinct(Zip)) %>% filter(Zips > 2.75), by = c("Grant Number")) %>%
#inner_join(df.expand %>% filter(Year == 2015) %>% mutate(ZipFilter = ifelse(Total > median(Total), 2, 1)) %>% select(`Grant Number`, ZipFilter), by = c("Grant Number")) %>%
  mutate(l.Sites = log(Sites),
         l.Total = log(Total))

k = 1  # 1 for larger group, 0 for smaller clinic group OOOOOOORRRRRRR 2 for large, 1 for medium, and 0 for small

# df.qual.exp <- inner_join(uds.sites %>% select(`Grant Number`, `Site State`) %>% unique(), df.qual.exp, by = c("Grant Number")) %>%
# mutate(ACA = ifelse((`Site State` %in% c('MI','NH') & Year >= 2014) |
#                     (`Site State` %in% c('PA','IN','AK') & Year >= 2015) |
#                     (`Site State` %in% c('MT','LA') & Year >= 2016) |
#                     (`Site State` %in% c('VA','ME') & Year >= 2019) |
#                     (`Site State` %in% c('ID','UT','NE') & Year >= 2020) |
#                     (`Site State` %in% c('OK','MO') & Year >= 2021) |
#                     (`Site State` %in% c('AZ','AR','CA','CO','CT','DE','DC','HI','IL','IA','KY','MD','MA','MN','NV','NJ','NM','NY','ND','OH','OR','PA','RI','VT','VA','WA') & Year >= 2014), 1, 0)) %>%
#   select(`Grant Number`, ZipFilter, value:ACA)

# %>%
#   group_by(`Grant Number`) %>%
#   filter(max(ACA) == 0 | min(ACA) == 1) %>%
#   ungroup()

df.qual.exp <- df.qual.exp %>% 
  mutate(
    None = None / Total * 100,
         MCR = MCR / Total * 100,
         COM = COM / Total * 100,
         MCD = MCD / Total * 100,

         SwitchEMR = SwitchEMR * 100, 
         eScripts = eScripts* 100, 
         eDecisionSupport = eDecisionSupport * 100, 
         InterOrgHITexchange = InterOrgHITexchange * 100, 
         PatientPortals = PatientPortals * 100, 
         UDSQualRep.highlowuse = UDSQualRep.highlowuse * 100) %>% 
  left_join(uds.covariates, by = c("Grant Number","Year")) %>% 
  left_join(uds.7, by = c("Grant Number","Year")) %>% 
  mutate(Pct.HBW = BirthsGTE2500 / TotalDeliveries * 100, 
         Pct.HTN.CNTRL = Controlcount / Sample * 100) %>%

  left_join(cw %>% select(`Grant Number`, EIN), by = c("Grant Number")) %>% 
  left_join(unique(rbind(chc.finances, chc.fin.2) %>% group_by(ID, year) %>% summarise_all(max)) %>% 
  arrange(ID, year), by = c("EIN" = "ID", "Year" = "year")) %>% 
  mutate(Liabilities = `Liabilities (as pct of assets)` * `Total net assets`, 
         Staffing = `Total expenses before depreciation` * Personnel, 
         LBE.pct = `Gross land, buildings and equipment (LBE)` / `Total net assets`, 
         `Liabilities (as pct of assets)` = `Liabilities (as pct of assets)` * 100, 
         `Surplus as a pct of before dep expenses` = `Surplus as a pct of before dep expenses` * 100, 
         Qcomp = (IMM2yo + Pap + ChildBMI + AdultBMI + Tobac + CRC + DepScrn) / 7, 
         CostOfCare = `Total expenses before depreciation` / Total) %>%

  mutate(Liquidity =  `Total net assets` / (Cash + Investments + Receivables + `Gross land, buildings and equipment (LBE)`), 
         Profitability =  `Unrestricted surplus (deficit) before depreciation` / (Cash + Investments + Receivables + `Gross land, buildings and equipment (LBE)`), 
         Efficiency = (`Total revenue (unrestricted & restricted)` - `Total expenses before depreciation`) / (Cash + Investments + Receivables + `Gross land, buildings and equipment (LBE)`), 
         NetWorth = (100 - `Liabilities (as pct of assets)`) / `Liabilities (as pct of assets)` / 100) %>% 
  mutate(Altmans.Z = 6.56 * Liquidity + 3.26 * Profitability + 6.72 * Efficiency + 1.05 * NetWorth) %>% # 6.56, 3.26, 6.72, 1.05  OR 0.18, 0.3, 0.81, 0.14
  mutate(Liquidity = ifelse(Liquidity == 'Inf', NA, Liquidity), 
         Profitability = ifelse(Profitability == 'Inf', NA, Profitability), 
         Efficiency = ifelse(Efficiency == 'Inf', NA, Efficiency), 
         NetWorth = ifelse(NetWorth == 'Inf', NA, NetWorth), 
         Altmans.Z = ifelse(Altmans.Z == 'Inf', NA, Altmans.Z)) #%>%   filter(!is.na(Altmans.Z) & Altmans.Z != 'Inf')

df.qual.exp %>%
  group_by(Att, Expandor) %>%
  summarise(n() / 6)

df.qual.exp %>% 
  group_by(Tx) %>% 
  summarise(mean(Altmans.Z, na.rm = TRUE), sd(Altmans.Z, na.rm = TRUE))

##########
## Using penetration rate to filter instead of zip codes
##########
#IncomeBelowPoverty

years <- unique(df.qual.exp$Year)

#FQHC penetration rate by zip code
temp1 <- uds.zip %>%
  filter(`Reporting Year` == 2015 ) %>%
  inner_join(uds.sites[!is.na(uds.sites$Zip),] %>%
              filter(Year %in% years), by = c("Grant Number","Zip Code" = "Zip", "Reporting Year" = "Year")) %>%
  select(`Grant Number`, `Reporting Year`, `Zip Code`, `Total Number of Patients`) %>%
  left_join(census, by = c("Zip Code" = "ZipCode","Reporting Year" = "YEAR")) %>% #,"Reporting Year" = "YEAR"
  mutate(MarketShare = `Total Number of Patients` / IncomeBelowPoverty) %>%#IncomeBelowPoverty or TotalPop
  unique()

temp1 %>% 
  filter(`Reporting Year` == 2020)

census %>% filter(YEAR == 2020)
# FQHC weighted average penetration rate conditional on site in zip

temp2 <- temp1 %>%
  mutate(Num = IncomeBelowPoverty * MarketShare) %>%
  group_by(`Grant Number`, `Reporting Year`) %>%
  summarise(Num = sum(Num, na.rm = TRUE)) %>%
  left_join(temp1 %>%
  group_by(`Grant Number`, `Reporting Year`) %>%
  summarise(YTotal = sum(IncomeBelowPoverty, na.rm = TRUE)), by = c("Grant Number","Reporting Year")) %>% #IncomeBelowPoverty
  mutate(MarketShare = Num / YTotal) %>%
  select(`Grant Number`, `Reporting Year`, MarketShare)

df.qual.exp <- df.qual.exp %>%
  select(`Grant Number`, value:Altmans.Z) %>% 
  left_join(temp2, by = c("Grant Number")) %>% #,"Year" = "Reporting Year")) %>%
  #mutate(ZipFilter = ifelse(MarketShare > median(MarketShare, na.rm = FALSE), 2,  # - 0.25
  #                   ifelse(is.na(MarketShare) == T, 1, 1))) %>%
  mutate(ZipFilter = ifelse(MarketShare >= median(MarketShare, na.rm = TRUE), 2, 1)) %>% #median(MarketShare, na.rm = TRUE)
  select(`Grant Number`, ZipFilter, MarketShare, value:Altmans.Z) %>%
  # inner_join(df.qual.exp %>% filter(Year == 2016) %>% mutate(Filter = ifelse(`Surplus as a pct of after dep expenses` < 0, 1, 
  #                                                                     ifelse(`Surplus as a pct of after dep expenses` > 0.03, 2, NA))) %>% select(`Grant Number`, Filter), by = c("Grant Number")) %>% 
  # select(`Grant Number`, ZipFilter, Filter, MarketShare, value:Altmans.Z) %>% 
  #mutate(ZipFilter = Filter) %>% 
  unique()

summary(df.qual.exp$MarketShare)
summary(df.qual.exp$Zips)
summary(df.qual.exp$`Surplus as a pct of after dep expenses`)

df.qual.exp %>% mutate(Treat = ifelse(Att > 0, 'Treat','Control')) %>% group_by(ZipFilter, Treat) %>% summarise(n(), n_distinct(ID))
df.qual.exp %>% filter(`Grant Number` == "H80CS00225")

cnames <- colnames(df.qual.exp)

compare <- read.csv("comeptition.data.final.csv")[,-1]
colnames(compare) <- cnames

# df.qual.exp <- compare
# df.qual.exp <- df.qual.exp %>%
#   #select(`Grant Number`:Year) %>%
#   inner_join(compare %>% select(`Grant Number`, Year, value), by = c("Grant Number","Year")) %>%
#   filter(value.x == value.y) %>%
#   select(`Grant Number`:Altmans.Z) %>%
#   rename(value = value.x)

df.qual.exp <- df.qual.exp %>% rbind(compare %>% filter(`Grant Number` %in% c("H80CS00765","H80CS00881","H80CS02323") & Year %in% unique(df.qual.exp$Year) )) %>% 
  filter(`Grant Number` != "H80CS00225")

uds.sites[!is.na(uds.sites$Zip),] %>% 
  #select(`Grant Number`, Zip, Year) %>% 
  unique() %>% filter(`Grant Number` %in% c("H80CS00881") & Year == 2016)

# df.qual.exp %>% 
#   select(`Grant Number`, Year, value) %>% 
#   full_join(compare %>% select(`Grant Number`, Year, value), by = c("Grant Number","Year")) %>% 
#   filter(is.na(value.x) | is.na(value.y))

#write.csv(df.qual.exp, file = "comeptition.data.final_2.csv") DO NOT UNCOMMENT THIS LINE AS IT WILL OVERWRITE MY BACKUP DATA

#df.qual.exp <- read.csv("comeptition.data.final_2.csv")

## Add in urbanicity
all.density <- uds.sites %>% 
  filter(Year == 2016) %>% 
  select(`Grant Number`, Zip) %>% 
  unique()

zipsize2 <- zctas(cb = TRUE, year = 2010)

density.calc <- all.density %>% 
  left_join(census %>% filter(YEAR == 2016), by = c("Zip" = "ZipCode")) %>% 
  left_join(zipsize2 %>% select(ZCTA5, CENSUSAREA), by = c("Zip" = "ZCTA5")) %>% 
  group_by(`Grant Number`) %>% 
  mutate(num = sum(TotalPop, na.rm = TRUE), denom = sum(CENSUSAREA, na.rm = TRUE)) %>% 
  mutate(Density = num / denom) %>% 
  mutate(Pct.Rural = ifelse(Density < 500, 1, 0), 
         Pct.Urban = ifelse(Density > 2500, 1, 0)) %>% 
  select(`Grant Number`, Density, Pct.Rural, Pct.Urban) %>% 
  unique()

df.qual.exp <- df.qual.exp %>% 
  left_join(density.calc, by = c("Grant Number")) #%>% mutate(ZipFilter = ifelse(Density >= median(Density), 2, 1))

df.expand %>% 
  filter(`Grant Number` == "H80CS00101")

uds.sites %>% filter(`Grant Number` == "H80CS00101") %>% 
  group_by(Zip) %>% 
  summarise(min(Year))
```

## What's happening with the closures?
```{r}

closures <- uds.sites[!is.na(uds.sites$Zip),] %>% 
  select(`Grant Number`, Zip, Year) %>% 
  unique() %>% 
  left_join(clinics.in.zips, by = c("Zip", "Year")) %>% 
  group_by(`Grant Number`, Year) %>% 
  summarise(Clinics = max(Clinics)) %>% 
  pivot_wider(id_cols = `Grant Number`, names_from = Year, values_from = Clinics) %>% 
  filter(`2014` == 1 & `2015` == 1 & `2016` == 1 ) %>% 
  left_join(chc.term, by = c("Grant Number" = "ClinicID")) %>% 
  filter(!is.na(NewClinicID))

closures <- closures %>% pivot_longer(`2014`:`2021`) %>%
  mutate(Outcome = ifelse(!is.na(NewClinicID) & is.na(value), 1, 0))

closures <- closures %>% 
  left_join(closures %>% 
    group_by(`Grant Number`) %>% 
    filter(value > 1) %>% 
    filter(name == min(name)) %>% 
    select(`Grant Number`, name), by = c("Grant Number")) %>% 
  #rename(Year = name.x) %>% 
  mutate(Att = as.numeric(ifelse(is.na(name.y), "0", name.y)), 
         Year = as.numeric(name.x)) %>% 
    select(`Grant Number`, Year, Att, Outcome) %>% 
  rbind( df.qual.exp %>% 
  mutate(Outcome = 0) %>% 
  select(`Grant Number`, Year, Att, Outcome)) %>% 
  mutate(ID = as.numeric(factor(`Grant Number`)), 
         Tx = ifelse(Year >= Att, 1, 0))

summary(lm(data = closures, Outcome ~ Tx + as.factor(Year)))

cs.output.expand <- att_gt( yname = "Outcome",
                      tname = "Year",
                      idname = "ID",
                      gname = "Att",
                      xformla = ~1,
                      data = closures, 
                      control_group = "nevertreated",
                      panel = FALSE, 
                      clustervars = "ID",
                      base_period = "universal"
                      )
aggte(cs.output.expand, type = "simple", na.rm = TRUE)
```

## Play environment
```{r}

cs.output.expand <- att_gt( yname = "Qcomp",
                      tname = "Year",
                      idname = "ID",
                      gname = "Att",
                      xformla = ~Pct.NHB.y + Pct.HL.y + Pct.LessThan100pctFPL.y + Pct.FarmWorker.y + Pct.Homeless.y + Pct.Children.y,
                      data = df.qual.exp %>% 
  left_join(cw %>% left_join(staffing, by = c("EIN")) %>% 
  select(`Grant Number`, Year, Staff) %>% mutate(Year = as.numeric(Year), Staff = as.numeric(Staff)), by = c("Grant Number","Year")) %>% 
    left_join(df.qual.exp %>% filter(Year == 2016) %>% select(`Grant Number`, Pct.NHA:Pct.Children), by = c("Grant Number")) %>% 
    filter(ZipFilter == 1), # / (Cash + Investments + Receivables + `Gross land, buildings and equipment (LBE)`) * 1000), # %>% inner_join(df.qual.exp %>% filter(Year == 2016 & `Surplus as a pct of before dep expenses`< 2) %>% select(`Grant Number`), by = c("Grant Number")),  # %>% select(Year, ID, Att, `Gross land, buildings and equipment (LBE)`) %>% filter(!is.na(`Gross land, buildings and equipment (LBE)`)), ### Split surplus pct after dep at 0 and 0.05 in 2016 plus ZipFilter ==2
                      control_group = "nevertreated",
                      panel = TRUE, 
                      clustervars = "ID",
                      base_period = "universal"
                      )
aggte(cs.output.expand, type = "simple", na.rm = TRUE)

cs.output.expand.agg.dyn.0 <- aggte(cs.output.expand, type = "simple", na.rm = TRUE)
cs.output.expand.agg.dyn <- aggte(cs.output.expand, type = "dynamic", na.rm = TRUE)
data.frame(Year = cs.output.expand.agg.dyn$egt, 
           Estimate = cs.output.expand.agg.dyn$att.egt, 
           CILB = cs.output.expand.agg.dyn$att.egt - cs.output.expand.agg.dyn$crit.val.egt * cs.output.expand.agg.dyn$se.egt,
           CIUB = cs.output.expand.agg.dyn$att.egt + cs.output.expand.agg.dyn$crit.val.egt * cs.output.expand.agg.dyn$se.egt) %>% 
  # mutate(CILB = replace(CILB, Year == -1, NA), 
  #        CIUB = replace(CIUB, Year == -1, NA)) %>% 
  #filter(abs(Year) < 4) %>% 
  ggplot(mapping = aes(x = Year, y = Estimate)) + 
  geom_vline(xintercept=-0.5, linetype = "dashed", color = "red") + 
  geom_hline(yintercept=0, linetype = "dashed", color = "grey") + 
  geom_point() + geom_errorbar(aes(ymin=CILB, ymax=CIUB), width=.2) + theme_tufte_revised()

summary(df.qual.exp[df.qual.exp$Year %in% c(2016:2019),]$`Liabilities (as pct of assets)`)

# Surplus as a pct of after dep expenses
# Liabilities (as pct of assets)
# Qcomp
# Gross land, buildings and equipment (LBE)

df.qual.exp %>%
  #filter(Att == 2020 & Year == 2020) %>%
  select(Att, Year, `Gross land, buildings and equipment (LBE)`, Sites, Altmans.Z, None) %>%
  mutate(Treat = ifelse(Att > 0, 1, 0)) %>%
  group_by(Att, Year) %>%
  summarize(Outcome = mean(None, na.rm = TRUE)) %>%
 # mutate(Outcome = ifelse(Outcome < -5, -5, Outcome)) %>%
  ggplot(aes(x = Year, y = Outcome, group = factor(Att), color = factor(Att))) + geom_point() #+ geom_jitter()

# uds.zip.totals %>% 
#   mutate(LowInc = (None + MCD) / Total) %>% 
#   filter(LowInc < 0.5 & `Reporting Year` == 2019)
#summary(lm(data = df.qual.exp, Qcomp ~ MarketShare + factor(Year) + factor(`Grant Number`)))

#df.qual.exp %>% filter(Year == 2016 & `Surplus as a pct of after dep expenses` < 0 & ZipFilter == 2)

```

## Quick Matching to see if shit changes
```{r}

# df.match <- df.qual.exp %>% filter(Year == 2015) %>%
#   select(`Grant Number`, Att, MCD, Total, Sites, Pct.NHB, Pct.HL, Pct.Children, Pct.LessThan100pctFPL)
# 
# 
# FIPS <- df.match$`Grant Number`
# X <- as.matrix(df.match[,c(3:9)])
# Tr <- ifelse(df.match$Att > 0, 1, 0)
# 
# summary(X)
# 
# m <- Match(Tr = Tr, X = X, M = 1) #, replace = TRUE)
# #mb <- MatchBalance(Treatment ~ . - FIPS, data=df.match, match.out=m, nboots=500)
# 
# df.matched <- data.frame(FIPS[m$index.treated], FIPS[m$index.control]) %>%
#   mutate(Paired = paste(FIPS[m$index.treated], FIPS[m$index.control], sep = "-"))
# 
# df.matched <- rbind(df.matched %>%
#         select(FIPS.m.index.treated., Paired) %>%
#         rename(FIPS = FIPS.m.index.treated.) %>%
#         mutate(Treatment = 1),
#       df.matched %>%
#         select(FIPS.m.index.control., Paired) %>%
#         rename(FIPS = FIPS.m.index.control.) %>%
#         mutate(Treatment = 0)) %>%
#   rename(`Grant Number` = FIPS)
# 
# df.qual.exp <- inner_join(df.matched %>% select(`Grant Number`, Treatment), df.qual.exp, by = c("Grant Number")) # %>% unique()

```

# Figures

## Prep for the robustness tests
```{r}
atts <- c("Primo", rep(0, 6))

# cbind("Sites", round(cbind(cs.output.expand.agg.dyn.0$overall.att, 
#       cs.output.expand.agg.dyn.0$overall.att - 1.96 * cs.output.expand.agg.dyn.0$overall.se, 
#       cs.output.expand.agg.dyn.0$overall.att + 1.96 * cs.output.expand.agg.dyn.0$overall.se, 
#       cs.output.expand.agg.dyn$overall.att, 
#       cs.output.expand.agg.dyn$overall.att - cs.output.expand.agg.dyn$crit.val.egt * cs.output.expand.agg.dyn$overall.se, 
#       cs.output.expand.agg.dyn$overall.att + cs.output.expand.agg.dyn$crit.val.egt * cs.output.expand.agg.dyn$overall.se), 3))
z <- c("Primo", 0) #cs.output.expand$Wpval

```

## Figure 0 - Demographics
```{r}
df.qual.exp %>% 
  mutate(Treat = ifelse(Att > 0, 1, 0)) %>% 
  mutate(Groups = ifelse(Treat == 1 & Expandor == 0, "Treated", 
                  ifelse(Treat == 0 & Expandor == 0, "Control", 
                  ifelse(Treat == 0 & Expandor == 1, "Expandor Comp", "Expandor No Comp")))) %>% 
  filter(Year == 2015) %>%  #z & ZipFilter == 2) %>% 
  select(Sites, Zips, #MarketShare, 
         Total, None:MCD, IMM2yo:DepScrn, Groups, Pct.NHA:Pct.NHW, Pct.LessThan200pctFPL, Pct.Homeless:Pct.Children, `Liabilities (as pct of assets)`, `Gross land, buildings and equipment (LBE)`, `Surplus as a pct of after dep expenses`) %>% #SwitchEMR, InterOrgHITexchange, UDSQualRep.highlowuse, 
  tbl_summary(by = Groups, 
              statistic = all_continuous() ~ "{mean} ({sd})") %>% 
  add_p(test = everything() ~ "t.test")

df.qual.exp %>% mutate(Treat = ifelse(Att > 0, 1, 0)) %>% group_by(Treat) %>% summarise(n())

df.qual.exp %>% 
  filter(Year %in% c(2015, 2016)) %>% 
  mutate(Groups = ifelse(is.na(Personnel), "NoFins", "Fins")) %>% 
 #z & ZipFilter == 2) %>% 
  select(Sites, Zips, #MarketShare, 
         Total, None:MCD, #IMM2yo:DepScrn, 
         Groups, Pct.NHA:Pct.NHW, Pct.LessThan200pctFPL, Pct.Homeless:Pct.Children) %>% #SwitchEMR, InterOrgHITexchange, UDSQualRep.highlowuse, 
  tbl_summary(by = Groups, 
              statistic = all_continuous() ~ "{mean} ({sd})", digits = all_continuous() ~ 2) %>% 
  add_p(test = everything() ~ "t.test")

df.qual.exp %>% 
  mutate(Groups = ifelse(is.na(Personnel), "NoFins", "Fins")) %>% 
  group_by(Groups) %>% 
  summarise(n())

R.version

442*5
37*5
518/1877

442+37

```

## Figure 1 - Trends over time

```{r}
## Plot
p <- uds.sites[!is.na(uds.sites$Zip),] %>% 
  select(`Grant Number`, Zip, Year) %>% 
  unique() %>% 
  left_join(clinics.in.zips, by = c("Zip", "Year")) %>% 
  group_by(`Grant Number`, Year) %>% 
  summarise(Clinics = max(Clinics)) %>% 
  group_by(Year, Clinics) %>% 
  summarise(Count = n()) %>% 
  mutate(CompetitorsInArea = factor(ifelse( Clinics < 4, Clinics - 1, "3+"))) %>% 
  group_by(Year, CompetitorsInArea) %>% 
  summarise(Count = sum(Count))

plot <- p %>% 
  ggplot(aes(x = Year, y = Count, group = CompetitorsInArea, color = CompetitorsInArea)) + 
  geom_point() + geom_line() + theme_minimal() + 
  ggtitle("Figure 1 - Counts of FQHCs by number of competing FQHCs in service area, 2014-2021") + ylab("Number of FQHCs") 
plot + scale_color_discrete(name = "Competitors")

## Table
uds.sites[!is.na(uds.sites$Zip),] %>% 
  select(`Grant Number`, Zip, Year) %>% 
  unique() %>% 
  left_join(clinics.in.zips, by = c("Zip", "Year")) %>% 
  group_by(`Grant Number`, Year) %>% 
  summarise(Clinics = max(Clinics)) %>% 
  group_by(Year, Clinics) %>% 
  summarise(Count = n()) %>% 
  mutate(CompetitorsInArea = factor(ifelse( Clinics < 4, Clinics - 1, "3+"))) %>% 
  group_by(Year, CompetitorsInArea) %>% 
  summarise(Count = sum(Count)) %>% 
  pivot_wider(id_cols = CompetitorsInArea, names_from = Year, values_from = Count)

plot0 <- p %>% 
  arrange(CompetitorsInArea) %>% 
  ggplot(aes(x = Year, y = Count, group = CompetitorsInArea, color = CompetitorsInArea)) + ylab("FQHCs") + 
  geom_point() + geom_line() + theme_tufte_revised() + ggtitle("FQHCs by Number of Rival Clinics") + 
  geom_dl(label=c(rep("0 Competitors", 8), rep("1 Competitor", 8), rep("2 Competitors", 8), rep("3+ Competitors", 8)), 
          method = list("last.points", hjust = 1, vjust = 3, cex = 0.75), inherit.aes=T ) + 
  guides(colour = "none")

plot1 <- uds.sites[!is.na(uds.sites$Zip),] %>% 
            group_by(Zip, Year) %>% 
  summarise(Counts = n()) %>% 
  #summarise(Counts = n_distinct(`Grant Number`)) %>% 
  ungroup() %>% 
  group_by(Year, Counts) %>% 
  summarise(Zips = n()) %>% 
  mutate(Counts.f = as.factor(ifelse(Counts >= 2, "2+", Counts))) %>% 
  group_by(Counts.f, Year) %>% 
  summarise(Zips = sum(Zips)) %>% 
  ggplot(aes(x = Year, y = Zips, group = Counts.f, color = Counts.f)) + 
  geom_point() + geom_line() + theme_tufte_revised() + ggtitle("Zip Codes by Number of FQHC Sites") + 
  geom_dl(label=c(rep("1 Site", 8), rep("2+ Sites", 8)), method = list("last.points", hjust = 1, vjust = -0.5, cex = 0.75), inherit.aes=T ) + 
  guides(colour = "none")

p.desc <- ggarrange(plot1, plot0)

annotate_figure(p.desc, top = text_grob("Figure 2 - FQHC Growth and Density by Zip and Clinic", face = "bold", size = 14))

# rbind(uds.sites %>% 
#   group_by(`Grant Number`, `Site Name`) %>% 
#   summarise(Year = min(Year)) %>% 
#   group_by(Year) %>% 
#   summarise(Counts = n()) %>% 
#   mutate(Type = "Opened"), 
#   uds.sites %>% 
#   group_by(`Grant Number`, `Site Name`) %>% 
#   summarise(Year = max(Year)) %>% 
#   group_by(Year) %>% 
#   summarise(Counts = n()) %>% 
#   mutate(Type = "Closed")) %>% 
#   filter(Year %in% c(2015:2019))

```

```{r}

# df.z %>% 
#   group_by(as.factor(Att), Year) %>% 
#   summarise(outcome = mean(Total.z, na.rm = TRUE)) %>% 
#   ggplot(aes(x = Year, y = outcome, color = `as.factor(Att)`)) + geom_point()

# 
# cs.output.expand <- att_gt( yname = "Total.z",
#                       tname = "Year",
#                       idname = "ID",
#                       gname = "Att",
#                       xformla = ~1,
#                       data = df.z, # %>%  group_by(`Grant Number`, Year, Att) %>% summarise(Zips = n(), ID = min(ID), Total.z = sum(Total.z, na.rm = TRUE) ), # %>% filter(Year > 2015),
#                       control_group = "nevertreated",
#                       clustervars = "ID",
#                       panel = FALSE,
#                       base_period = "universal"
#                       )
# aggte(cs.output.expand, type = "simple", na.rm = TRUE)
# 
# 
# cs.output.expand <- att_gt( yname = "Total.z",
#                       tname = "Year",
#                       idname = "ID",
#                       gname = "Att.z",
#                       xformla = ~1,
#                       data = df.z %>% group_by(`Grant Number`, Year, Att, Att.z) %>% summarise(Zips = n(), ID = min(ID), Total.z = sum(Total.z, na.rm = TRUE) ) %>% filter(Att == Att.z), # %>% filter(Year > 2015),
#                       control_group = "nevertreated",
#                       clustervars = "ID",
#                       panel = FALSE,
#                       base_period = "universal"
#                       )
# aggte(cs.output.expand, type = "simple", na.rm = TRUE)
# 
# #uds.zip %>% filter(`Grant Number` == "H80CS00030" & `Zip Code` == '48264')

```

## Figure 2 - Focused analysis, patient attribution in service area (where they have sites)
```{r}

## Treatment vs control, only in zip code of interest
df.z <- left_join(df.qual.exp #%>% filter(`Grant Number` != "H80CS00225")
                  , #uds.zip %>% mutate(Zip = `Zip Code`) %>% select(`Grant Number`, Zip) %>% unique() , by = c("Grant Number")) %>%
                  uds.sites[!is.na(uds.sites$Zip),] %>%
                    filter(Year %in% c(2015:2019)) %>% 
              select(`Grant Number`, Zip) %>%
              unique() , by = c("Grant Number")) %>%
  left_join(uds.zip %>%
            mutate(  Total.z = `Total Number of Patients`, 
                     MCD.z = `Medicaid_CHIP_Other Public Patients`,
                     MCR.z = `Medicare Patients`,
                     COM.z = `Private Patients`, 
                     None.z = `None_Uninsured Patients`) %>%
            select(`Grant Number`, `Reporting Year`, `Zip Code`, Total.z:None.z), by = c("Grant Number", "Year" = "Reporting Year", "Zip" = "Zip Code")) %>%
  left_join(clinics.in.zips %>% select(Zip:Clinics), by = c("Zip","Year")) %>%
  mutate(Tx.z = ifelse(Clinics == 1 , 0, 
                ifelse(is.na(Clinics), NA, 1)), # | is.na(Clinics)
         ID = as.numeric(factor(paste(`Grant Number`, Zip, sep = "-")))) %>%
  select(`Grant Number`:Year, ZipFilter, Att, ID, Zip:Tx.z) %>%
  arrange(`Grant Number`, Zip, Year) %>% 
  unique() #%>%  inner_join(df.qual.exp %>% filter(Year == 2015 & Total > 12050) %>% select(`Grant Number`), by = c("Grant Number"))  #& Zips < 3.75 or 4.75
  # %>%   filter(MarketShare > median(MarketShare, na.rm = TRUE) - 0.25)

df.z <- df.z %>%
  left_join(df.z %>%
            filter(Tx.z == 1) %>%
            group_by(ID, Zip) %>%
            summarise(Att.z = min(Year)), by = c("ID", "Zip")) %>%
  mutate(Att.z = ifelse(is.na(Att.z), 0, Att.z)) #%>%
# inner_join(df.z %>% filter(!is.na(Clinics)) %>% group_by(ID) %>% summarise(n()) %>% filter(`n()` == length(unique(df.qual.exp$Year))), by = c("ID")) #%>% filter(Att == Att.z)

### NOTE: The aabove commented lines are very important, they decide if only the service area or whole catchment is included in analysis.  
# cs.output.expand <- att_gt( yname = "Outcome",
#                       tname = "Year",
#                       idname = "ID",
#                       gname = "Att",
#                       xformla = ~1,
#                       data = df.z %>% group_by(`Grant Number`, Year, Att) %>% summarise(ID = min(ID), Outcome = sum(MCD.z, na.rm = TRUE) ), # %>% filter(Year > 2015),
#                       control_group = "nevertreated",
#                       clustervars = "ID",
#                       panel = TRUE,
#                       base_period = "universal"
#                       )
# aggte(cs.output.expand, type = "simple", na.rm = TRUE)

## Whole zip as one measure

df.z.full <- df.z %>% ############## NOTE: Original +500 effect was recoved with Att.z, not Att
  select(Zip, ZipFilter, Att.z, Year) %>% 
  unique() %>% 
  left_join(uds.zip, by = c("Zip" = "Zip Code", "Year" = "Reporting Year")) %>% 
  left_join(uds.6b.rates %>% 
            mutate(Quality = (IMM2yo + Pap + ChildBMI + AdultBMI + Tobac + CRC + DepScrn) / 7) %>% 
            select(GrantNumber, Year, Quality), by = c("Grant Number" = "GrantNumber", "Year")) %>% 
  group_by(Zip, ZipFilter, Year, Att.z) %>% 
  summarise(
    Total.z = sum(`Total Number of Patients`, na.rm = TRUE), 
    MCD.z = sum(`Medicaid_CHIP_Other Public Patients`, na.rm = TRUE),
    MCR.z = sum(`Medicare Patients`, na.rm = TRUE),
    COM.z = sum(`Private Patients`, na.rm = TRUE),
    None.z = sum(`None_Uninsured Patients`, na.rm = TRUE), 
    Quality.z = sum(`Total Number of Patients` * Quality, na.rm = TRUE) / Total.z) %>% 
  ungroup() %>% 
  mutate(ID = as.numeric(as.factor(Zip)))# %>% group_by(`Grant Number`, Year, Att) %>% summarise(ID = min(ID), Total.z = sum(Total.z, na.rm = TRUE)) # %>% filter(Year > 2015),

df.zip.1 <- df.z %>% select(ID, ZipFilter, Year, Att, Total.z:None.z)
df.zip.2 <- df.z.full %>% select(ID, ZipFilter, Year, Att.z, Total.z:Quality.z)

## Start the loop

for (k in 1:3){

atts.zip.1 <- rep(0, 6)
atts.zip.2 <- rep(0, 6)

if (k == 1){temp.1.root <- df.zip.1
            temp.2.root <- df.zip.2} else 
  if (k == 2){temp.1.root <- df.zip.1 %>% filter(ZipFilter == 2)
              temp.2.root <- df.zip.2 %>% filter(ZipFilter == 2)} else
    if (k == 3){temp.1.root <- df.zip.1 %>% filter(ZipFilter == 1)
                temp.2.root <- df.zip.2 %>% filter(ZipFilter == 1)}

for (i in 5:9){

  temp.1 <- temp.1.root[,c(1:4, i)]
  colnames(temp.1) <- c("ID", "ZipFilter", "Year","Att","Outcome")

   temp.2 <- temp.2.root[,c(1:4, i)]
  colnames(temp.2) <- c("ID", "ZipFilter", "Year","Att","Outcome")

    # Incumbent only
cs.output.expand <- att_gt( yname = "Outcome",
                      tname = "Year",
                      idname = "ID",
                      gname = "Att",
                      xformla = ~1,
                      data = temp.1,
                      control_group = "nevertreated",
                      clustervars = "ID",
                      panel = FALSE,
                      base_period = "universal"
                      )

cs.output.expand.agg.dyn.0 <- aggte(cs.output.expand, type = "simple", na.rm = TRUE)
cs.output.expand.agg.dyn <- aggte(cs.output.expand, type = "dynamic", na.rm = TRUE)

data.frame(Year = cs.output.expand.agg.dyn$egt,
           Estimate = cs.output.expand.agg.dyn$att.egt,
           CILB = cs.output.expand.agg.dyn$att.egt - cs.output.expand.agg.dyn$crit.val.egt * cs.output.expand.agg.dyn$se.egt,
           CIUB = cs.output.expand.agg.dyn$att.egt + cs.output.expand.agg.dyn$crit.val.egt * cs.output.expand.agg.dyn$se.egt) %>%
  ggplot(mapping = aes(x = Year, y = Estimate)) +
  geom_vline(xintercept=-0.5, linetype = "dashed", color = "red") +
  geom_hline(yintercept=0, linetype = "dashed", color = "grey") +
  geom_point() +
  geom_errorbar(aes(ymin=CILB, ymax=CIUB), width=.2) +
  ggtitle("Total Zip") +
  theme_tufte_revised()

atts.zip.1 <- rbind(atts.zip.1, 
                    cbind(colnames(df.zip.1)[i], 
                          round(mean(temp.1[temp.1$Att == 0 & temp.1$Year == min(temp.1$Year),]$Outcome, na.rm = TRUE), 3), 
                          round(mean(temp.1[temp.1$Att > 0 & temp.1$Year == min(temp.1$Year),]$Outcome, na.rm = TRUE) - mean(temp.1[temp.1$Att == 0 & temp.1$Year == min(temp.1$Year),]$Outcome, na.rm = TRUE), 3), 
                          round(cbind(cs.output.expand.agg.dyn.0$overall.att, 
                          cs.output.expand.agg.dyn.0$overall.att - (1.96 * cs.output.expand.agg.dyn.0$overall.se), 
                          cs.output.expand.agg.dyn.0$overall.att + (1.96 * cs.output.expand.agg.dyn.0$overall.se)), 3)))

# Incumbent + Competitor
cs.output.expand <- att_gt( yname = "Outcome",
                      tname = "Year",
                      idname = "ID",
                      gname = "Att",
                      xformla = ~1,
                      data = temp.2, 
                      control_group = "nevertreated",
                      clustervars = "ID",
                      panel = FALSE,
                      base_period = "universal"
                      )

cs.output.expand.agg.dyn.0 <- aggte(cs.output.expand, type = "simple", na.rm = TRUE)
cs.output.expand.agg.dyn <- aggte(cs.output.expand, type = "dynamic", na.rm = TRUE)

data.frame(Year = cs.output.expand.agg.dyn$egt,
           Estimate = cs.output.expand.agg.dyn$att.egt,
           CILB = cs.output.expand.agg.dyn$att.egt - cs.output.expand.agg.dyn$crit.val.egt * cs.output.expand.agg.dyn$se.egt,
           CIUB = cs.output.expand.agg.dyn$att.egt + cs.output.expand.agg.dyn$crit.val.egt * cs.output.expand.agg.dyn$se.egt) %>%
  ggplot(mapping = aes(x = Year, y = Estimate)) +
  geom_vline(xintercept=-0.5, linetype = "dashed", color = "red") +
  geom_hline(yintercept=0, linetype = "dashed", color = "grey") +
  geom_point() +
  geom_errorbar(aes(ymin=CILB, ymax=CIUB), width=.2) +
  ggtitle("Total Zip") +
  theme_tufte_revised()

atts.zip.2 <- rbind(atts.zip.2, cbind(colnames(df.zip.2)[i], 
                                      round(mean(temp.2[temp.2$Att == 0 & temp.2$Year == min(temp.2$Year),]$Outcome, na.rm = TRUE), 3), 
                                      round(mean(temp.2[temp.2$Att > 0 & temp.2$Year == min(temp.2$Year),]$Outcome, na.rm = TRUE) - mean(temp.2[temp.2$Att == 0 & temp.2$Year == min(temp.2$Year),]$Outcome, na.rm = TRUE), 3), 
                                      round(cbind(cs.output.expand.agg.dyn.0$overall.att,
      cs.output.expand.agg.dyn.0$overall.att - (1.96 * cs.output.expand.agg.dyn.0$overall.se),
      cs.output.expand.agg.dyn.0$overall.att + (1.96 * cs.output.expand.agg.dyn.0$overall.se)), 3)))

}

## For the full sample only

## For Larger vs Smaller FQHCs only
if (k == 1){atts.zip.full <-  cbind(atts.zip.1[-1,], atts.zip.2[-1,]) } else 
  if (k == 2){atts.zip.larger <- cbind(atts.zip.1[-1,], atts.zip.2[-1,])} else
    if (k == 3){atts.zip.smaller <- cbind(atts.zip.1[-1,], atts.zip.2[-1,])}

}

df.zip.1 %>% mutate(Treat = ifelse(Att > 0, 'Treat','Control')) %>% group_by(Treat) %>% summarise(n(), n_distinct(ID))
df.zip.2 %>% mutate(Treat = ifelse(Att.z > 0, 'Treat','Control')) %>% group_by(Treat) %>% summarise(n(), n_distinct(ID))

table.names <- rbind(cbind((df.zip.1 %>% mutate(Treat = ifelse(Att > 0, 'Treat','Control')) %>% group_by(Treat) %>% summarise(n()))[2,2], 
            (df.zip.1 %>% mutate(Treat = ifelse(Att > 0, 'Treat','Control')) %>% group_by(Treat) %>% summarise(n()))[1,2]), 
      cbind((df.zip.2 %>% mutate(Treat = ifelse(Att.z > 0, 'Treat','Control')) %>% group_by(Treat) %>% summarise(n()))[2,2], 
            (df.zip.2 %>% mutate(Treat = ifelse(Att.z > 0, 'Treat','Control')) %>% group_by(Treat) %>% summarise(n()))[1,2]))

## Figure 2 - Full Figure
rbind(atts.zip.full, atts.zip.larger, atts.zip.smaller) %>% 
   kable("html", booktabs = T, col.names = c("Outcome", "C_0", "T_0 - C_0", "DID Estimate", "LB", "UB", "Outcome", "C_0", "T_0 - C_0", "DID Estimate", "LB", "UB")) %>%
  kable_styling(latex_options = c("striped")) %>%
  pack_rows(index = c("Full Sample" = 5, "More Market Share Only" = 5, "Less Market Share Only" = 5)) %>% 
  add_header_above(c(" " = 1, "Incumbent vs Controls (FQHC-Zip units)" = 6, "Incumbent & Competitor vs Controls (zip units)" = 6)) %>% 
  add_footnote(c("C_0 and T_0 are mean values for each sample in control and treatment groups in first year of panel",
                 paste("Full Sample Incumbent vs Control has ", table.names[1,1], "treatment and ", table.names[1,2], " control FQHC-Zip-Years"), 
                 paste("Full Sample Incumbent & Control vs Control has ", table.names[2,1], "treatment and ", table.names[2,2], " control Zip-Years")))

cbind(rep(c("Total","Medicaid","Medicare","Commerical","Uninsured"), 3), rbind(atts.zip.full, atts.zip.larger, atts.zip.smaller)[,2:6]) %>% 
   kable("html", booktabs = T, col.names = c("Outcome", "C_0", "T_0 - C_0", "DID Estimate", "LB", "UB")) %>%
  kable_styling(latex_options = c("striped")) %>%
  row_spec(c(1, 2, 6, 7, 10, 11, 12),bold=T, italic = T, hline_after = T) %>% 
  pack_rows(index = c("Panel A: Full Sample" = 5, "Panel B: Higher Low-Income Market Share Subgroup" = 5, "Panel B: Lower Low-Income Market Share Subgroup" = 5)) %>% 
  add_footnote(c("C_0 and T_0 are mean values for each sample in control and treatment groups in first year of panel",
                 paste("Full Sample Incumbent vs Control has ", table.names[1,1], "treatment and ", table.names[1,2], " control FQHC-Zip-Years"), 
                 paste("Full Sample Incumbent & Control vs Control has ", table.names[2,1], "treatment and ", table.names[2,2], " control Zip-Years")))

```

## Nicer Table

```{r}
df.zip.1 %>% 
  filter(Year %in% c(2016, 2019)) %>% 
  mutate(Tx = ifelse(Att > 0, 1, 0)) %>% 
  group_by(Year, Tx) %>% 
  summarise(mean(Total.z, na.rm = TRUE), 
            mean(MCD.z, na.rm = TRUE), 
            mean(MCR.z, na.rm = TRUE), 
            mean(COM.z, na.rm = TRUE), 
            mean(None.z, na.rm = TRUE)) %>% 
  arrange(Tx, Year) %>% 
  t() %>% 
  as.data.frame() %>% 
  mutate(unadj = (V4 - V3) - (V2 - V1) )

cbind(c("Total","Medicaid","Medicare","Commerical","Uninsured"), atts.zip.full[,2:6]) %>% 
   kable("html", booktabs = T, col.names = c("Outcome", "C_0", "T_0 - C_0", "DID Estimate", "LB", "UB")) %>%
  kable_styling(latex_options = c("striped")) %>%
  row_spec(c(1, 2),bold=T, italic = T, hline_after = T) %>% 
  #pack_rows(index = c("Panel A: Full Sample" = 5, "Panel B: Higher Low-Income Market Share Subgroup" = 5, "Panel B: Lower Low-Income Market Share Subgroup" = 5)) %>% 
  add_footnote(c("C_0 and T_0 are mean values for each sample in control and treatment groups in first year of panel",
                 paste("Full Sample Incumbent vs Control has ", table.names[1,1], "treatment and ", table.names[1,2], " control FQHC-Zip-Years"), 
                 paste("Full Sample Incumbent & Control vs Control has ", table.names[2,1], "treatment and ", table.names[2,2], " control Zip-Years")))
```

## Look at the Market as a whole

```{r}

df.market <- df.z.full %>% #filter(ZipFilter == 1)
                        left_join(df.z %>% select(`Grant Number`,Zip) %>% unique(), by = c("Zip")) %>% 
  mutate(V1 = 1, V2 = ifelse(ZipFilter == 2, 1, 0), V3 = ifelse(ZipFilter == 1, 1, 0)) %>% 
                       select(Year, ID, Att.z, Quality.z, `Grant Number`, Total.z:None.z, V1:V3) %>%
  # inner_join(df.qual.exp %>% filter(Year == 2015) %>%
  #                                           mutate(ZipFilter = ifelse(Total < 12000, 1,
  #                                                              ifelse(Total > 12000, 2, NA))) %>% select(`Grant Number`, ZipFilter), by = c("Grant Number")) %>%
  inner_join(df.qual.exp %>% filter(Year == 2016) %>% mutate(ZipFilter = ifelse(`Surplus as a pct of before dep expenses` < 2, 1, 2)) %>% select(`Grant Number`, ZipFilter), by = c("Grant Number")) %>% 
  select(Year, ZipFilter, ID, Att.z, Quality.z, Total.z:None.z, V1:V3) %>% 
  mutate(V4 = ifelse(ZipFilter == 2, 1, 0), V5 = ifelse(ZipFilter == 1, 1, 0)) %>% 
  unique()

### Incumbnet + Competitors, Quality 
cs.output.expand <- att_gt( yname = "Quality.z",
                      tname = "Year",
                      idname = "ID",
                      gname = "Att.z",
                      xformla = ~1,
                      data = df.market %>% filter(V5 == 1), 
                      control_group = "nevertreated",
                      clustervars = "ID",
                      panel = TRUE,
                      base_period = "universal"
                      )

cs.output.expand.agg.dyn.0 <- aggte(cs.output.expand, type = "simple", na.rm = TRUE)
cs.output.expand.agg.dyn <- aggte(cs.output.expand, type = "dynamic", na.rm = TRUE)

data.frame(Year = cs.output.expand.agg.dyn$egt,
           Estimate = cs.output.expand.agg.dyn$att.egt,
           CILB = cs.output.expand.agg.dyn$att.egt - cs.output.expand.agg.dyn$crit.val.egt * cs.output.expand.agg.dyn$se.egt,
           CIUB = cs.output.expand.agg.dyn$att.egt + cs.output.expand.agg.dyn$crit.val.egt * cs.output.expand.agg.dyn$se.egt) %>%
  ggplot(mapping = aes(x = Year, y = Estimate)) +
  geom_vline(xintercept=-0.5, linetype = "dashed", color = "red") +
  geom_hline(yintercept=0, linetype = "dashed", color = "grey") +
  geom_point() +
  geom_errorbar(aes(ymin=CILB, ymax=CIUB), width=.2) +
  ggtitle("Total Change in Market Quality") +
  theme_tufte_revised()

```

## Figure 2 plots that are nicer
```{r}

## Plot of this shit, all payors
f2.picture.full <- rbind(rbind(atts.zip.full, atts.zip.larger, atts.zip.smaller)[,c(1:6)]) %>%  #, rbind(atts.zip.full, atts.zip.larger, atts.zip.smaller)[,c(7:12)]) %>% 
  as.data.frame() %>% 
  mutate(Model = c(rep("Incumbent Only", 3*5)), #, rep("Incumbent & Competitor", 3*5)), 
         Sample = rep(c(rep("Full Sample", 5), rep("More Market Share", 5), rep("Less Market Share", 5)), 1))

#### Clean up the data
row.names(f2.picture.full) <- NULL
colnames(f2.picture.full) <- c("Payor", "C_0", "T_0 - C_0", "Estimate", "LB", "UB", "Model","Sample")
f2.picture.full[,4:6] <- sapply(f2.picture.full[,4:6], as.numeric)
f2.picture.full$Sample <- factor(f2.picture.full$Sample, levels = c("Full Sample", "More Market Share","Less Market Share"))
f2.picture.full$Payor <- gsub(f2.picture.full$Payor, pattern = ".z", replacement = "")
f2.picture.full$Payor <- factor(f2.picture.full$Payor, levels = c("None", "MCR","COM","MCD","Total"))
f2.picture.full$Model <- factor(f2.picture.full$Model, levels = c("Incumbent Only","Incumbent & Competitor"))
f2.picture.full$Total <- ifelse(f2.picture.full$Payor == "Total", "Total", "Payor")
f2.picture.full$Total <- factor(f2.picture.full$Total, levels = c("Total","Payor"))

### Plot
f2.p.totals.inc <- ggplot(data=f2.picture.full #%>% filter(Sample == "Full Sample")
                          , aes(y=Payor, x=Estimate, xmin=LB, xmax=UB)) +
geom_point() +

#this changes the features of the overall effects
#one could resize, reshape, or recolor these points if desired
xlab("DID Estimates - Patients") + 
ylab(NULL) + 
#adds the CIs
geom_errorbarh(height=.1) +

#sets the scales
#note that I reverse the y axis to correctly order the effect #sizes based on my index variable
#scale_x_continuous(limits=c(-2.5,1), breaks = c(-2.5:1), name=xname)+
#scale_y_continuous(name = "", breaks=1:9, labels = t9.forest$X1, trans="reverse")+

#adding a vertical line at the effect = 0 mark
geom_vline(xintercept=0, color="black", linetype="dashed", alpha=.5)+

#faceting based on my subgroups
  #facet_wrap(~X1) +
facet_grid(Total~Sample, scales= "free_y", switch = 'y', space = "free_y")+

#thematic stuff
#ggtitle("Figure 3 - ATT Estimates by SubGroups")+
theme_tufte_revised()+
   theme(strip.placement = "outside") +
#theme(text=element_text(family="Times",size=12, color="black"))+
theme(panel.spacing = unit(1, "lines")) +
  ggtitle("Figure 3, CS-DID Estimates on Patient Panel and Payor (FQHC-zip-year units)")

f2.p.totals.inc #+ theme(panel.spacing = unit(1, "lines"),  title =  element_text(face = "bold", size = 12))

```

## Figure 3 - Primary Results

```{r}

df.sum <- df.qual.exp %>% #filter(`Grant Number` != "H80CS00225") %>%

  ungroup() %>% 
  #filter(MarketShare < median(MarketShare, na.rm = TRUE) - 0.25) %>% 
  #inner_join(df.qual.exp %>% filter(Year == 2015 & Total < 12000) %>% select(`Grant Number`), by = c("Grant Number")) %>%  #& Zips < 3.75 or 4.75
  # select(`Grant Number`, Year, ID, Att, Sites, Qcomp, Altmans.Z) %>%
  # inner_join(df.qual.exp %>% filter(Year == 2015) %>% mutate(ZipFilter = ifelse(Zips < 1.75, 1, 
  #                                                                        ifelse(Zips > 4.75, 2, NA))) %>% select(`Grant Number`, ZipFilter), by = c("Grant Number")) %>%
  select(Year, ZipFilter, ID, Att, Sites, Qcomp, Altmans.Z)

for (k in 1:3){

atts.sum <- rep(0, 7)
fig3.sum <- rep(0, 5)

if (k ==1){df.sum.root <- df.sum} else
  if (k == 2){df.sum.root <- df.sum %>% filter(ZipFilter == 2)} else
    if (k == 3){df.sum.root <- df.sum %>% filter(ZipFilter == 1)}

for (i in 5:ncol(df.sum)){

  temp <- df.sum.root[,c(1:4, i)]
  colnames(temp) <- c("Year", "ZipFilter", "ID","Att", "Outcome")
  temp$Outcome

  # temp <- temp %>%
  #   mutate(Outcome = ifelse(Outcome < quantile(temp$Outcome, probs = c(0.05, 0.95), na.rm = TRUE)[1], quantile(temp$Outcome, probs = c(0.05, 0.95), na.rm = TRUE)[1],
  #                    ifelse(Outcome > quantile(temp$Outcome, probs = c(0.05, 0.95), na.rm = TRUE)[2], quantile(temp$Outcome, probs = c(0.05, 0.95), na.rm = TRUE)[2], Outcome)))

cs.output.expand <- att_gt( yname = "Outcome",
                      tname = "Year",
                      idname = "ID",
                      gname = "Att",
                      xformla = ~1,
                      data = temp, # %>% filter(Year >= 2015 & Year <= 2020), 
                      control_group = "nevertreated",
                      clustervars = "ID",
                      panel = FALSE, 
                      base_period = "universal"
                      )
cs.output.expand.agg.dyn.0 <- aggte(cs.output.expand, type = "simple", na.rm = TRUE)
cs.output.expand.agg.dyn <- aggte(cs.output.expand, type = "dynamic", na.rm = TRUE)

atts.sum <- rbind(atts.sum, cbind(colnames(df.sum)[i], round(mean(temp[temp$Att == 0 & temp$Year == min(temp$Year),]$Outcome, na.rm = TRUE), 3), round(mean(temp[temp$Att > 0 & temp$Year == min(temp$Year),]$Outcome, na.rm = TRUE) - mean(temp[temp$Att == 0 & temp$Year == min(temp$Year),]$Outcome, na.rm = TRUE), 3), round(cbind(cs.output.expand.agg.dyn.0$overall.att, 
      cs.output.expand.agg.dyn.0$overall.att - 1.96 * cs.output.expand.agg.dyn.0$overall.se, 
      cs.output.expand.agg.dyn.0$overall.att + 1.96 * cs.output.expand.agg.dyn.0$overall.se), 3)))

atts <- rbind(atts, cbind(colnames(df.sum)[i], round(cbind(cs.output.expand.agg.dyn.0$overall.att, 
      cs.output.expand.agg.dyn.0$overall.att - 1.96 * cs.output.expand.agg.dyn.0$overall.se, 
      cs.output.expand.agg.dyn.0$overall.att + 1.96 * cs.output.expand.agg.dyn.0$overall.se, 
      cs.output.expand.agg.dyn$overall.att, 
      cs.output.expand.agg.dyn$overall.att - cs.output.expand.agg.dyn$crit.val.egt * cs.output.expand.agg.dyn$overall.se, 
      cs.output.expand.agg.dyn$overall.att + cs.output.expand.agg.dyn$crit.val.egt * cs.output.expand.agg.dyn$overall.se), 3)))
z <- rbind(z, c(colnames(df.sum)[i], cs.output.expand$Wpval))

fig3.sum <- rbind(fig3.sum, 
                 data.frame(Outcome = colnames(df.sum)[i], 
                      Year = cs.output.expand.agg.dyn$egt, 
           Estimate = cs.output.expand.agg.dyn$att.egt, 
           CILB = cs.output.expand.agg.dyn$att.egt - cs.output.expand.agg.dyn$crit.val.egt * cs.output.expand.agg.dyn$se.egt,
           CIUB = cs.output.expand.agg.dyn$att.egt + cs.output.expand.agg.dyn$crit.val.egt * cs.output.expand.agg.dyn$se.egt))
}

rownames(fig3.sum) <- NULL
if (k == 1){atts.sum.full <- atts.sum[-1,]} else 
 if (k == 2){atts.sum.larger <- atts.sum[-1,]} else 
   {atts.sum.smaller <- atts.sum[-1,]}

}

#Event Study Plots
fig3.sum[-1,] %>%
  ggplot(mapping = aes(x = Year, y = Estimate)) +
  facet_wrap(~Outcome, scales = "free_y") +
  ylab("Percentage Points") +
  geom_vline(xintercept=-0.5, linetype = "dashed", color = "red") +
  geom_hline(yintercept=0, linetype = "dashed", color = "grey") +
  geom_point() + geom_errorbar(aes(ymin=CILB, ymax=CIUB), width=.2) +
  ggtitle("fig3 - Event Study") +
  theme_tufte_revised()

table.names <- rbind(cbind((df.sum %>% mutate(Treat = ifelse(Att > 0, 'Treat','Control')) %>% group_by(Treat) %>% summarise(n()))[2,2], 
            (df.sum %>% mutate(Treat = ifelse(Att > 0, 'Treat','Control')) %>% group_by(Treat) %>% summarise(n()))[1,2]), 
      cbind((df.sum %>% mutate(Treat = ifelse(Att > 0, 'Treat','Control')) %>% group_by(ZipFilter, Treat) %>% summarise(n()))[4,3], 
            (df.sum %>% mutate(Treat = ifelse(Att > 0, 'Treat','Control')) %>% group_by(ZipFilter, Treat) %>% summarise(n()))[3,3]), 
      cbind((df.sum %>% mutate(Treat = ifelse(Att > 0, 'Treat','Control')) %>% group_by(ZipFilter, Treat) %>% summarise(n()))[2,3], 
            (df.sum %>% mutate(Treat = ifelse(Att > 0, 'Treat','Control')) %>% group_by(ZipFilter, Treat) %>% summarise(n()))[1,3]))

## Table
rbind(atts.sum.full, atts.sum.larger, atts.sum.smaller) %>% 
   kable("html", booktabs = T, col.names = c("Outcome", "C_0", "T_0 - C_0", "DID Estimate", "LB", "UB"), caption = "Fig 3 - FQHC Responses by Market Share") %>%
  kable_styling(latex_options = c("striped")) %>%
  pack_rows(index = c("Full" = 3, "Above Median Market Share FQHCs" = 3, "Below Median Market Share FQHCs" = 3)) %>% 
  add_footnote(c("Atlman's Z is a measure of Financial Distress",
                 "C_0 and T_0 are mean values for each sample in control and treatment groups in first year of panel",
                 paste("Full Sample has ", table.names[1,1], "treatment and ", table.names[1,2], " control FQHC-Years")))

## Better Figure
f3.picture.full <- rbind(cbind("Full Sample", atts.sum.full), cbind("More Market Share", atts.sum.larger), cbind("Less Market Share", atts.sum.smaller) ) %>% 
  as.data.frame()

### Clean up the data
row.names(f3.picture.full) <- NULL
colnames(f3.picture.full) <- c("MarketShare", "Measure", "C_0", "T_0 - C_0", "Estimate", "LB", "UB")
f3.picture.full[,3:7] <- sapply(f3.picture.full[,3:7], as.numeric)
f3.picture.full$MarketShare <- factor(f3.picture.full$MarketShare, levels = c("Full Sample", "More Market Share","Less Market Share"))
# f3.picture.full$Payor <- gsub(f3.picture.full$Payor, pattern = ".z", replacement = "")
# f3.picture.full$Payor <- factor(f3.picture.full$Payor, levels = c("None", "MCR","COM","MCD","Total"))
# f3.picture.full$Model <- factor(f3.picture.full$Model, levels = c("Incumbent Only","Incumbent & Competitor"))
# f3.picture.full$Total <- ifelse(f3.picture.full$Payor == "Total", "Total", "Payor")
# f3.picture.full$Total <- factor(f3.picture.full$Total, levels = c("Total","Payor"))

### Plot
f3.p.totals.full <- ggplot(data=f3.picture.full, aes(y=Measure, x=Estimate, xmin=LB, xmax=UB)) +
geom_point() +

#this changes the features of the overall effects
#one could resize, reshape, or recolor these points if desired
xlab("DID Estimates") + 
ylab(NULL) + 
#adds the CIs
geom_errorbarh(height=.1) +

#sets the scales
#note that I reverse the y axis to correctly order the effect #sizes based on my index variable
#scale_x_continuous(limits=c(-2.5,1), breaks = c(-2.5:1), name=xname)+
#scale_y_continuous(name = "", breaks=1:9, labels = t9.forest$X1, trans="reverse")+

#adding a vertical line at the effect = 0 mark
geom_vline(xintercept=0, color="black", linetype="dashed", alpha=.5)+

#faceting based on my subgroups
  #facet_wrap(~X1) +
facet_grid(~MarketShare, scales= "free", switch = 'y', space = "free_y")+

#thematic stuff
#ggtitle("Figure 3 - ATT Estimates by SubGroups")+
theme_tufte_revised()+
   theme(strip.placement = "outside") +
#theme(text=element_text(family="Times",size=12, color="black"))+
theme(panel.spacing = unit(1, "lines")) +
  ggtitle("Figure 3, FQHC Performance Reactions by Market Share")

f3.p.totals.full

```

## Figure 4 - Split by profitability
```{r}

df.sum <- df.qual.exp %>% #filter(`Grant Number` != "H80CS00225") %>% 
  ungroup() %>% 
  #rename(`Grant Number` = Grant.Number, `Surplus as a pct of before dep expenses` = Surplus.as.a.pct.of.after.dep.expenses) %>% 
  #filter(MarketShare < 0.4) %>%  #median(MarketShare, na.rm = TRUE) - 0.25
  #inner_join(df.qual.exp %>% filter(Year == 2015 & Total < 15000) %>% select(`Grant Number`), by = c("Grant Number")) %>%  #& Zips < 1.75 or 7.75
  select(`Grant Number`, Year, ID, Att, Sites, Qcomp, Altmans.Z) %>%
  # inner_join(df.qual.exp %>% filter(Year == 2015) %>%
  #                                           mutate(ZipFilter = ifelse(Total < 12000, 1,
  #                                                              ifelse(Total > 12000, 2, NA))) %>% select(`Grant Number`, ZipFilter), by = c("Grant Number")) %>%
  inner_join(df.qual.exp %>% filter(Year == 2016) %>% mutate(ZipFilter = ifelse(`Surplus as a pct of before dep expenses` < 2, 1, 2)) %>% select(`Grant Number`, ZipFilter), by = c("Grant Number")) %>% 
  select(Year, ZipFilter, ID, Att, Sites, Qcomp, Altmans.Z) %>% 
  unique()

#summary(df.qual.exp[df.qual.exp$Year == 2015,]$Qcomp)

for (k in 2:3){

atts.sum <- rep(0, 7)
fig4.sum <- rep(0, 5)

if (k ==1){df.sum.root <- df.sum} else
  if (k == 2){df.sum.root <- df.sum %>% filter(ZipFilter == 2)} else
    if (k == 3){df.sum.root <- df.sum %>% filter(ZipFilter == 1)}

for (i in 5:ncol(df.sum)){

  temp <- df.sum.root[,c(1:4, i)]
  colnames(temp) <- c("Year", "ZipFilter", "ID","Att", "Outcome")
  temp$Outcome

  # temp <- temp %>%
  #   mutate(Outcome = ifelse(Outcome < quantile(temp$Outcome, probs = c(0.05, 0.95), na.rm = TRUE)[1], quantile(temp$Outcome, probs = c(0.05, 0.95), na.rm = TRUE)[1],
  #                    ifelse(Outcome > quantile(temp$Outcome, probs = c(0.05, 0.95), na.rm = TRUE)[2], quantile(temp$Outcome, probs = c(0.05, 0.95), na.rm = TRUE)[2], Outcome)))

cs.output.expand <- att_gt( yname = "Outcome",
                      tname = "Year",
                      idname = "ID",
                      gname = "Att",
                      xformla = ~1,
                      data = temp , 
                      control_group = "nevertreated",
                      clustervars = "ID",
                      panel = FALSE, 
                      base_period = "universal"
                      )
cs.output.expand.agg.dyn.0 <- aggte(cs.output.expand, type = "simple", na.rm = TRUE)
cs.output.expand.agg.dyn <- aggte(cs.output.expand, type = "dynamic", na.rm = TRUE)

atts.sum <- rbind(atts.sum, cbind(colnames(df.sum)[i], round(mean(temp[temp$Att == 0 & temp$Year == min(temp$Year),]$Outcome, na.rm = TRUE), 3), round(mean(temp[temp$Att > 0 & temp$Year == min(temp$Year),]$Outcome, na.rm = TRUE) - mean(temp[temp$Att == 0 & temp$Year == min(temp$Year),]$Outcome, na.rm = TRUE), 3), round(cbind(cs.output.expand.agg.dyn.0$overall.att, 
      cs.output.expand.agg.dyn.0$overall.att - 1.96 * cs.output.expand.agg.dyn.0$overall.se, 
      cs.output.expand.agg.dyn.0$overall.att + 1.96 * cs.output.expand.agg.dyn.0$overall.se), 3)))

atts <- rbind(atts, cbind(colnames(df.sum)[i], round(cbind(cs.output.expand.agg.dyn.0$overall.att, 
      cs.output.expand.agg.dyn.0$overall.att - 1.96 * cs.output.expand.agg.dyn.0$overall.se, 
      cs.output.expand.agg.dyn.0$overall.att + 1.96 * cs.output.expand.agg.dyn.0$overall.se, 
      cs.output.expand.agg.dyn$overall.att, 
      cs.output.expand.agg.dyn$overall.att - cs.output.expand.agg.dyn$crit.val.egt * cs.output.expand.agg.dyn$overall.se, 
      cs.output.expand.agg.dyn$overall.att + cs.output.expand.agg.dyn$crit.val.egt * cs.output.expand.agg.dyn$overall.se), 3)))
z <- rbind(z, c(colnames(df.sum)[i], cs.output.expand$Wpval))

fig4.sum <- rbind(fig4.sum, 
                 data.frame(Outcome = colnames(df.sum)[i], 
                      Year = cs.output.expand.agg.dyn$egt, 
           Estimate = cs.output.expand.agg.dyn$att.egt, 
           CILB = cs.output.expand.agg.dyn$att.egt - cs.output.expand.agg.dyn$crit.val.egt * cs.output.expand.agg.dyn$se.egt,
           CIUB = cs.output.expand.agg.dyn$att.egt + cs.output.expand.agg.dyn$crit.val.egt * cs.output.expand.agg.dyn$se.egt))
}

rownames(fig4.sum) <- NULL
if (k == 1){atts.sum.full <- atts.sum[-1,]} else 
 if (k == 2){atts.sum.larger <- atts.sum[-1,]} else 
   {atts.sum.smaller <- atts.sum[-1,]}

}

#Event Study Plots
fig4.sum[-1,] %>%
  ggplot(mapping = aes(x = Year, y = Estimate)) +
  facet_wrap(~Outcome, scales = "free_y") +
  ylab("Percentage Points") +
  geom_vline(xintercept=-0.5, linetype = "dashed", color = "red") +
  geom_hline(yintercept=0, linetype = "dashed", color = "grey") +
  geom_point() + geom_errorbar(aes(ymin=CILB, ymax=CIUB), width=.2) +
  ggtitle("fig4 - Event Study") +
  theme_tufte_revised()

table.names <- rbind(cbind((df.sum %>% mutate(Treat = ifelse(Att > 0, 'Treat','Control')) %>% group_by(Treat) %>% summarise(n()))[2,2], 
            (df.sum %>% mutate(Treat = ifelse(Att > 0, 'Treat','Control')) %>% group_by(Treat) %>% summarise(n()))[1,2]), 
      cbind((df.sum %>% mutate(Treat = ifelse(Att > 0, 'Treat','Control')) %>% group_by(ZipFilter, Treat) %>% summarise(n()))[4,3], 
            (df.sum %>% mutate(Treat = ifelse(Att > 0, 'Treat','Control')) %>% group_by(ZipFilter, Treat) %>% summarise(n()))[3,3]), 
      cbind((df.sum %>% mutate(Treat = ifelse(Att > 0, 'Treat','Control')) %>% group_by(ZipFilter, Treat) %>% summarise(n()))[2,3], 
            (df.sum %>% mutate(Treat = ifelse(Att > 0, 'Treat','Control')) %>% group_by(ZipFilter, Treat) %>% summarise(n()))[1,3]))

## Table
rbind(atts.sum.larger, atts.sum.smaller) %>% 
   kable("html", booktabs = T, col.names = c("Outcome", "C_0", "T_0 - C_0", "DID Estimate", "LB", "UB")) %>%
  kable_styling(latex_options = c("striped")) %>%
  pack_rows(index = c("Profitable FQHCs" = 3, "Unprofitable/Budget Neutral FQHCs" = 3)) %>% 
  add_footnote(c("Atlman's Z is a measure of Financial Distress",
                 "C_0 and T_0 are mean values for each sample in control and treatment groups in first year of panel",
                 paste("Profitable FQHCs Sample has ", table.names[2,1], "treatment and ", table.names[2,2], " control FQHC-Years"), 
                 paste("Unprofitable and Budget Neutral FQHCs Sample has ", table.names[3,1], "treatment and ", table.names[3,2], " control FQHC-Years")))

## Better Figure
f4.picture.full <- rbind(cbind("Full Sample", atts.sum.full), cbind("Profitable", atts.sum.larger), cbind("Unprofitable/Budget Neutral", atts.sum.smaller) ) %>% 
  as.data.frame()

### Clean up the data
row.names(f4.picture.full) <- NULL
colnames(f4.picture.full) <- c("Profitability", "Measure", "C_0", "T_0 - C_0", "Estimate", "LB", "UB")
f4.picture.full[,3:7] <- sapply(f4.picture.full[,3:7], as.numeric)
# f4.picture.full$Sample <- factor(f4.picture.full$Sample, levels = c("Full Sample", "More Market Share","Less Market Share"))
# f4.picture.full$Payor <- gsub(f4.picture.full$Payor, pattern = ".z", replacement = "")
# f4.picture.full$Payor <- factor(f4.picture.full$Payor, levels = c("None", "MCR","COM","MCD","Total"))
# f4.picture.full$Model <- factor(f4.picture.full$Model, levels = c("Incumbent Only","Incumbent & Competitor"))
# f4.picture.full$Total <- ifelse(f4.picture.full$Payor == "Total", "Total", "Payor")
# f4.picture.full$Total <- factor(f4.picture.full$Total, levels = c("Total","Payor"))

### Plot
f4.p.totals.full <- ggplot(data=f4.picture.full, aes(y=Measure, x=Estimate, xmin=LB, xmax=UB)) +
geom_point() +

#this changes the features of the overall effects
#one could resize, reshape, or recolor these points if desired
xlab("DID Estimates") + 
ylab(NULL) + 
#adds the CIs
geom_errorbarh(height=.1) +

#sets the scales
#note that I reverse the y axis to correctly order the effect #sizes based on my index variable
#scale_x_continuous(limits=c(-2.5,1), breaks = c(-2.5:1), name=xname)+
#scale_y_continuous(name = "", breaks=1:9, labels = t9.forest$X1, trans="reverse")+

#adding a vertical line at the effect = 0 mark
geom_vline(xintercept=0, color="black", linetype="dashed", alpha=.5)+

#faceting based on my subgroups
  #facet_wrap(~X1) +
facet_grid(~Profitability, scales= "free", switch = 'y', space = "free_y")+

#thematic stuff
#ggtitle("Figure 3 - ATT Estimates by SubGroups")+
theme_tufte_revised()+
   theme(strip.placement = "outside") +
#theme(text=element_text(family="Times",size=12, color="black"))+
theme(panel.spacing = unit(1, "lines")) +
  ggtitle("Figure 4, FQHC Performance Reactions by Profitability")

f4.p.totals.full
```

## True Figure 3 - Responses
```{r}

fig3 <- rbind(f3.picture.full %>% rename(Subgroup = MarketShare), f4.picture.full[-c(1:3),] %>% rename(Subgroup = Profitability)) %>% 
  mutate(Measure = ifelse(Measure == "Sites", "Number of Sites", 
                   ifelse(Measure == "Qcomp", "Quality of Care", 
                   ifelse(Measure == "Altmans.Z", "Financial Stability", NA))))

f3.full <- ggplot(data=fig3[1:3,], aes(y=Measure, x=Estimate, xmin=LB, xmax=UB)) +
  geom_point() +

#this changes the features of the overall effects
#one could resize, reshape, or recolor these points if desired
xlab("DID Estimates") + 
ylab(NULL) + 
#adds the CIs
geom_errorbarh(height=.1) +

#sets the scales
#note that I reverse the y axis to correctly order the effect #sizes based on my index variable
#scale_x_continuous(limits=c(-2.5,1), breaks = c(-2.5:1), name=xname)+
#scale_y_continuous(name = "", breaks=1:9, labels = t9.forest$X1, trans="reverse")+

#adding a vertical line at the effect = 0 mark
geom_vline(xintercept=0, color="black", linetype="dashed", alpha=.5)+

#faceting based on my subgroups
  #facet_wrap(~X1) +
facet_grid(~Subgroup, scales= "free", switch = 'y', space = "free_y")+

#thematic stuff
#ggtitle("Figure 3 - ATT Estimates by SubGroups")+
theme_tufte_revised()+
   theme(strip.placement = "outside") +
#theme(text=element_text(family="Times",size=12, color="black"))+
theme(panel.spacing = unit(1, "lines"))

f3.mark <- ggplot(data=fig3[4:9,], aes(y=Measure, x=Estimate, xmin=LB, xmax=UB)) +
  geom_point() +

#this changes the features of the overall effects
#one could resize, reshape, or recolor these points if desired
xlab("DID Estimates") + 
ylab(NULL) + 
#adds the CIs
geom_errorbarh(height=.1) +

#sets the scales
#note that I reverse the y axis to correctly order the effect #sizes based on my index variable
#scale_x_continuous(limits=c(-2.5,1), breaks = c(-2.5:1), name=xname)+
#scale_y_continuous(name = "", breaks=1:9, labels = t9.forest$X1, trans="reverse")+

#adding a vertical line at the effect = 0 mark
geom_vline(xintercept=0, color="black", linetype="dashed", alpha=.5)+

#faceting based on my subgroups
  #facet_wrap(~X1) +
facet_grid(~Subgroup, scales= "free", switch = 'y', space = "free_y")+

#thematic stuff
#ggtitle("Figure 3 - ATT Estimates by SubGroups")+
theme_tufte_revised()+
   theme(strip.placement = "outside") +
#theme(text=element_text(family="Times",size=12, color="black"))+
theme(panel.spacing = unit(1, "lines"))

f3.prof <- ggplot(data=fig3[10:15,], aes(y=Measure, x=Estimate, xmin=LB, xmax=UB)) +
  geom_point() +

#this changes the features of the overall effects
#one could resize, reshape, or recolor these points if desired
xlab("DID Estimates") + 
ylab(NULL) + 
#adds the CIs
geom_errorbarh(height=.1) +

#sets the scales
#note that I reverse the y axis to correctly order the effect #sizes based on my index variable
#scale_x_continuous(limits=c(-2.5,1), breaks = c(-2.5:1), name=xname)+
#scale_y_continuous(name = "", breaks=1:9, labels = t9.forest$X1, trans="reverse")+

#adding a vertical line at the effect = 0 mark
geom_vline(xintercept=0, color="black", linetype="dashed", alpha=.5)+

#faceting based on my subgroups
  #facet_wrap(~X1) +
facet_grid(~Subgroup, scales= "free", switch = 'y', space = "free_y")+

#thematic stuff
#ggtitle("Figure 3 - ATT Estimates by SubGroups")+
theme_tufte_revised()+
   theme(strip.placement = "outside") +
#theme(text=element_text(family="Times",size=12, color="black"))+
theme(panel.spacing = unit(1, "lines"))

gt <- grid.arrange(f3.full,                            
             f3.mark, f3.prof,                         
             ncol = 3, nrow = 2, 
             layout_matrix = rbind(c(1,  2, 2), c(1,  3, 3)))

# Add labels to the arranged plots
p <- as_ggplot(gt) +                                # transform to a ggplot
 draw_plot_label(label = c("Panel A", "Panel B", "Panel C"), size = 15,
                 x = c(0, 0.33, 0.33), y = c(1, 1, 0.5)) # Add labels
p

```

## Figure for Paper

```{r}
df.qual.exp %>% 
  filter(Year %in% c(2016, 2019)) %>% 
  mutate(Tx = ifelse(Att > 0, 1, 0)) %>% 
  group_by(Year, Tx) %>% 
  summarise(mean(Sites, na.rm = TRUE), 
            mean(Qcomp, na.rm = TRUE), 
            mean(Altmans.Z, na.rm = TRUE), 
            mean(Efficiency, na.rm = TRUE)) %>% 
  arrange(Tx, Year) %>% 
  t() %>% 
  as.data.frame() %>% 
  mutate(unadj = (V4 - V3) - (V2 - V1) )
```

## Figure for Poster

```{r}

poster.fig3 <- cbind(Group = c(rep("Panel A", 3),rep("Panel B", 6),rep("Panel C", 6)), fig3) %>% 
  mutate(Measure = as.factor(Measure)) %>% 
  mutate(Measure = fct_reorder(Measure, desc(Measure))) %>%

ggplot(aes(y=Subgroup, x=Estimate, xmin=LB, xmax=UB)) +
geom_point() +

#this changes the features of the overall effects
#one could resize, reshape, or recolor these points if desired
xlab("DID Estimates") + 
ylab(NULL) + 
#adds the CIs
geom_errorbarh(height=.1) +

#sets the scales
#note that I reverse the y axis to correctly order the effect #sizes based on my index variable
#scale_x_continuous(limits=c(-2.5,1), breaks = c(-2.5:1), name=xname)+
#scale_y_continuous(name = "", breaks=1:9, labels = t9.forest$X1, trans="reverse")+

#adding a vertical line at the effect = 0 mark
geom_vline(xintercept=0, color="black", linetype="dashed", alpha=.5)+

#faceting based on my subgroups
  #facet_wrap(~X1) +
facet_grid(Group~Measure, scales = "free", switch = 'y', space = "free_y")+

#thematic stuff
#ggtitle("Figure 3 - ATT Estimates by SubGroups")+
theme_tufte_revised()+
   theme(strip.placement = "outside") +
#theme(text=element_text(family="Times",size=12, color="black"))+
theme(panel.spacing = unit(1, "lines"),  strip.text =  element_text(face = "bold", size = 16))

poster.fig3
```

## Figure 5 - Financing

```{r}

df.comp <- df.qual.exp %>% #filter(`Grant Number` != "H80CS00225") %>% 
  ungroup() %>% 
  #inner_join(df.qual.exp %>% filter(Year == 2015 & Zips < 4.75) %>% select(`Grant Number`), by = c("Grant Number")) %>%  #& Zips < 3.75 or 4.75
  #filter(MarketShare > median(MarketShare, na.rm = TRUE)) %>% 
    select(Year, ZipFilter, ID, Att, `Total revenue (unrestricted & restricted)`, `Total net assets`, `Surplus as a pct of before dep expenses`, 
           `Staffing`, `Liabilities`, `Gross land, buildings and equipment (LBE)`, Liquidity, Profitability, Efficiency, NetWorth)

for (k in 1:3){

atts.comp <- rep(0, 7)
fig3.comp <- rep(0, 5)

if (k ==1){df.comp.root <- df.comp} else
  if (k == 2){df.comp.root <- df.comp %>% filter(ZipFilter == 2)} else
    if (k == 3){df.comp.root <- df.comp %>% filter(ZipFilter == 1)}

for (i in 5:ncol(df.comp)){

 # if (i %in% c(6, 7, 19)){temp <- df.comp.root[df.comp.root$Year %in% c(2015:2019), c(1:4, i)]} else {temp <- df.comp.root[,c(1:4, i)]}
  temp <- df.comp.root[,c(1:4, i)]
  #temp <- df.comp.root[df.comp.root$Year %in% c(2015:2019), c(1:4, i)]
  colnames(temp) <- c("Year", "ZipFilter", "ID","Att", "Outcome")
  temp$Outcome

  # temp <- temp %>%
  #   mutate(Outcome = ifelse(Outcome < quantile(temp$Outcome, probs = c(0.05, 0.95), na.rm = TRUE)[1], quantile(temp$Outcome, probs = c(0.05, 0.95), na.rm = TRUE)[1],
  #                    ifelse(Outcome > quantile(temp$Outcome, probs = c(0.05, 0.95), na.rm = TRUE)[2], quantile(temp$Outcome, probs = c(0.05, 0.95), na.rm = TRUE)[2], Outcome)))

cs.output.expand <- att_gt( yname = "Outcome",
                      tname = "Year",
                      idname = "ID",
                      gname = "Att",
                      xformla = ~1,
                      data = temp, # %>% filter(Year >= 2016), 
                      control_group = "nevertreated",
                      clustervars = "ID",
                      panel = FALSE, 
                      base_period = "universal"
                      )
cs.output.expand.agg.dyn.0 <- aggte(cs.output.expand, type = "simple", na.rm = TRUE)
cs.output.expand.agg.dyn <- aggte(cs.output.expand, type = "dynamic", na.rm = TRUE)

atts.comp <- rbind(atts.comp, cbind(colnames(df.comp)[i], round(mean(temp[temp$Att == 0 & temp$Year == min(temp$Year),]$Outcome, na.rm = TRUE), 3), round(mean(temp[temp$Att > 0 & temp$Year == min(temp$Year),]$Outcome, na.rm = TRUE) - mean(temp[temp$Att == 0 & temp$Year == min(temp$Year),]$Outcome, na.rm = TRUE), 3), round(cbind(cs.output.expand.agg.dyn.0$overall.att, 
      cs.output.expand.agg.dyn.0$overall.att - 1.96 * cs.output.expand.agg.dyn.0$overall.se, 
      cs.output.expand.agg.dyn.0$overall.att + 1.96 * cs.output.expand.agg.dyn.0$overall.se), 3)))

atts <- rbind(atts, cbind(colnames(df.comp)[i], round(cbind(cs.output.expand.agg.dyn.0$overall.att, 
      cs.output.expand.agg.dyn.0$overall.att - 1.96 * cs.output.expand.agg.dyn.0$overall.se, 
      cs.output.expand.agg.dyn.0$overall.att + 1.96 * cs.output.expand.agg.dyn.0$overall.se, 
      cs.output.expand.agg.dyn$overall.att, 
      cs.output.expand.agg.dyn$overall.att - cs.output.expand.agg.dyn$crit.val.egt * cs.output.expand.agg.dyn$overall.se, 
      cs.output.expand.agg.dyn$overall.att + cs.output.expand.agg.dyn$crit.val.egt * cs.output.expand.agg.dyn$overall.se), 3)))
z <- rbind(z, c(colnames(df.comp)[i], cs.output.expand$Wpval))

fig3.comp <- rbind(fig3.comp, 
                 data.frame(Outcome = colnames(df.comp)[i], 
                      Year = cs.output.expand.agg.dyn$egt, 
           Estimate = cs.output.expand.agg.dyn$att.egt, 
           CILB = cs.output.expand.agg.dyn$att.egt - cs.output.expand.agg.dyn$crit.val.egt * cs.output.expand.agg.dyn$se.egt,
           CIUB = cs.output.expand.agg.dyn$att.egt + cs.output.expand.agg.dyn$crit.val.egt * cs.output.expand.agg.dyn$se.egt))
}

rownames(fig3.comp) <- NULL
if (k == 1){atts.comp.full <- atts.comp[-1,]} else 
 if (k == 2){atts.comp.larger <- atts.comp[-1,]} else 
   {atts.comp.smaller <- atts.comp[-1,]}

}

#Event Study Plots
fig3.comp[-1,] %>%
  ggplot(mapping = aes(x = Year, y = Estimate)) +
  facet_wrap(~Outcome, scales = "free_y") +
  ylab("Percentage Points") +
  geom_vline(xintercept=-0.5, linetype = "dashed", color = "red") +
  geom_hline(yintercept=0, linetype = "dashed", color = "grey") +
  geom_point() + geom_errorbar(aes(ymin=CILB, ymax=CIUB), width=.2) +
  ggtitle("fig3 - Event Study") +
  theme_tufte_revised()

 cbind(atts.comp.full, atts.comp.larger, atts.comp.smaller) %>% 
   kable("html", booktabs = T, col.names = rep(c("Outcome", "C_0", "T_0 - C_0", "DID Estimate", "LB", "UB"), 3)) %>%
  kable_styling(latex_options = c("striped")) %>%
  #pack_rows(index = c("Financial" = 5, "" = 5, "Profitability" = 5)) %>% 
  add_header_above(c(" " = 1, "Full" = 6, "More Market Share Clinics" = 6, "Less Market Share Clinics" = 6)) %>% 
  #pack_rows(index = c("Larger FQHCs" = 15, "Smaller FQHCs" = 15)) %>% 
  add_footnote(c(""))

```

## Figure 5 - Financing these changes

```{r}

df.fin <- df.qual.exp %>% #filter(`Grant Number` != "H80CS00225") %>% 
  ungroup() %>% 
  #filter(MarketShare > median(MarketShare, na.rm = TRUE)) %>% 
  select(`Grant Number`, Year, ID, Att, `Total revenue (unrestricted & restricted)`, `Total net assets`, `Surplus as a pct of before dep expenses`, 
           `Staffing`, `Liabilities`, `Gross land, buildings and equipment (LBE)`, Liquidity, Profitability, Efficiency, NetWorth) %>%
  inner_join(df.qual.exp %>% filter(Year == 2016) %>% mutate(ZipFilter = ifelse(`Surplus as a pct of before dep expenses` < 2, 1, 2)) %>% select(`Grant Number`, ZipFilter), by = c("Grant Number")) %>% 
    select(Year, ZipFilter, ID, Att, `Total revenue (unrestricted & restricted)`, `Total net assets`, `Surplus as a pct of before dep expenses`, 
           `Staffing`, `Liabilities`, `Gross land, buildings and equipment (LBE)`, Liquidity, Profitability, Efficiency, NetWorth)

for (k in 1:3){

atts.fin <- rep(0, 7)
fig3.fin <- rep(0, 5)

if (k ==1){df.fin.root <- df.fin} else
  if (k == 2){df.fin.root <- df.fin %>% filter(ZipFilter == 2)} else
    if (k == 3){df.fin.root <- df.fin %>% filter(ZipFilter == 1)}

for (i in 5:ncol(df.fin)){

  temp <- df.fin.root[,c(1:4, i)]
  #temp <- df.fin.root[df.fin.root$Year %in% c(2015:2019), c(1:4, i)]
  colnames(temp) <- c("Year", "ZipFilter", "ID","Att", "Outcome")
  temp$Outcome

  # temp <- temp %>%
  #   mutate(Outcome = ifelse(Outcome < quantile(temp$Outcome, probs = c(0.05, 0.95), na.rm = TRUE)[1], quantile(temp$Outcome, probs = c(0.05, 0.95), na.rm = TRUE)[1],
  #                    ifelse(Outcome > quantile(temp$Outcome, probs = c(0.05, 0.95), na.rm = TRUE)[2], quantile(temp$Outcome, probs = c(0.05, 0.95), na.rm = TRUE)[2], Outcome)))

cs.output.expand <- att_gt( yname = "Outcome",
                      tname = "Year",
                      idname = "ID",
                      gname = "Att",
                      xformla = ~1,
                      data = temp, # %>% filter(Year >= 2016), 
                      control_group = "nevertreated",
                      clustervars = "ID",
                      panel = FALSE, 
                      base_period = "universal"
                      )
cs.output.expand.agg.dyn.0 <- aggte(cs.output.expand, type = "simple", na.rm = TRUE)
cs.output.expand.agg.dyn <- aggte(cs.output.expand, type = "dynamic", na.rm = TRUE)

atts.fin <- rbind(atts.fin, cbind(colnames(df.fin)[i], round(mean(temp[temp$Att == 0 & temp$Year == min(temp$Year),]$Outcome, na.rm = TRUE), 3), round(mean(temp[temp$Att > 0 & temp$Year == min(temp$Year),]$Outcome, na.rm = TRUE) - mean(temp[temp$Att == 0 & temp$Year == min(temp$Year),]$Outcome, na.rm = TRUE), 3), round(cbind(cs.output.expand.agg.dyn.0$overall.att, 
      cs.output.expand.agg.dyn.0$overall.att - 1.96 * cs.output.expand.agg.dyn.0$overall.se, 
      cs.output.expand.agg.dyn.0$overall.att + 1.96 * cs.output.expand.agg.dyn.0$overall.se), 3)))

atts <- rbind(atts, cbind(colnames(df.fin)[i], round(cbind(cs.output.expand.agg.dyn.0$overall.att, 
      cs.output.expand.agg.dyn.0$overall.att - 1.96 * cs.output.expand.agg.dyn.0$overall.se, 
      cs.output.expand.agg.dyn.0$overall.att + 1.96 * cs.output.expand.agg.dyn.0$overall.se, 
      cs.output.expand.agg.dyn$overall.att, 
      cs.output.expand.agg.dyn$overall.att - cs.output.expand.agg.dyn$crit.val.egt * cs.output.expand.agg.dyn$overall.se, 
      cs.output.expand.agg.dyn$overall.att + cs.output.expand.agg.dyn$crit.val.egt * cs.output.expand.agg.dyn$overall.se), 3)))
z <- rbind(z, c(colnames(df.fin)[i], cs.output.expand$Wpval))

fig3.fin <- rbind(fig3.fin, 
                 data.frame(Outcome = colnames(df.fin)[i], 
                      Year = cs.output.expand.agg.dyn$egt, 
           Estimate = cs.output.expand.agg.dyn$att.egt, 
           CILB = cs.output.expand.agg.dyn$att.egt - cs.output.expand.agg.dyn$crit.val.egt * cs.output.expand.agg.dyn$se.egt,
           CIUB = cs.output.expand.agg.dyn$att.egt + cs.output.expand.agg.dyn$crit.val.egt * cs.output.expand.agg.dyn$se.egt))
}

rownames(fig3.fin) <- NULL
if (k == 1){atts.fin.full <- atts.fin[-1,]} else 
 if (k == 2){atts.fin.larger <- atts.fin[-1,]} else 
   {atts.fin.smaller <- atts.fin[-1,]}

}

#Event Study Plots
fig3.fin[-1,] %>%
  ggplot(mapping = aes(x = Year, y = Estimate)) +
  facet_wrap(~Outcome, scales = "free_y") +
  ylab("Percentage Points") +
  geom_vline(xintercept=-0.5, linetype = "dashed", color = "red") +
  geom_hline(yintercept=0, linetype = "dashed", color = "grey") +
  geom_point() + geom_errorbar(aes(ymin=CILB, ymax=CIUB), width=.2) +
  ggtitle("fig3 - Event Study") +
  theme_tufte_revised()

 cbind(atts.fin.full, atts.fin.larger, atts.fin.smaller) %>% 
   kable("html", booktabs = T, col.names = rep(c("Outcome", "C_0", "T_0 - C_0", "DID Estimate", "LB", "UB"), 3)) %>%
  kable_styling(latex_options = c("striped")) %>%
  pack_rows(index = c("General Measures" = 4, "Altmans Z Components" = 5)) %>% 
  add_header_above(c(" " = 1, "Full" = 6, "Profitable" = 6, "Unprofitable" = 6)) %>% 
  #pack_rows(index = c("Larger FQHCs" = 15, "Smaller FQHCs" = 15)) %>% 
  add_footnote(c(""))
```

## True Figure 4
```{r}
fields <- 10
fig4 <- cbind(c(rep("Full", fields), rep("More Market Share", fields), rep("Less Market Share", fields), rep('Profitable', fields), rep('Unprofitable/Budget Neutral', fields)), 
      rbind(atts.comp.full, atts.comp.larger, atts.comp.smaller, atts.fin.larger, atts.fin.smaller)) %>% 
  data.frame()

colnames(fig4) <- c("Subgroup", "Measure", "C_0", "T_0 - C_0", "Estimate", "LB", "UB")

fig4$Measure <- ifelse(fig4$Measure == 'Gross land, buildings and equipment (LBE)', 'Property and Buildings', fig4$Measure)

fig4$C_0 <- as.numeric(fig4$C_0)
fig4$`T_0 - C_0` <- as.numeric(fig4$`T_0 - C_0`)
fig4$Estimate <- as.numeric(fig4$Estimate)
fig4$LB <- as.numeric(fig4$LB)
fig4$UB <- as.numeric(fig4$UB)

fig4$Groups <- ifelse(fig4$Measure %in% c('Total revenue (unrestricted & restricted)','Property and Buildings','Staffing', 'Liabilities'), 'Mechanisms', 
               ifelse(fig4$Measure %in% c('Liquidity','Efficiency'), 'Financing', NA))

rownames(fig4) <- NULL

fig4 %>% 
  filter(Groups == 'Mechanisms' ) %>% 
ggplot(aes(y=Measure, x=Estimate, xmin=LB, xmax=UB)) +
geom_point() +

#this changes the features of the overall effects
#one could resize, reshape, or recolor these points if desired
xlab("DID Estimates") + 
ylab(NULL) + 
#adds the CIs
geom_errorbarh(height=.1) +

#sets the scales
#note that I reverse the y axis to correctly order the effect #sizes based on my index variable
#scale_x_continuous(limits=c(-2.5,1), breaks = c(-2.5:1), name=xname)+
#scale_y_continuous(name = "", breaks=1:9, labels = t9.forest$X1, trans="reverse")+

#adding a vertical line at the effect = 0 mark
geom_vline(xintercept=0, color="black", linetype="dashed", alpha=.5) +

#faceting based on my subgroups
  #facet_wrap(~X1) +
facet_grid(~Subgroup, scales = "free_y", switch = 'y', space = "free_y")+

#thematic stuff
#ggtitle("Figure 3 - ATT Estimates by SubGroups")+
theme_tufte_revised()+
   theme(strip.placement = "outside") +
#theme(text=element_text(family="Times",size=12, color="black"))+
theme(panel.spacing = unit(1, "lines")) +
  ggtitle("Figure 2, Incumbent Only (FQHC-zip-year units)")

# fig4 %>% 
#   select(Measure:UB) %>% 
#   filter(Measure %notin% c('Total revenue (unrestricted & restricted)', 'Total expenses before depreciation', 'Surplus as a pct of before dep expenses')) %>% 
#   kable("html", booktabs = T, col.names = c("Outcome", "C_0", "T_0 - C_0", "DID Estimate", "LB", "UB")) %>%
#   kable_styling(latex_options = c("striped")) %>%
#   pack_rows(index = c("General Measures" = 4, "Altmans Z Components" = 5)) %>% 
#   add_header_above(c(" " = 1, "Full" = 6, "Profitable" = 6, "Unprofitable" = 6)) %>% 
#   #pack_rows(index = c("Larger FQHCs" = 15, "Smaller FQHCs" = 15)) %>% 
#   add_footnote(c(""))

library(formattable)
fig4$C_0 <- currency(fig4$C_0, digits = 0L)
fig4$`T_0 - C_0` <- currency(fig4$`T_0 - C_0`, digits = 0L)
fig4$Estimate <- currency(fig4$Estimate, digits = 0L)
fig4$LB <- currency(fig4$LB, digits = 0L)
fig4$UB <- currency(fig4$UB, digits = 0L)

fig4 %>% 
  filter(Groups == 'Mechanisms' ) %>% 
  select(Measure:UB) %>% 
  kable("html", booktabs = T, col.names = c("Outcome", "C_0", "T_0 - C_0", "DID Estimate", "LB", "UB")) %>%
  kable_styling(latex_options = c("striped")) %>%
  row_spec(c(6, 14),bold=T, italic = T, hline_after = T) %>% 
  row_spec(c(19),bold=F, italic = T, hline_after = T) %>% 
  pack_rows(index = c("Panel A: Full Sample" = 4, "Panel B: Higher Low-Income Market Share Subgroup" = 4, "Panel B: Lower Low-Income Market Share Subgroup" = 4, 
                      "Panel C: Profitable Subgroup" = 4, "Panel C: Unprofitable/Budget Neutral Subgroup" = 4)) %>% 
  #add_header_above(c(" " = 1, "Full" = 6, "Profitable" = 6, "Unprofitable" = 6)) %>% 
  #pack_rows(index = c("Larger FQHCs" = 15, "Smaller FQHCs" = 15)) %>% 
  add_footnote(c(""))

```

# Robustness Testing

## Service Area Overlap measurement - Plausible Exogeneity

- >99% of zips in sample have unmet Low income, Medicaid, or Uninsured need at baseline
- Distance of Rival's new zip code to nearest existing zip code in service area,  75% are below 10km. All those > 10km to nearest site are in rural areas traveling across very low density land
```{r}

## Need
df.z.full %>% 
  filter(Year == 2015) %>% 
  left_join(census, by = c("Zip" = "ZipCode", "Year" = "YEAR")) %>% 
  mutate(Pct.Total = Total.z / TotalPop, Pct.LowInc = Total.z / (IncomeBelowPoverty / PovertyStatusDet * TotalPop), Pct.MCD = MCD.z / (Medicaid / MCD.denom * TotalPop), Pct.None = None.z / (Uninsured / Unins.denom * TotalPop) ) %>% 
  select(Zip:Att.z, Pct.Total:Pct.None) %>% 
  mutate(NeedMore = ifelse(Pct.LowInc < 1 | Pct.MCD < 1 | Pct.None < 1, 1, 0), 
         Tx = ifelse(Att.z > 0, 1, 0)) %>% 
  #group_by(Tx) %>% 
  mutate(#n(), 
            NeedMore = NeedMore, 
            NeedLowIncome = ifelse(Pct.LowInc < 1, 1, 0), 
            NeedMCD = ifelse(Pct.MCD < 1, 1, 0),
            NeedNone = ifelse(Pct.None < 1, 1, 0)) %>% 
  select(Tx, NeedMore:NeedNone) %>% 
  filter(!is.na(NeedMore)) %>% 
  tbl_summary(by = Tx) %>% 
  add_p()

## Distance of Competitor to New Site
df.z.dist <- df.z.full %>% 
  filter(Att.z == Year) %>% 
  select(Zip:Total.z) %>% 
  left_join(uds.sites %>% select(`Grant Number`, Year, Zip) %>% unique(), by = c("Zip","Year")) %>% 
  left_join(df.z %>% filter(Tx == 1) %>% select(`Grant Number`, Year, Zip) %>% unique(), by = c("Zip","Year")) %>% 
  filter(`Grant Number.x` != `Grant Number.y`)

## Read in Centroid Data
zip.centroids <- read.csv("ZIP_Code_Population_Weighted_Centroids.csv") %>% 
  mutate(ZipCode = str_pad(STD_ZIP5, 5, pad = "0")) %>% 
  select(ZipCode, X, Y)
## Add missing centroids
zip.centroids <- rbind(zip.centroids, c(ZipCode = "37243", X = -86.783, Y = 36.165), 
                                      c(ZipCode = "95951", X = -121.98, Y = 39.73)) %>% 
  mutate(X = as.numeric(X), Y = as.numeric(Y))

df.z.dist <- df.z.dist %>% mutate(DistToClosestSite = NA, ZipOfClosestSite = NA)

for (g in 1:nrow(df.z.dist)){
#g <- 3
 temp <- rbind(zip.centroids %>% filter(ZipCode == df.z.dist[g,]$Zip), 
          zip.centroids %>% filter(ZipCode %in% unique(uds.sites[uds.sites$`Grant Number` == df.z.dist[g,]$`Grant Number.x` & uds.sites$Year == df.z.dist[g,]$Year,]$Zip))) %>% 
   unique()
  mm <- as.matrix(dist(as.matrix(temp$X, temp$Y), method = "euclidean"))[-1,1]
  df.z.dist[g,]$DistToClosestSite <-  min(mm) * 111.139
  df.z.dist[g,]$ZipOfClosestSite  <- ifelse(is.na(temp[2,]$ZipCode), temp[1,]$ZipCode, unlist(temp[-1,]$ZipCode[rank(mm) == 1]))
}

df.z.dist[19,]$DistToClosestSite <- 0

mean(df.z.dist$DistToClosestSite)
quantile(df.z.dist$DistToClosestSite, probs = c(0, 0.1, 0.25, 0.5, 0.75, 0.9, 1))

mean(df.z.dist[df.z.dist$ZipFilter == 1,]$DistToClosestSite)
mean(df.z.dist[df.z.dist$ZipFilter == 2,]$DistToClosestSite)

quantile(df.z.dist[df.z.dist$ZipFilter == 1,]$DistToClosestSite, probs = c(0, 0.1, 0.25, 0.5, 0.75, 0.9, 1))
quantile(df.z.dist[df.z.dist$ZipFilter == 2,]$DistToClosestSite, probs = c(0, 0.1, 0.25, 0.5, 0.75, 0.9, 1))

zipsize <- zctas(starts_with = c(df.z.dist$Zip, df.z.dist$ZipOfClosestSite), year = 2010)

## Original Figure
df.z.dist.plot <- df.z.dist %>% 
  left_join(zipsize %>% select(GEOID10, ALAND10), by = c("Zip" = "GEOID10")) %>% 
  left_join(uds.sites %>% select(Zip, `Site State`) %>% unique(), by = c("Zip")) %>% 
  left_join(census %>% select(ZipCode, YEAR, TotalPop), by = c("Year" = "YEAR", "Zip" = "ZipCode")) %>% 
  arrange(desc(DistToClosestSite)) %>% 
  mutate(Area.SQmiles = ALAND10 / 2.59e6) %>% 
  mutate(Density.NewZip = TotalPop / Area.SQmiles, 
         l.density = log(TotalPop / Area.SQmiles))

df.z.dist.plot %>% 
  ggplot(aes(x = log(Density.NewZip), y = DistToClosestSite)) + geom_point(aes(size = Area.SQmiles)) +
  ggtitle("Distance from Rival's newly competing zip ocde to closest incumbent zip code by Density and Area of newly competing zip code")

summary(lm(DistToClosestSite ~ log(Density.NewZip), data = df.z.dist.plot))

# Updated Figure
df.z.dist.plot <- df.z.dist.plot %>% 
  select(Zip:ZipOfClosestSite, Density.NewZip) %>% 
  left_join(zipsize %>% select(GEOID10, ALAND10), by = c("ZipOfClosestSite" = "GEOID10")) %>% 
  left_join(uds.sites %>% select(Zip, `Site State`) %>% unique(), by = c("ZipOfClosestSite" = "Zip")) %>% 
  left_join(census %>% select(ZipCode, YEAR, TotalPop), by = c("Year" = "YEAR", "ZipOfClosestSite" = "ZipCode")) %>% 
  arrange(desc(DistToClosestSite)) %>% 
  mutate(Area.SQmiles = ALAND10 / 2.59e6) %>% 
  mutate(Density.Origin = TotalPop / Area.SQmiles)

df.z.dist.plot %>% 
  ggplot(aes(x = log(Density.NewZip), y = DistToClosestSite, color = log(Density.Origin))) + geom_point() +
  scale_colour_gradientn(colours = terrain.colors(10))+ theme_minimal()+
  ylab("Distance from Competative to closest Rival Site (km)") +
  ggtitle("Distance from newly competative Zip to Rival's closest existing zip, by log Density of areas")

p <- df.z.dist.plot %>% 
  ggplot(aes(x = log(Density.Origin), y = log(Density.NewZip))) + 
  geom_point(aes(size = DistToClosestSite, colour = DistToClosestSite)) + 
  scale_colour_gradientn(colours = c("green","lightblue","purple"),
                         values = c(1.0,0.7,0.1,0))  +#scale_colour_gradientn(colours = terrain.colors(10))+ 
  theme_minimal() + ylab("Log Population Density of New Zip Code") + 
  xlab("Log Population Density of Nearest Service Area Zip Code") + 
  labs(color = "Distance Travelled to\nOpenNew Site (km)") + 
  guides(size = FALSE) + xlim(0.5, 11.5) + ylim(0.5, 11.5)
p

summary(lm(log(Density.Origin) ~ log(Density.NewZip), data = df.z.dist.plot))

```

## Estimate effects of competition on nearest FQHC, takes 20 minutes to run code
```{r}

## Build the first part of the dataset

df.spill <- df.z %>% 
  left_join(df.z.dist %>% 
  select(`Grant Number.y`, Zip) %>% 
    unique() %>% 
    rename(ZipOfComp = Zip), by = c("Grant Number" = "Grant Number.y")) %>% 
  #filter(!is.na(ZipOfComp)) %>% 
  mutate(ZipOfComp = ifelse(`Grant Number` == "H80CS00765", 95991, ZipOfComp)) %>% 
  left_join(zip.centroids, by = c("Zip" = "ZipCode")) %>% 
  left_join(zip.centroids, by = c("ZipOfComp" = "ZipCode")) %>% 
  mutate(Dist = sqrt((X.x - X.y) ^ 2 + (Y.x - Y.y) ^ 2) * 111.139) %>% 
  mutate(Dist.f = ifelse(is.na(Dist), "-1",
                  ifelse(Dist == 0, 0, 
                  ifelse(Dist > 0 & Dist < 20, 1, 
                  ifelse(Dist >= 20 & Dist < 50, 2, 3))))) %>% 
  #filter(is.na(Dist.f) | Dist.f == 2) %>% 
  left_join(uds.sites %>% 
              filter(`Location Setting` %in% c("All Other Clinic Types")) %>% 
              mutate(Type.num = ifelse(`Location Type` == "Seasonal", 0.33, 
                                ifelse(`Location Type` == "Mobile Van", 0.66, 1)), 
                     Schedule = ifelse(`Operational Schedule` == "Full-Time", 1, 
                                ifelse(`Operational Schedule` == "Part-Time", 0.5, 0)),
                     Calendar = ifelse(`Calendar Schedule` == "Year-Round", 1,
                                ifelse(`Calendar Schedule` == "Seasonal", 0.5, 0))) %>% 
              group_by(`Grant Number`, Year, Zip) %>% 
              summarise(Sites = n(), Mean.Hours = mean(Access, na.rm = TRUE), Total.Hours = sum(Access, na.rm = TRUE), 
                        Perminance = sum(Type.num, na.rm = TRUE), Schedule = sum(Schedule, na.rm = TRUE), Calendar = sum(Calendar, na.rm = TRUE)), by = c("Grant Number","Year", "Zip")) %>% 
  mutate(Sites = ifelse(is.na(Sites), 0, Sites), 
         Mean.Hours = ifelse(is.na(Mean.Hours), 0, Mean.Hours), 
         Total.Hours = ifelse(is.na(Total.Hours), 0, Total.Hours), 
         Perminance = ifelse(is.na(Perminance), 0, Perminance), 
         Schedule = ifelse(is.na(Schedule), 0, Schedule), 
         Calendar = ifelse(is.na(Calendar), 0, Calendar))

dist.vecs <- df.spill[!is.na(df.spill$Dist),]$Dist

df.spill %>% 
  group_by(Dist.f) %>% 
  summarise(n())

## Find neighbors based on some circle around FQHC zip

mm <- dist(zip.centroids[,2:3])
mm <- as.matrix(mm) * 111.139

mm.2 <- mm[zip.centroids$ZipCode %in% unique(df.spill$Zip),]

mm.2 <- as.matrix((mm.2 < 50) + 0 )
rowSums(mm.2)

mm.3 <- c(CenterZip = 0, Year = 0, Competing.Sites = 0)
for (g in 1:dim(mm.2)[1]){
  mm.3 <- rbind(mm.3, cbind(CenterZip = zip.centroids[rownames(mm.2)[g],]$ZipCode, zip.centroids[names(mm.2[g,][which(as.vector(mm.2[g,] == 1))]),]) %>% 
  left_join(uds.sites %>% filter(Year %in% c(2015:2019)) %>% group_by(Year, Zip) %>% summarise(Sites.Total = n()), by = c("ZipCode" = "Zip")) %>% 
  left_join(df.spill %>% select(`Grant Number`, Year, Zip, Sites), by = c("ZipCode" = "Zip", "Year")) %>% 
  mutate(Sites.Total = ifelse(is.na(Sites.Total), 0, Sites.Total), 
         Sites = ifelse(is.na(Sites), 0, Sites), 
         Competing.Sites = Sites.Total - Sites) %>% 
  filter(CenterZip != ZipCode) %>% 
  group_by(CenterZip, Year) %>% 
  summarise(Competing.Sites = sum(Competing.Sites)) %>% 
    as.matrix())
}

mm.3 <- mm.3[-1,]

#mm.100 <- mm.3 %>% data.frame() 
#mm.60 <- mm.3 %>% data.frame() 
#mm.50 <- mm.3 %>% data.frame() 
#mm.20 <- mm.3 %>% data.frame() 
#mm.10 <- mm.3 %>% data.frame()

#mm.3 <- mm.50

```

```{r}
cs.output.expand <- att_gt( yname = "Total.Hours",
                      tname = "Year",
                      idname = "ID",
                      gname = "Att",
                      xformla = ~1,
                      data = df.spill %>% filter(Dist.f %in% c("-1",3)), # %>% filter(Year > 2015),
                      control_group = "nevertreated",
                      clustervars = "ID",
                      panel = TRUE,
                      base_period = "universal"
                      )
aggte(cs.output.expand, type = "simple", na.rm = TRUE)
```

## Control for spatial Spillovers
```{r}

df.spill2 <- df.spill %>% 
  left_join(mm.3 %>% data.frame() %>% mutate(Year = as.numeric(Year)), by = c("Zip" = "CenterZip","Year")) %>% 
  mutate(Competing.Sites = as.numeric(ifelse(is.na(Competing.Sites), 0, Competing.Sites)), 
         D = ifelse(Competing.Sites > 0, 1, 0))

#length(unique(df.spill2[is.na(df.spill2$Dist),]$Zip))  1893 or 1893 * 5

set.seed(234)  #232
df.spill2 <- df.spill2 %>% mutate(Dist = ifelse(is.na(Dist), rep(sample(dist.vecs, 1893, replace = T), each = 5), Dist), 
                                  Att.viz = ifelse(Att == 0, rep(sample(c(2017:2019), 1893, replace = T), each = 5), Att)) %>% 
  mutate(Dist.rounded = as.factor(ceiling(Dist / 10)))

df.spill2 <- df.spill2 %>% 
  ungroup() %>% 
  #filter(MarketShare < median(MarketShare, na.rm = TRUE)) %>% 
  select(`Grant Number`, Tx, D, Att, Att.viz, Dist, Dist.rounded, Year, Zip, Total.Hours, Sites, Perminance, Schedule, Calendar, Total.z, MCD.z, MCR.z, COM.z, None.z) %>%
  inner_join(df.qual.exp %>% filter(Year == 2016) %>% mutate(ZipFilter = ifelse(`Surplus as a pct of before dep expenses` < 2, 1, 2)) %>% select(`Grant Number`, ZipFilter), by = c("Grant Number")) %>% 
    select(Tx, D, Dist, Year, Zip, `Grant Number`, ZipFilter, Att, Att.viz, Dist.rounded, Total.Hours, Sites, Perminance, Schedule, Calendar, Total.z, MCD.z, MCR.z, COM.z, None.z) # %>% filter(ZipFilter == 2)

  i <- 11
  df.temp <- df.spill2[,c(1:5, i)]
  names(df.temp) <- c("Tx", "D", "Dist", "Year", "Zip", 'Outcome')
  model_feols_clustered <- feols(Outcome ~ Tx*as.numeric(Dist / 10) + D | Year + Zip, cluster = ~ Zip, data = df.temp)
  TEs <- data.frame(model_feols_clustered$coeftable)[1,]
  TEs_Dist <- data.frame(model_feols_clustered$coeftable)[nrow(model_feols_clustered$coeftable),]

for (i in 12:20){
  df.temp <- df.spill2[,c(1:5, i)]
  names(df.temp) <- c("Tx", "D", "Dist", "Year", "Zip", 'Outcome')
  model_feols_clustered <- feols(Outcome ~ Tx*as.numeric(Dist / 10) + D | Year + Zip, cluster = ~ Zip, data = df.temp)
  TEs <- rbind(TEs, data.frame(model_feols_clustered$coeftable)[1,])
  TEs_Dist <- rbind(TEs_Dist, data.frame(model_feols_clustered$coeftable)[nrow(model_feols_clustered$coeftable),])
}

p.spill <- df.spill2 %>% 
  mutate(Treats = ifelse(Att == 0 & Year >=  Att.viz, 1, Tx)) %>% 
  filter(Att > 0) %>% 
  group_by(Dist.rounded, Treats) %>% 
  summarise(Counts = n(), Sites = mean(Sites, na.rm = TRUE), Total.z = mean(Total.z, na.rm = TRUE), Total.Hours = mean(Total.Hours, na.rm = TRUE), MCD.z = mean(MCD.z, na.rm = TRUE)) %>% 
  pivot_wider(id_cols = Dist.rounded, names_from = Treats, values_from = c(Counts, Sites, Total.z, Total.Hours, MCD.z)) %>% 
  mutate(Sites = Sites_1 - Sites_0, 
         Total.z = Total.z_1 - Total.z_0, 
         Total.Hours = Total.Hours_1 - Total.Hours_0, 
         MCD.z = MCD.z_1 - MCD.z_0) %>% 
  left_join(df.spill2 %>% 
  mutate(Treats = ifelse(Att == 0 & Year >= Att.viz, 1, Tx)) %>% 
  filter(Att == 0) %>% 
  group_by(Dist.rounded, Treats) %>% 
  summarise(Counts = n(), Sites = mean(Sites, na.rm = TRUE), Total.z = mean(Total.z, na.rm = TRUE), Total.Hours = mean(Total.Hours, na.rm = TRUE), MCD.z = mean(MCD.z, na.rm = TRUE)) %>% 
  pivot_wider(id_cols = Dist.rounded, names_from = Treats, values_from = c(Counts, Sites, Total.z, Total.Hours, MCD.z)) %>% 
  mutate(Sites.control = Sites_1 - Sites_0, 
         Total.z.control = Total.z_1 - Total.z_0, 
         Total.Hours.control = Total.Hours_1 - Total.Hours_0, 
         MCD.z.control = MCD.z_1 - MCD.z_0) %>% 
    select(Dist.rounded, Sites.control:MCD.z.control), by = c("Dist.rounded")) %>% 
  mutate( Locations = Sites - Sites.control, 
         `Patients` = Total.z - Total.z.control, 
         `Patients - Medicaid Only` = MCD.z - MCD.z.control, 
         `Hours per Week` = Total.Hours - Total.Hours.control) %>% 
  select(Dist.rounded:Counts_1, `Locations`:`Hours per Week`) %>% 
  mutate(Dist.rounded = as.numeric(paste(Dist.rounded, 0, sep = ""))) %>% 
  pivot_longer(cols = c( `Locations`:`Hours per Week`)) %>% 
  mutate(`Treated Zips` = Counts_0 / 3) %>% 
  ggplot(aes(x = Dist.rounded, y = value)) + geom_point(aes(size = `Treated Zips`)) + facet_wrap(~name, scales = "free_y", ncol = 2) + xlim(0, 200) + 
  xlab("Kilometers to Competitive Zip") + ylab("Unadjusted Difference-in-differences") + theme_tufte_revised() + 
  geom_smooth(method = "lm", se = F, quiet = TRUE, formula = "y ~ x", mapping = aes(weight = `Treated Zips`)) + 
  geom_hline(yintercept=0, linetype = "dashed", color = "grey") + theme(text = element_text(size = 16)) + theme(title = element_text(face = "bold")) + 
  ggtitle("Figure 4 - Unadjusted DID Estimates by Distance to Competitive Zip")

suppressWarnings(print(p.spill))

output <- cbind(TEs %>% 
  mutate(`95% LB` = Estimate - `Std..Error` * 1.96, 
         `95% UB` = Estimate + `Std..Error` * 1.96) %>% 
    select(Estimate,`95% LB`, `95% UB`) ,

TEs_Dist %>% 
  mutate(`95% LB` = Estimate - `Std..Error` * 1.96, 
         `95% UB` = Estimate + `Std..Error` * 1.96) %>% 
    select(Estimate,`95% LB`, `95% UB`)) 
rownames(output) <- c("Hours Per Week","Locations","Perminance","Weekly Schedule","Calendar Schedule","Total","Medicaid","Medicare","Commerical","Uninsured")

round(output[c(6:10, 2,3, 1, 4, 5),], digits = 3) %>% 
  kable("html", booktabs = T, col.names = c("DID Estimate", "95% LB", "95% UB", "DID Estimate", "95% LB", "95% UB")) %>%
  kable_styling(latex_options = c("striped")) %>%
  pack_rows(index = c("Patient Panel" = 5, "Extensive Margin" = 2, "Intensive Margin" = 3)) %>% 
  add_header_above(c(" " = 1, "Treatment Effect" = 3, "Treatment Effect * Distance" = 3)) %>% 
  add_footnote(c("Treatment Effect is the ATT estimated in the zip code where the rival opens their new location","Treatment Effect * Distance is the change in the ATT for every 10km an outlying zip code in the incumbent's service area is from the competitive zip "))
```

## Preperiod parallel trends
```{r}

i <- 4
df.temp2 <- df.spill2 %>% filter(Att == 0 | Year < Att) %>% mutate(Treat = ifelse(Att > 0, 1, 0)) %>% 
                   select(Year, Treat, Zip, Total.Hours:None.z)
df.temp2 <- df.temp2[,c(1:3, i)]
names(df.temp2) <- c('Year','Treat','Zip','Outcome')

parallel.trends.set <- data.frame(feols(Outcome ~ Treat * Year, 
                 data = df.temp2)$coeftable)[4,]

for (i in 5:13){
  df.temp2 <- df.spill2 %>% filter(Att == 0 | Year < Att) %>% mutate(Treat = ifelse(Att > 0, 1, 0)) %>% 
                   select(Year, Treat, Zip, Total.Hours:None.z)
df.temp2 <- df.temp2[,c(1:3, i)]
names(df.temp2) <- c('Year','Treat','Zip','Outcome')

parallel.trends.set <- rbind(parallel.trends.set, data.frame(feols(Outcome ~ Treat * Year , 
                 data = df.temp2)$coeftable)[4,])
}
parallel.trends.set
```

## Again, with did2s
```{r}
did2s(
      data = df.spill2, 
      yname = "MCD.z", 
      treatment = "Tx",
        first_stage = ~ 0 | Zip + Year + D, 
        second_stage = ~ i(Tx, ref = FALSE) + as.numeric(Dist / 10) + I(Tx *as.numeric(Dist / 10)),
        cluster_var = "Zip",
      verbose = FALSE
    )

```

## How do close versus far zip codes compare

```{r}

df.temp3 <- df.spill2 %>% 
  left_join(census, by = c("Zip" = "ZipCode", "Year" = "YEAR" )) %>% 
    filter(Year %in% c(2015:2016) & Att > 0) %>% 
  mutate(Pct.Age = Age65andOver / TotalPop * 100, 
         Pct.Pov = IncomeBelowPoverty / PovertyStatusDet * 100, 
         Pct.HL = Hispanic / TotalPop * 100, 
         Pct.Black = Black / TotalPop * 100,
         Pct.White = White / TotalPop * 100,
         Pct.Unemp = Unemployed / Unemp_denom * 100,
         Pct.MCD = Medicaid / MCD.denom * 100, 
         Pct.NoIns = Uninsured / Unins.denom * 100) %>% 
  select(Dist, Pct.Age, Pct.HL, Pct.Black, Pct.White, Pct.Pov, Pct.Unemp, Pct.MCD, Pct.NoIns)

ziptypes <- data.frame(0, 0, 0, 0)
for (k in 2:ncol(df.temp3)){
  df.temp4 <- df.temp3[,c(1, k)]
  names(df.temp4) <- c('Dist','Outcome')

  ziptypes <- rbind(ziptypes, summary(lm(data = df.temp4, Outcome ~ as.numeric(Dist / 10) ))$coeff[2,])
}
ziptypes <- ziptypes[-1,]

rownames(ziptypes) <- c("% age over 65+","% Hispanic/Latino","% Black","% White", "% Poverty","% Unemployed","% Medicaid","% Uninsured")

ziptypes %>% 
  mutate(Estimate = X0, 
         `95% LB` = X0 - X0.1 * 1.96, 
         `95% UB` = X0 + X0.1 * 1.96, 
         `p-value` = X0.3) %>% 
  select(Estimate:`p-value`) %>% 
  kable("html", booktabs = T) %>%
  kable_styling(latex_options = c("striped")) %>%
  pack_rows(index = c("Demographics" = 4, "Income" = 2, "Insurance" = 2)) %>% 
  row_spec(c(1, 4), bold = T, hline_after = T) %>% 
  row_spec(c(3, 5:8), italic = T, hline_after = T) %>% 
  add_footnote(c("For every 10 km between the competing zip code and the zip code of interest, the covariate rate will change by the estimated coefficient"))

```

## Balance on Covariates
```{r}

df.qual.exp %>% 
  mutate(Treat = ifelse(Att > 0, 1, 0)) %>% 
  mutate(Groups = ifelse(Treat == 1 & Expandor == 0, "Treated", 
                  ifelse(Treat == 0 & Expandor == 0, "Control", 
                  ifelse(Treat == 0 & Expandor == 1, "Expandor Comp", "Expandor No Comp")))) %>% 
  filter(Year == 2015) %>%  #z & ZipFilter == 2) %>% 
  select(Sites, Zips, #MarketShare, 
         Total, Density, Pct.Rural, Pct.Urban, None:MCD, IMM2yo:DepScrn, Groups, Pct.NHA:Pct.NHW, Pct.LessThan200pctFPL, Pct.Homeless:Pct.Children, `Liabilities (as pct of assets)`, `Gross land, buildings and equipment (LBE)`, `Surplus as a pct of after dep expenses`) %>% #SwitchEMR, InterOrgHITexchange, UDSQualRep.highlowuse, 
  tbl_summary(by = Groups, 
              statistic = all_continuous() ~ "{mean} ({sd})") %>% 
  add_p(test = everything() ~ "t.test")

```

## Treated vs Rivals
```{r}

# rbind(uds.covariates %>% filter(Year == 2015 & `Grant Number` %in% c(unique(df.z.dist.plot$`Grant Number.x`)) ) %>% mutate(Class = "Incumbent"), 
#       uds.covariates %>% filter(Year == 2015 & `Grant Number` %in% c(unique(df.z.dist.plot$`Grant Number.y`)) ) %>% mutate(Class = "Rival")) %>% 
#   tbl_summary(by = Class) %>% 
#   add_p()

```

## Mapping Quality Function
```{r}
p <- seq(0, 1, 0.05)
v <- c(1:10)

for (i in 1:10){
  p <- i / 10
  y <- 1 - (1 - p)^v
  y.prime = -(1 - p)^v*log(1 - p) - y/v
  #plot(v, y, main = paste("p = ", p, sep = "")) ## For regular trade offs
  plot(v, y.prime, main = paste("p = ", p, sep = "")) ## for denominator sign checking
}

## Checking the total differential on quality to see what happens
dd <- merge(merge(seq(0.2, 0.8, 0.05), c(1:5)) %>% rename(p = x, v = y), 
        c(seq(-5, -1, 0.5), seq(1, 5, 0.5)) ) %>% 
  rename(dv = y) %>% 
  mutate(dp = (-(1 - p)^v*log(1 - p) - (1 - (1 - p)^v)/v)*( v*(1 - p)^(v-1) )^-1 * dv) %>% 
    mutate(dq = -(1 - p)^v*log(1 - p) * dv + v*(1 - p)^(v-1) * dp) %>% 
  filter(dv > -abs(v))

summary(dd[complete.cases(dd),]$dq)

dd %>% filter(dq == min(dd$dq, na.rm = TRUE))
dd %>% filter(dq == max(dd$dq, na.rm = TRUE))

plot(dd[dd$v == 3,]$dv, dd[dd$v == 3,]$dq)

ggplot(data = dd, aes(x = p, y = dq, color = dv)) + geom_point() + facet_wrap(~v)

summary(df.qual.exp$Qcomp)

```

## Placebo

```{r}

df.placebo <- df.qual.exp %>% 
  #filter(ZipFilter == 2) %>% 
select(Year, ID, Att, Pct.NHA:Pct.Children) 
# %>% 
#   inner_join(df.qual.exp %>% filter(Year == 2016) %>% mutate(ZipFilter = ifelse(`Surplus as a pct of before dep expenses` < 2, 1, 2)) %>% select(ID, ZipFilter), by = c("ID")) %>% 
#   filter(ZipFilter == 2) %>% 
#     select(Year, ID, Att, Pct.NHA:Pct.Children)

atts.placebo <- rep(0, 7)
for (i in 5:ncol(df.placebo)){

  temp <- df.placebo[,c(1:3, i)]
  colnames(temp) <- c("Year","ID","Att","Outcome")

cs.output.expand <- att_gt( yname = "Outcome",
                      tname = "Year",
                      idname = "ID",
                      gname = "Att",
                      xformla = ~1,
                      data = temp,
                      control_group = "nevertreated",
                      clustervars = "ID",
                      panel = FALSE, 
                      base_period = "universal"
                      )
cs.output.expand.agg.dyn.0 <- aggte(cs.output.expand, type = "simple", na.rm = TRUE)
cs.output.expand.agg.dyn <- aggte(cs.output.expand, type = "dynamic", na.rm = TRUE)

atts.placebo <- rbind(atts.placebo, cbind(colnames(df.placebo[,i]), round(cbind(cs.output.expand.agg.dyn.0$overall.att, 
      cs.output.expand.agg.dyn.0$overall.att - 1.96 * cs.output.expand.agg.dyn.0$overall.se, 
      cs.output.expand.agg.dyn.0$overall.att + 1.96 * cs.output.expand.agg.dyn.0$overall.se, 
      cs.output.expand.agg.dyn$overall.att, 
      cs.output.expand.agg.dyn$overall.att - cs.output.expand.agg.dyn$crit.val.egt * cs.output.expand.agg.dyn$overall.se, 
      cs.output.expand.agg.dyn$overall.att + cs.output.expand.agg.dyn$crit.val.egt * cs.output.expand.agg.dyn$overall.se), 3)))

}

#if (k == 1){atts.placebo.larger <- atts.placebo[-1,]} else {atts.placebo.smaller <- atts.placebo[-1,]}

#rownames(atts.placebo.larger) <- NULL
#atts.placebo.larger
#atts.placebo.smaller

as.data.frame(atts.placebo[-1,]) %>% 
  rename(Measure = V1, Estimate = V2, `95%CI_LB` = V3, `95%CI_UB` = V4)
```

## multiple hypothesis correction, pre-period parallel trends
```{r}

# rownames(atts) <- rep(NULL, nrow(atts))
# if (k == 1){atts.larger <- atts} else {atts.smaller <- atts}
#

# p-values rankings, Benjamini-Hochberg procedure
cbind(Sample = c(rep("Panel A: Full Sample", 3), 
                    rep("Panel B: Higher Low-Income Market Share Subgroup", 3), 
                    rep("Panel B: Lower Low-Income Market Share Subgroup", 3), 
                    rep("Panel C: Profitable Subgroup", 3), 
                    rep("Panel C: Unprofitable/Budget Neutral Subgroup", 3)), 
  data.frame(Outcome = atts[c(2:16),c(1)], Estimate = atts[c(2:16),c(2)], CI = paste("(", atts[c(2:16),3], ", ", atts[c(2:16),4], ")", sep = ""), p.value =  c(2*pnorm(-abs(as.numeric(atts[c(2:16),2]) / ((as.numeric(atts[c(2:16),4]) - as.numeric(atts[c(2:16),3])) / (2 * 1.96)))))))      %>% 
     arrange(p.value) %>%  
    mutate(rank = row_number(), 
           rank / 14 * 0.1) %>%

  kable("html", booktabs = T, col.names = c("Sample","Outcome", "Estimate", "95% CI", "p-value", "Rank", "BH-procedure Threshold")) %>%
  kable_styling(latex_options = c("striped")) %>% 
  row_spec(c(1:6), italic = T, hline_after = T, background = "lightgrey")

# parallel trends
if (k == 1){z.larger <- z[-1,]} else {z.smaller <- z[-1,]}

# length(z.larger)
# length(z.smaller)
cbind(atts[c(2:16),1], z[c(2:16),2])
cbind(atts[atts[,1] %in% c('Total revenue (unrestricted & restricted)','Total net assets','Staffing', 'Liabilities'),1], z[atts[,1] %in% c('Total revenue (unrestricted & restricted)','Total net assets','Staffing', 'Liabilities'),2])

```

## What's happening in the control? Let's think

```{r}

df.space <- df.qual.exp %>% 
  filter(Att > 0) %>% 
  select(`Grant Number`, Year) %>% 
  unique() %>% 
  left_join(uds.sites %>% select(`Grant Number`, Year, Zip), by = c("Grant Number","Year")) %>% 
  left_join(clinics.in.zips, by = c("Zip","Year")) %>%

#  filter(`Grant Number` == "H80CS00881")
  group_by(Zip) %>% 
  filter(max(Clinics) > 1) %>% 
  unique() %>% 
  left_join(uds.sites %>% select(`Grant Number`, Zip, Year), by = c("Zip","Year")) %>% 
  rename(Incumbent = `Grant Number.x`, Competitor = `Grant Number.y`) %>% 
  filter(Incumbent != Competitor) %>% 
  group_by(Zip, Incumbent, Competitor) %>% 
  summarise(YearFirstComp = min(Year))%>% 
  arrange(Incumbent)

#df.space %>%  filter(Incumbent == "H80CS00881")

df.z.2 <- df.space %>% 
  pivot_longer(Incumbent:Competitor) %>% 
  left_join(uds.zip.totals, by = c("value" = "Grant Number")) %>% 
    ungroup() %>% 
  mutate(ID = as.numeric(as.factor(paste(Zip, name, value)))) %>% 
  mutate(year.rel = `Reporting Year` - YearFirstComp, 
         Treat = ifelse(name == "Incumbent", 1, 0), 
         Post = ifelse(year.rel >= 0, 1, 0))

df.z.2$year.rel <- relevel(as.factor(df.z.2$year.rel), ref = "-1")

# Are incumbents and controls really that different?  not really
df.z.2 %>% 
  select(value, name, `Reporting Year`, None:Total) %>% 
  unique() %>% 
  left_join(uds.covariates, by = c("value" = "Grant Number", "Reporting Year" = "Year")) %>% 
  left_join(uds.6b.rates, by = c("value" = "GrantNumber", "Reporting Year" = "Year")) %>% 
  filter(`Reporting Year` == 2015) %>% 
  filter(value != 'H80CS28975') %>% 
  select(name, Total, None:Pct.Children, IMM2yo:DepScrn) %>% 
  mutate(None = None / Total, 
         MCD = MCD / Total, 
         MCR = MCR / Total, 
         COM = COM / Total) %>% 
  tbl_summary(by = name, 
              statistic = all_continuous() ~ "{mean} ({sd})") %>% 
  add_p(test = everything() ~ "t.test")

summary(lm(data = df.z.2, Total ~ Treat*year.rel))

# Competitors in the competative zip

df.space %>% 
  left_join(df.qual.exp %>% select(`Grant Number`, Att), by = c("Incumbent" = "Grant Number")) %>% 
  left_join(uds.zip, by = c("Zip" = "Zip Code","Competitor" = "Grant Number")) %>% 
  filter(`Reporting Year` %in% c(2015:2019)) %>% 
  mutate(year.rel = `Reporting Year` - Att, 
         Tx = ifelse(`Reporting Year` - Att >= 0 , 1, 0)) %>% 
  ggplot(aes(x = year.rel, y = `Total Number of Patients`, group = Tx)) + geom_point() + geom_smooth(method = "lm") #+ ylim(0, 2000)

df.space1 <- df.space %>% 
  left_join(df.qual.exp %>% select(`Grant Number`, ID, Att) %>% unique(), by = c("Incumbent" = "Grant Number")) %>% 
  merge(2015:2019) %>% 
  left_join(uds.zip, by = c("Zip" = "Zip Code","Competitor" = "Grant Number", "y" = "Reporting Year")) %>% 
  #filter(y %in% c(2015:2019)) %>% 
  mutate(year.rel = y- Att, 
         Tx = ifelse(year.rel >= 0 , 1, 0)) %>% 
  mutate(None.z = ifelse(is.na(`None_Uninsured Patients`), 0, `None_Uninsured Patients`),
          MCR.z = ifelse(is.na(`Medicare Patients`), 0, `Medicare Patients`),
          COM.z = ifelse(is.na(`Private Patients`), 0, `Private Patients`),
          MCD.z = ifelse(is.na(`Medicaid_CHIP_Other Public Patients`), 0, `Medicaid_CHIP_Other Public Patients`),
          Total.z = ifelse(is.na(`Total Number of Patients`), 0, `Total Number of Patients`),
         Grant = Competitor) %>% 
  rename(Year = y) %>% 
  select(Grant, Zip, ID, Att, Year, None.z:Total.z) %>% 
  arrange(Grant, Zip, Year) %>% 
  rbind(df.spill %>% 
  filter(Att == 0) %>% 
  rename(Grant = `Grant Number`) %>% 
  select(Grant, Zip, ID, Att, Year, None.z, MCR.z, COM.z, MCD.z, Total.z) %>% 
    inner_join(df.spill2 %>% filter(Dist == 0) %>% select(`Grant Number`) %>% unique(), by = c("Grant" = "Grant Number"))) %>% 
  mutate(Tx = ifelse(Att > 0 & Year >= Att, 1, 0))

feols(data = df.space1, Total.z ~ Tx | Zip + Year, cluster = ~Zip)

cs.output.expand <- att_gt( yname = "Total.z",
                      tname = "Year",
                      idname = "ID",
                      gname = "Att",
                      xformla = ~1,
                      data = df.space1, # %>% filter(Year > 2015),
                      control_group = "nevertreated",
                      clustervars = "ID",
                      panel = TRUE,
                      base_period = "universal"
                      )
aggte(cs.output.expand, type = "dynamic", na.rm = TRUE)
```

## Change in Full Market Benefit
```{r}

df.z.dist.plot %>% 
  left_join(uds.zip %>% select(`Grant Number`, `Reporting Year`, `Zip Code`, `Medicaid_CHIP_Other Public Patients`, `Total Number of Patients`), 
            by = c("Grant Number.y" = "Grant Number", "Zip" = "Zip Code")) %>% 
  left_join(uds.zip %>% select(`Grant Number`, `Reporting Year`, `Zip Code`, `Medicaid_CHIP_Other Public Patients`, `Total Number of Patients`), 
            by = c("Grant Number.x" = "Grant Number", "Zip" = "Zip Code", "Reporting Year")) %>% 
  rename(MCD.Rival = `Medicaid_CHIP_Other Public Patients.y`, 
         Total.Rival = `Total Number of Patients.y`, 
         MCD.Incumbent = `Medicaid_CHIP_Other Public Patients.x`, 
         Total.Incumbent = `Total Number of Patients.x`) %>% 
  left_join(uds.6b.rates %>% 
            mutate(Quality.Incumbent = (IMM2yo + Pap + ChildBMI + AdultBMI + Tobac + CRC + DepScrn) / 7) %>% 
            select(GrantNumber, Year, Quality.Incumbent), by = c("Grant Number.y" = "GrantNumber", "Reporting Year" = "Year")) %>% 
  left_join(uds.6b.rates %>% 
            mutate(Quality.Rival = (IMM2yo + Pap + ChildBMI + AdultBMI + Tobac + CRC + DepScrn) / 7) %>% 
            select(GrantNumber, Year, Quality.Rival), by = c("Grant Number.x" = "GrantNumber", "Reporting Year" = "Year")) %>% 
  mutate(YearDiff = `Reporting Year` - Year, 
         Post = ifelse(YearDiff >= 0, 1, 0)) %>% 
  filter(`Reporting Year` %in% c(2015:2019)) %>% 
  select(Zip, Att.z, `Grant Number.x`, `Grant Number.y`, `Reporting Year`:Post) %>% 
  mutate(Quality.Market = Total.Incumbent / (Total.Incumbent + Total.Rival) * Quality.Incumbent + Total.Rival / (Total.Incumbent + Total.Rival) * Quality.Rival) %>% 
  group_by(Post) %>% 
  summarise(mean(Quality.Market, na.rm = TRUE))

```