### 2023-08-13 code/data_cleaning.Rmd
sending up some data

---
title: "220410_Competition"
author: "Justin"
date: "4/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Library 
```{r}
library(tidygeocoder)
library(tidyverse)
library(did)
library(tidycensus)
library(gtsummary)
library(readxl)
library(sandwich)
library(lmtest)
library(ggpubr)
library(kableExtra)
library(tigris)

library(gridExtra)
library(cowplot)
library(Matching)
library(fixest)
library(directlabels)

`%notin%` <- Negate(`%in%`)

select <- dplyr::select

#Sys.setenv('R_MAX_VSIZE'=32000000000)

```

## Define new Tufte data theme
```{r, echo = FALSE, results = "asis", message=FALSE, warning = FALSE, fig.align="center"}

is.extrafont.installed <- function(){
  if(is.element("extrafont", installed.packages()[,1])){
    library(extrafont)
    # probably need something here to run font_import()
    return(T)
  }else{
    warning("Library extrafont installed; using system sans/serif libraries as fallback fonts. 
    To enable full font support, run: 
      install.packages('extrafont') 
      font_import()")
    return(F)
  }
}

# set font
base_font_family_tufte <- function(){
  if(is.extrafont.installed()){
    library(extrafont)
    tuftefont <- choose_font(c("Gill Sans MT", "Gill Sans", "GillSans", "Verdana", "serif"), quiet = FALSE)  
  }else{
    tuftefont <- "serif"
  }
  return(tuftefont)
}

theme_tufte_revised <- function(base_size = 11, base_family = base_font_family_tufte(), ticks = TRUE) {

  ret <- theme_bw(base_family = base_family, base_size = base_size) + 
    theme(
      axis.line = element_line(color = 'black'),
      axis.title.x = element_text(vjust = -0.3), 
      axis.title.y = element_text(vjust = 0.8),
      legend.background = element_blank(), 
      legend.key = element_blank(), 
      legend.title = element_text(face="plain"),
      panel.background = element_blank(), 
      panel.border = element_blank(),
      panel.grid = element_blank(),
      plot.background = element_blank(),
      strip.background = element_blank()
    )

  if (!ticks) {
    ret <- ret + theme(axis.ticks = element_blank())
  }

  ret
}

```

# Data

## Read in UDS

```{r}

## Zips
locs <- paste("//Users/jhm65/Dropbox/Git/UDS/UDS-", c(2014:2021), "-Full-Dataset.xlsx", sep = "")
uds.zip <- read_excel(locs[1], sheet = "HealthCenterZipCodes")

new.zip.names <- c("BHCMIS ID", "Grant Number", "Reporting Year" , "Zip Code", "Zip Code Type",  "None_Uninsured Patients" ,  "Medicaid_CHIP_Other Public Patients", "Medicare Patients", "Private Patients", "Total Number of Patients" ) 
colnames(uds.zip) <-    new.zip.names

for(i in 2:length(locs)){
  temp.zip <- read_excel(locs[i], sheet = "HealthCenterZipCodes")
  colnames(temp.zip) <- new.zip.names
  uds.zip <- rbind(uds.zip, temp.zip %>% mutate(`Zip Code` = str_pad(`Zip Code`, 5, pad = "0")))
}

## Rename
new.zip.names <- c("BHCMIS ID", "Grant Number", "Reporting Year" , "Zip Code", "Zip Code Type",  "None_Uninsured Patients" ,  "Medicaid_CHIP_Other Public Patients", "Medicare Patients", "Private Patients", "Total Number of Patients" )
colnames(uds.zip) <-    new.zip.names

## Clean

uds.zip[uds.zip$`Total Number of Patients`== '-' | uds.zip$`Total Number of Patients`== '--',]$`Total Number of Patients` <- NA
uds.zip$`Total Number of Patients` <- as.numeric(uds.zip$`Total Number of Patients`)
uds.zip <- uds.zip[!is.na(uds.zip$`Total Number of Patients`),] %>%
    mutate(`Reporting Year` = as.numeric(`Reporting Year`))

uds.zip$`None_Uninsured Patients` <- as.numeric(uds.zip$`None_Uninsured Patients`)
uds.zip$`Medicaid_CHIP_Other Public Patients` <- as.numeric(uds.zip$`Medicaid_CHIP_Other Public Patients`)
uds.zip$`Medicare Patients` <- as.numeric(uds.zip$`Medicare Patients`)
uds.zip$`Private Patients` <- as.numeric(uds.zip$`Private Patients`)

uds.zip <- cbind(uds.zip, broken.total = rowSums(uds.zip[,c(6:9)], na.rm = TRUE), missings = rowSums(is.na(uds.zip[,c(6:9)]))) %>%
  mutate(`Total Number of Patients` = as.numeric(`Total Number of Patients`)) %>%
  mutate(`Total Number of Patients` = ifelse(is.na(`Total Number of Patients`), round(missings*16/4), `Total Number of Patients`)) %>%
  mutate(`None_Uninsured Patients` = ifelse(is.na(`None_Uninsured Patients`), (`Total Number of Patients` - broken.total) * 1/missings, `None_Uninsured Patients`),
         `Medicaid_CHIP_Other Public Patients` = ifelse(is.na(`Medicaid_CHIP_Other Public Patients`), (`Total Number of Patients` - broken.total) * 1/missings, `Medicaid_CHIP_Other Public Patients`),
         `Medicare Patients` = ifelse(is.na(`Medicare Patients`), (`Total Number of Patients` - broken.total) * 1/missings, `Medicare Patients`),
         `Private Patients` = ifelse(is.na(`Private Patients`), (`Total Number of Patients` - broken.total) * 1/missings, `Private Patients`))

# ## Old version with missingness still in there  
# uds.zip$`Reporting Year` <- as.numeric(uds.zip$`Reporting Year`)
# uds.zip$`Total Number of Patients` <- as.numeric(uds.zip$`Total Number of Patients`)
# uds.zip$`None_Uninsured Patients` <- as.numeric(uds.zip$`None_Uninsured Patients`)
# uds.zip$`Medicaid_CHIP_Other Public Patients` <- as.numeric(uds.zip$`Medicaid_CHIP_Other Public Patients`)
# uds.zip$`Medicare Patients` <- as.numeric(uds.zip$`Medicare Patients`)
# uds.zip$`Private Patients` <- as.numeric(uds.zip$`Private Patients`)
# 
# 
# uds.zip[is.na(uds.zip$`Total Number of Patients`),]$`Total Number of Patients` <- 0
# uds.zip[grep("-", uds.zip$`None_Uninsured Patients`),]$`None_Uninsured Patients` <- 0
# uds.zip[grep("-", uds.zip$`Medicaid_CHIP_Other Public Patients`),]$`Medicaid_CHIP_Other Public Patients` <- 0
# uds.zip[grep("-", uds.zip$`Medicare Patients`),]$`Medicare Patients` <- 0
# uds.zip[grep("-", uds.zip$`Private Patients`),]$`Private Patients` <- 0
# 
uds.zip.totals <- uds.zip %>%
  select(`Grant Number`, `Reporting Year`, `None_Uninsured Patients`:`Total Number of Patients`) %>%
  group_by(`Grant Number`, `Reporting Year`) %>% summarise(None = sum(`None_Uninsured Patients`, na.rm = T),
                                                           MCR = sum(`Medicare Patients`, na.rm = T),
                                                           COM = sum(`Private Patients`, na.rm = T),
                                                           MCD = sum(`Medicaid_CHIP_Other Public Patients`, na.rm = T),
                                                           Total = sum(`Total Number of Patients`, na.rm = T))

uds.zip <- uds.zip[uds.zip$`Zip Code Type` == "ZipCode",]

## Sites
locs <- paste("//Users/jhm65/Dropbox/Git/UDS/UDS-", c(2014:2021), "-Full-Dataset.xlsx", sep = "")

uds.sites <- read_excel(locs[1], sheet = "HealthCenterSiteInfo") %>% 
  select(GrantNumber, SiteName, SiteType, LocationType, LocationSetting, OperationalSchedule, CalendarSchedule, TotalWeeklyHoursOfOperation, SiteStreetAddress, SiteCity, SiteState, SiteZIPCode) %>% 
  mutate(Year = 2014)

colnames(uds.sites) <- c("Grant Number", "Site Name", "Site Type", "Location Type", "Location Setting", "Operational Schedule", "Calendar Schedule", "Total Weekly Hours Of Operation", "Site Street Address", "Site City", "Site State", "Site ZIP Code", "Year")

for(i in 2:(length(locs)-1)){
  uds.sites <- rbind(uds.sites, read_excel(locs[i], sheet = "HealthCenterSiteInfo") %>% 
                       select("Grant Number", "Site Name", "Site Type", "Location Type", "Location Setting", "Operational Schedule", "Calendar Schedule", "Total Weekly Hours Of Operation", "Site Street Address", "Site City", "Site State", "Site ZIP Code") %>% 
  mutate(Year = 2013 + i))
}

temp.sites <- read_excel(locs[i+1], sheet = "HealthCenterSiteInfo" )%>% 
  select(GrantNumber, SiteName, SiteType, LocationType, LocationSetting, OperationalSchedule, CalendarSchedule, TotalWeeklyHoursOfOperation, SiteStreetAddress, SiteCity, SiteState, SiteZIPCode) %>% 
  mutate(Year = 2021)
colnames(temp.sites) <- c("Grant Number", "Site Name", "Site Type", "Location Type", "Location Setting", "Operational Schedule", "Calendar Schedule", "Total Weekly Hours Of Operation", "Site Street Address", "Site City", "Site State", "Site ZIP Code", "Year")

uds.sites <- rbind(uds.sites, temp.sites)

uds.sites$Zip <- substr(uds.sites$`Site ZIP Code`, 1, 5)

uds.sites <- uds.sites %>% 
  filter(`Site State` %notin% c('AK','HI','GU','VI','PW','MH','AS','FM','PR','MP'))

uds.sites[grep(pattern = "-", uds.sites$Zip),]$Zip <- NA

```

## Clean up some of the missing zip codes

```{r}

tbcoded <- uds.sites[is.na(uds.sites$Zip),]
coded <- geo(street = tbcoded[c(1:1000),]$`Site Street Address`, 
             city = tbcoded[c(1:1000),]$`Site City`, 
             state = tbcoded[c(1:1000),]$`Site State`, 
             method = "census", 
             full_results = TRUE)

for (g in 1:9){
  g
coded <- rbind(coded, 
               geo(street = tbcoded[c((1 + g*1000):((g+1)*1000)),]$`Site Street Address`, 
             city = tbcoded[c((1 + g*1000):((g+1)*1000)),]$`Site City`, 
             state = tbcoded[c((1 + g*1000):((g+1)*1000)),]$`Site State`, 
             method = "census", 
             full_results = TRUE) )
}

coded <- rbind(coded, 
               geo(street = tbcoded[c(10001:10975),]$`Site Street Address`, 
             city = tbcoded[c(10001:10975),]$`Site City`, 
             state = tbcoded[c(10001:10975),]$`Site State`, 
             method = "census", 
             full_results = TRUE) )

coded[grep(pattern = "[[:digit:]]{5}$", coded$matched_address),]

uds.sites[is.na(uds.sites$Zip),]$Zip <- str_extract(coded$matched_address, pattern = "[[:digit:]]{5}$")
uds.sites[is.na(uds.sites$Zip),]

uds.sites.temp <- uds.sites

```

```{r}

uds.sites <- uds.sites.temp

uds.sites.temp %>% filter(`Grant Number` %in% c("H80CS00881") & Year == 2016)

chc <- read.csv("//Users/jhm65/Dropbox/Git/Competition/data_raw/chc.csv")

chc %>% filter(Health.Center.Number %in% c("H80CS00881"))

uds.sites[is.na(uds.sites$Zip),] <- uds.sites[is.na(uds.sites$Zip),] %>%
  left_join(chc, by = c("Grant Number" = "Health.Center.Number", "Site Street Address" = "Site.Address")) %>%
  mutate(Zip = ifelse(is.na(Zip), substr(Health.Center.Organization.ZIP.Code, 1, 5), Zip)) %>%
  select(`Grant Number`:Zip) %>%
  unique()

uds.sites[is.na(uds.sites$Zip),] <- uds.sites[is.na(uds.sites$Zip),] %>%
  left_join(chc, by = c("Grant Number" = "Health.Center.Number", "Site Name" = "Site.Name")) %>%
  mutate(Zip = ifelse(is.na(Zip), substr(Health.Center.Organization.ZIP.Code, 1, 5), Zip)) %>%
  select(`Grant Number`:Zip) %>%
  unique()

uds.sites[is.na(uds.sites$Zip),] 
```

```{r}

table(uds.sites.temp$`Location Setting`)

uds.sites <- uds.sites %>% #filter(`Location Type` == "Permanent" & `Location Setting` %in% c("All Other Clinic Types")) %>% 
  mutate(Access = as.numeric(ifelse(`Total Weekly Hours Of Operation` == "-" & `Operational Schedule` == "Full-Time", 40, 
                  ifelse(`Total Weekly Hours Of Operation` == "-" & `Operational Schedule` == "Part-Time", 10, `Total Weekly Hours Of Operation`))))

uds.sites[is.na(uds.sites$Zip),]

uds.sites %>% 
  arrange(`Total Weekly Hours Of Operation`, `Operational Schedule` )

# z <- geocode(chc.term$SitePhysicalAddress)
# 
# 
# y <- geo(address = chc.term$SitePhysicalAddress, method = "census")
# 
# write.csv(cbind(y[,1], y[,3], y[,2], z), file = "geocodedClosures.csv")

```

## Read in Quality data
```{r}

uds.6b.2012 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2014-Full-Dataset.xlsx", sheet = "Table6BClinicalmeasures", skip = 0) 
uds.6b.2013 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2014-Full-Dataset.xlsx", sheet = "Table6BClinicalmeasures", skip = 0)

uds.6b.2014 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2014-Full-Dataset.xlsx", sheet = "Table6BClinicalmeasures", skip = 0) 
uds.6b.2015 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2015-Full-Dataset.xlsx", sheet = "Table6BClinicalmeasures", skip = 1) 
uds.6b.2016 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2016-Full-Dataset.xlsx", sheet = "Table6BClinicalmeasures", skip = 0)

uds.6b.2017 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2017-Full-Dataset.xlsx", sheet = "Table6BClinicalmeasures", skip = 0) 
uds.6b.2018 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2018-Full-Dataset.xlsx", sheet = "Table6BClinicalmeasures", skip = 0) 
uds.6b.2019 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2019-Full-Dataset.xlsx", sheet = "Table6BClinicalmeasures", skip = 0)

uds.6b.2020 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2020-Full-Dataset.xlsx", sheet = "Table6BClinicalmeasures", skip = 0) 
uds.6b.2021 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2021-Full-Dataset.xlsx", sheet = "Table6BClinicalmeasures", skip = 0)

# rbind(
# colnames(uds.6b.2014[,c(2, grep(pattern = "^%", colnames(uds.6b.2014))[-6])]),
# colnames(uds.6b.2015[,c(2, grep(pattern = "^%", colnames(uds.6b.2015))[-6])]),
# colnames(uds.6b.2016[,c(2, grep(pattern = "^%", colnames(uds.6b.2016))[-6])]),
# colnames(uds.6b.2017[,c(2, grep(pattern = "^%", colnames(uds.6b.2017))[-6])]),
# colnames(uds.6b.2018[,c(2, grep(pattern = "^%", colnames(uds.6b.2018))[-6])]),
# colnames(uds.6b.2019[,c(2, grep(pattern = "^%", colnames(uds.6b.2019))[-6])]),
# colnames(uds.6b.2020[,c(2, grep(pattern = "^%", colnames(uds.6b.2020))[-c(3, 11, 13)])]))
# 
# 
# rbind(
# colnames(uds.6b.2014[,c(2, grep(pattern = "^Total", colnames(uds.6b.2014))[-6])]),
# colnames(uds.6b.2015[,c(2, grep(pattern = "^Total", colnames(uds.6b.2015))[-6])]),
# colnames(uds.6b.2016[,c(2, grep(pattern = "^Total", colnames(uds.6b.2016))[-6])]),
# colnames(uds.6b.2017[,c(2, grep(pattern = "^Total", colnames(uds.6b.2017))[-6])]),
# colnames(uds.6b.2018[,c(2, grep(pattern = "^Total", colnames(uds.6b.2018))[-6])]),
# colnames(uds.6b.2019[,c(2, grep(pattern = "^Total", colnames(uds.6b.2019))[-6])]),
# colnames(uds.6b.2020[,c(2, grep(pattern = "^Total", colnames(uds.6b.2020))[-c(3, 11, 13)])]))

namingshit <- c("GrantNumber", "IMM2yo", "Pap", "ChildBMI","AdultBMI","Tobac","CRC","DepScrn")
namingshit.denom <- c("GrantNumber", "IMM2yo.denom", "Pap.denom", "ChildBMI.denom","AdultBMI.denom","Tobac.denom","CRC.denom","DepScrn.denom")

uds.6b.rates <- rbind(
  uds.6b.2014[,c(2, grep(pattern = "^%", colnames(uds.6b.2014))[-6])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit), function(x) namingshit) %>% mutate(Year = 2014), 
  uds.6b.2015[,c(2, grep(pattern = "^%", colnames(uds.6b.2015))[-6])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit), function(x) namingshit) %>% mutate(Year = 2015),
  uds.6b.2016[,c(2, grep(pattern = "^%", colnames(uds.6b.2016))[-6])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit), function(x) namingshit) %>% mutate(Year = 2016),

  uds.6b.2017[,c(2, grep(pattern = "^%", colnames(uds.6b.2017))[-6])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit), function(x) namingshit) %>% mutate(Year = 2017),
  uds.6b.2018[,c(2, grep(pattern = "^%", colnames(uds.6b.2018))[-6])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit), function(x) namingshit) %>% mutate(Year = 2018),

  uds.6b.2019[,c(2, grep(pattern = "^%", colnames(uds.6b.2019))[-6])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit), function(x) namingshit) %>% mutate(Year = 2019),
  uds.6b.2020[,c(2, grep(pattern = "^%", colnames(uds.6b.2020))[-c(3, 11, 13)])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit), function(x) namingshit) %>% mutate(Year = 2020), 
  uds.6b.2021[,c(2, grep(pattern = "^%", colnames(uds.6b.2021))[-c(3, 11, 13)])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit), function(x) namingshit) %>% mutate(Year = 2021))

uds.6b.rates$IMM2yo <- gsub(pattern = "%$", replacement = "", uds.6b.rates$IMM2yo)
uds.6b.rates$Pap <- gsub(pattern = "%$", replacement = "", uds.6b.rates$Pap)
uds.6b.rates$ChildBMI <- gsub(pattern = "%$", replacement = "", uds.6b.rates$ChildBMI)
uds.6b.rates$AdultBMI <- gsub(pattern = "%$", replacement = "", uds.6b.rates$AdultBMI)

uds.6b.rates$Tobac <- gsub(pattern = "%$", replacement = "", uds.6b.rates$Tobac)
uds.6b.rates$CRC <- gsub(pattern = "%$", replacement = "", uds.6b.rates$CRC)
uds.6b.rates$DepScrn <- gsub(pattern = "%$", replacement = "", uds.6b.rates$DepScrn)

uds.6b.denom <- rbind(
  uds.6b.2014[,c(2, grep(pattern = "^Total", colnames(uds.6b.2014))[-6])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit.denom), function(x) namingshit.denom) %>% mutate(Year = 2014), 
  uds.6b.2015[,c(2, grep(pattern = "^Total", colnames(uds.6b.2015))[-6])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit.denom), function(x) namingshit.denom) %>% mutate(Year = 2015),
  uds.6b.2016[,c(2, grep(pattern = "^Total", colnames(uds.6b.2016))[-6])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit.denom), function(x) namingshit.denom) %>% mutate(Year = 2016),

  uds.6b.2017[,c(2, grep(pattern = "^Total", colnames(uds.6b.2017))[-6])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit.denom), function(x) namingshit.denom) %>% mutate(Year = 2017),
  uds.6b.2018[,c(2, grep(pattern = "^Total", colnames(uds.6b.2018))[-6])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit.denom), function(x) namingshit.denom) %>% mutate(Year = 2018),

  uds.6b.2019[,c(2, grep(pattern = "^Total", colnames(uds.6b.2019))[-6])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit.denom), function(x) namingshit.denom) %>% mutate(Year = 2019),
  uds.6b.2020[,c(2, grep(pattern = "^Total", colnames(uds.6b.2020))[-c(3, 11, 13)])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit.denom), function(x) namingshit.denom) %>% mutate(Year = 2020),
  uds.6b.2021[,c(2, grep(pattern = "^Total", colnames(uds.6b.2021))[-c(3, 11, 13)])][,c(1:6, 9, 11)] %>% 
  rename_all(vars(namingshit.denom), function(x) namingshit.denom) %>% mutate(Year = 2021))

for (k in 2:8) uds.6b.rates[grep("-", unlist(uds.6b.rates[,k])),k] <- NA
for (k in 2:8) uds.6b.denom[grep("-", unlist(uds.6b.denom[,k])),k] <- NA

uds.6b.rates <-  data.frame(GrantNumber = uds.6b.rates$GrantNumber, Year = uds.6b.rates$Year, lapply(uds.6b.rates[,c(2:8)], as.numeric)) 
uds.6b.denom <-  data.frame(GrantNumber = uds.6b.denom$GrantNumber, Year = uds.6b.denom$Year, lapply(uds.6b.denom[,c(2:8)], as.numeric))

uds.6b.rates[uds.6b.rates$Year == 2014,c(3:9)] <- uds.6b.rates[uds.6b.rates$Year == 2014,c(3:9)] * 100
uds.6b.rates[,c(3:9)] <- uds.6b.rates[,c(3:9)]

## Finding highest-scoring quality FQHC
# uds.6b.rates %>% 
#   mutate(Qcomp = (IMM2yo + Pap + ChildBMI + AdultBMI + Tobac + CRC + DepScrn) / 7) %>% 
#   arrange(desc(Qcomp)) %>% 
#   filter(Year == 2021) %>% 
#   select(GrantNumber, Year, Qcomp) %>% 
#   left_join(chc %>% select(Health.Center.Number, Health.Center.Name, Health.Center.Organization.City, Health.Center.Organization.State) %>% unique(), by = c("GrantNumber" = "Health.Center.Number")) %>% 
#   left_join(uds.6b.denom, by = c("GrantNumber","Year"))

```

## Read in Technology Data
```{r}

uds.HIT.2014 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2014-Full-Dataset.xlsx", sheet = "EHRInformation", skip = 0) %>% 
  select(GrantNumber, Tehr_L1b_Ca, Tehr_L2_Ca,	Tehr_L3_Ca, Tehr_L4_Ca, Tehr_L5_Ca, Tehr_L7_Ca)
uds.HIT.2015 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2015-Full-Dataset.xlsx", sheet = "EHRInformation", skip = 2) %>% 
  select(GrantNumber, Tehr_L1b_Ca, Tehr_L2_Ca,	Tehr_L3_Ca, Tehr_L4_Ca, Tehr_L5_Ca, Tehr_L7_Ca)
uds.HIT.2016 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2016-Full-Dataset.xlsx", sheet = "HITInformation", skip = 1) %>% 
  select(GrantNumber, Tehr_L1b_Ca, Tehr_L2_Ca,	Tehr_L3_Ca, Tehr_L4_Ca, Tehr_L5_Ca, Tehr_L7_Ca)

uds.HIT.2017 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2017-Full-Dataset.xlsx", sheet = "HITInformation", skip = 1) %>% 
  select(GrantNumber, Tehr_L1b_Ca, Tehr_L2_Ca,	Tehr_L3_Ca, Tehr_L4_Ca, Tehr_L5_Ca, Tehr_L7_Ca)
uds.HIT.2018 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2018-Full-Dataset.xlsx", sheet = "HITInformation", skip = 1) %>% 
  select(GrantNumber, Tehr_L1b_Ca, Tehr_L2_Ca,	Tehr_L3_Ca, Tehr_L4_Ca, Tehr_L5_Ca, Tehr_L7_Ca)
uds.HIT.2019 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2019-Full-Dataset.xlsx", sheet = "HITInformation", skip = 1) %>% 
  select(GrantNumber, Tehr_L1b_Ca, Tehr_L2_Ca,	Tehr_L3_Ca, Tehr_L4_Ca, Tehr_L5_Ca, Tehr_L7_Ca)

uds.HIT.2020 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2020-Full-Dataset.xlsx", sheet = "HITInformation", skip = 1) %>%
  select(GrantNumber, Tehr_L1b_Ca, Tehr_L4_Ca, Tehr_L7_Ca)
uds.HIT.2021 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2021-Full-Dataset.xlsx", sheet = "HITInformation", skip = 0) %>%
  select(GrantNumber, Tehr_L1b_Ca, Tehr_L4_Ca, Tehr_L7_Ca)

uds.HIT.2020 <- data.frame(GrantNumber = uds.HIT.2020[,1], Tehr_L1b_Ca = uds.HIT.2020[,2], Tehr_L2_Ca = NA,	Tehr_L3_Ca = NA, Tehr_L4_Ca = uds.HIT.2020[,3], Tehr_L5_Ca = NA, Tehr_L7_Ca = uds.HIT.2020[,4])
uds.HIT.2021 <- data.frame(GrantNumber = uds.HIT.2021[,1], Tehr_L1b_Ca = uds.HIT.2021[,2], Tehr_L2_Ca = NA,	Tehr_L3_Ca = NA, Tehr_L4_Ca = uds.HIT.2021[,3], Tehr_L5_Ca = NA, Tehr_L7_Ca = uds.HIT.2021[,4])

namevec <- c("Year", "Grant Number", "SwitchEMR","eScripts","eDecisionSupport","InterOrgHITexchange","PatientPortals","UDSQualRep")
uds.HIT <- rbind(cbind(Year = 2014, uds.HIT.2014),
                  cbind(Year = 2015, uds.HIT.2015),
                  cbind(Year = 2016, uds.HIT.2016),
                  cbind(Year = 2017, uds.HIT.2017),

                  cbind(Year = 2018, uds.HIT.2018),
                  cbind(Year = 2019, uds.HIT.2019), 
                  cbind(Year = 2020, uds.HIT.2020),
                  cbind(Year = 2021, uds.HIT.2021)) %>%

  rename_all(vars(namevec), function(x) namevec) %>% 
  mutate(UDSQualRep.levels = ifelse(UDSQualRep == "-", NA, 
                             ifelse(UDSQualRep == "We do not use the EHR", 0, 
                             ifelse(UDSQualRep == "We use the EHR but only to access individual patient charts", 1, 
                             ifelse(UDSQualRep == "We use the EHR in combination with another data analytic system", 3, 2))))) %>% 
  mutate(UDSQualRep.highlowuse = ifelse(UDSQualRep == "-", NA, 
                             ifelse(UDSQualRep == "We do not use the EHR" | UDSQualRep == "We use the EHR but only to access individual patient charts", 0, 1))) %>%  
  mutate(SwitchEMR = ifelse(SwitchEMR == "-", NA, ifelse(SwitchEMR == "Yes", 1, 0)),
         eScripts = ifelse(eScripts == "-" | eScripts == "Not Sure", NA, ifelse(eScripts == "Yes", 1, 0)),
         eDecisionSupport = ifelse(eDecisionSupport == "-" | eDecisionSupport == "Not Sure", NA, ifelse(eDecisionSupport == "Yes", 1, 0)),
         InterOrgHITexchange = ifelse(InterOrgHITexchange == "-" | InterOrgHITexchange == "Not Sure", NA, ifelse(InterOrgHITexchange %in% c("No","None of the above"), 0, 1)),
         PatientPortals = ifelse(PatientPortals == "-" | PatientPortals == "Not Sure", NA, ifelse(PatientPortals %in% c("No","No we do not engage patients using HIT"), 0, 1)))

uds.HIT[uds.HIT$UDSQualRep.levels == 0 & !is.na(uds.HIT$UDSQualRep.levels) & is.na(uds.HIT$SwitchEMR),]$SwitchEMR <- 0
uds.HIT[uds.HIT$UDSQualRep.levels == 0 & !is.na(uds.HIT$UDSQualRep.levels) & is.na(uds.HIT$eScripts),]$eScripts <- 0
uds.HIT[uds.HIT$UDSQualRep.levels == 0 & !is.na(uds.HIT$UDSQualRep.levels) & is.na(uds.HIT$eDecisionSupport),]$eDecisionSupport <- 0
uds.HIT[uds.HIT$UDSQualRep.levels == 0 & !is.na(uds.HIT$UDSQualRep.levels) & is.na(uds.HIT$InterOrgHITexchange),]$InterOrgHITexchange <- 0
uds.HIT[uds.HIT$UDSQualRep.levels == 0 & !is.na(uds.HIT$UDSQualRep.levels) & is.na(uds.HIT$PatientPortals),]$PatientPortals <- 0

# namevec <- c("Year", "Grant Number", "SwitchEMR","InterOrgHITexchange","UDSQualRep")
# uds.HIT <- rbind(cbind(Year = 2014, uds.HIT.2014 %>% select(GrantNumber, Tehr_L1b_Ca, Tehr_L4_Ca, Tehr_L7_Ca)),
#                   cbind(Year = 2015, uds.HIT.2015 %>% select(GrantNumber, Tehr_L1b_Ca, Tehr_L4_Ca, Tehr_L7_Ca)),
#                   cbind(Year = 2016, uds.HIT.2016 %>% select(GrantNumber, Tehr_L1b_Ca, Tehr_L4_Ca, Tehr_L7_Ca)),
#                   cbind(Year = 2017, uds.HIT.2017 %>% select(GrantNumber, Tehr_L1b_Ca, Tehr_L4_Ca, Tehr_L7_Ca)),
#                   
#                   cbind(Year = 2018, uds.HIT.2018 %>% select(GrantNumber, Tehr_L1b_Ca, Tehr_L4_Ca, Tehr_L7_Ca)),
#                   cbind(Year = 2019, uds.HIT.2019 %>% select(GrantNumber, Tehr_L1b_Ca, Tehr_L4_Ca, Tehr_L7_Ca)),          
#                   cbind(Year = 2020, uds.HIT.2020)) %>% 
#   
#   rename_all(vars(namevec), function(x) namevec) %>% 
#   mutate(UDSQualRep.levels = ifelse(UDSQualRep == "-", NA, 
#                              ifelse(UDSQualRep == "We do not use the EHR", 0, 
#                              ifelse(UDSQualRep == "We use the EHR but only to access individual patient charts", 1, 
#                              ifelse(UDSQualRep == "We use the EHR in combination with another data analytic system", 3, 2))))) %>% 
#   mutate(UDSQualRep.highlowuse = ifelse(UDSQualRep == "-", NA, 
#                              ifelse(UDSQualRep == "We do not use the EHR" | UDSQualRep == "We use the EHR but only to access individual patient charts", 0, 1))) %>%  
#   mutate(SwitchEMR = ifelse(SwitchEMR == "-", NA, ifelse(SwitchEMR == "Yes", 1, 0)),
#          InterOrgHITexchange = ifelse(InterOrgHITexchange == "-" | InterOrgHITexchange == "Not Sure", NA, ifelse(InterOrgHITexchange %in% c("No","None of the above"), 0, 1)))

```

## Read in Covariates
```{r}
# Table 3B
uds.3B.2014 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2014-Full-Dataset.xlsx", sheet = "Table3B", skip = 0) %>% select(GrantNumber, T3b_L1_Cb, T3b_L3_Cb, T3b_L8_Ca, T3b_L5_Cb, T3b_L8_Cd, T3b_L12_Ca) %>% rename(T3b_L26_Ca = T3b_L8_Cd)
uds.3B.2015 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2015-Full-Dataset.xlsx", sheet = "Table3B", skip = 2) %>% select(GrantNumber, T3b_L1_Cb, T3b_L3_Cb, T3b_L8_Ca, T3b_L5_Cb, T3b_L8_Cd, T3b_L12_Ca) %>% rename(T3b_L26_Ca = T3b_L8_Cd)
uds.3B.2016 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2016-Full-Dataset.xlsx", sheet = "Table3B", skip = 1) %>% select(GrantNumber, T3b_L1_Cb, T3b_L3_Cb, T3b_L8_Ca, T3b_L5_Cb, T3b_L26_Ca, T3b_L12_Ca)

uds.3B.2017 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2017-Full-Dataset.xlsx", sheet = "Table3B", skip = 1) %>% select(GrantNumber, T3b_L1_Cb, T3b_L3_Cb, T3b_L8_Ca, T3b_L5_Cb, T3b_L26_Ca, T3b_L12_Ca)
uds.3B.2018 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2018-Full-Dataset.xlsx", sheet = "Table3B", skip = 1) %>% select(GrantNumber, T3b_L1_Cb, T3b_L3_Cb, T3b_L8_Ca, T3b_L5_Cb, T3b_L26_Ca, T3b_L12_Ca)
uds.3B.2019 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2019-Full-Dataset.xlsx", sheet = "Table3B", skip = 1) %>% select(GrantNumber, T3b_L1_Cb, T3b_L3_Cb, T3b_L8_Ca, T3b_L5_Cb, T3b_L26_Ca, T3b_L12_Ca)

uds.3B.2020 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2020-Full-Dataset.xlsx", sheet = "Table3B", skip = 1) %>% select(GrantNumber, T3b_L1_Cb, T3b_L3_Cb, T3b_L8_Ca, T3b_L5_Cb, T3b_L26_Ca, T3b_L12_Ca)
uds.3B.2021 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2021-Full-Dataset.xlsx", sheet = "Table3B", skip = 0) %>% select(GrantNumber, T3b_L1_Cb, T3b_L3_Cb, T3b_L8_Ca, T3b_L5_Cb, T3b_L26_Ca, T3b_L12_Ca)

namevec <- c("Year", "Grant Number", "NHA", "NHB", "HL", "NHW", "Total", "Eng2ndLang")
uds.3B <- rbind(cbind(Year = 2014, uds.3B.2014),
                  cbind(Year = 2015, uds.3B.2015),
                  cbind(Year = 2016, uds.3B.2016),
                  cbind(Year = 2017, uds.3B.2017),

                  cbind(Year = 2018, uds.3B.2018),
                  cbind(Year = 2019, uds.3B.2019), 
                  cbind(Year = 2020, uds.3B.2020),
                  cbind(Year = 2021, uds.3B.2021)) %>%

  rename_all(vars(namevec), function(x) namevec)

# Table 4
uds.4.2014 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2014-Full-Dataset.xlsx", sheet = "Table4", skip = 0) %>% select(GrantNumber, T4_L1_Ca, T4_L2_Ca, T4_L3_Ca, T4_L4_Ca, T4_L5_Ca, T4_L16_Ca, T4_L23_Ca, T4_L12_Ca, T4_L12_Cb) 
uds.4.2015 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2015-Full-Dataset.xlsx", sheet = "Table4", skip = 2) %>% select(GrantNumber, T4_L1_Ca, T4_L2_Ca, T4_L3_Ca, T4_L4_Ca, T4_L5_Ca, T4_L16_Ca, T4_L23_Ca, T4_L12_Ca, T4_L12_Cb) 
uds.4.2016 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2016-Full-Dataset.xlsx", sheet = "Table4", skip = 1) %>% select(GrantNumber, T4_L1_Ca, T4_L2_Ca, T4_L3_Ca, T4_L4_Ca, T4_L5_Ca, T4_L16_Ca, T4_L23_Ca, T4_L12_Ca, T4_L12_Cb)

uds.4.2017 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2017-Full-Dataset.xlsx", sheet = "Table4", skip = 1) %>% select(GrantNumber, T4_L1_Ca, T4_L2_Ca, T4_L3_Ca, T4_L4_Ca, T4_L5_Ca, T4_L16_Ca, T4_L23_Ca, T4_L12_Ca, T4_L12_Cb) 
uds.4.2018 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2018-Full-Dataset.xlsx", sheet = "Table4", skip = 1) %>% select(GrantNumber, T4_L1_Ca, T4_L2_Ca, T4_L3_Ca, T4_L4_Ca, T4_L5_Ca, T4_L16_Ca, T4_L23_Ca, T4_L12_Ca, T4_L12_Cb) 
uds.4.2019 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2019-Full-Dataset.xlsx", sheet = "Table4", skip = 1) %>% select(GrantNumber, T4_L1_Ca, T4_L2_Ca, T4_L3_Ca, T4_L4_Ca, T4_L5_Ca, T4_L16_Ca, T4_L23_Ca, T4_L12_Ca, T4_L12_Cb)

uds.4.2020 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2020-Full-Dataset.xlsx", sheet = "Table4", skip = 1) %>% select(GrantNumber, T4_L1_Ca, T4_L2_Ca, T4_L3_Ca, T4_L4_Ca, T4_L5_Ca, T4_L16_Ca, T4_L23_Ca, T4_L12_Ca, T4_L12_Cb) 
uds.4.2021 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2021-Full-Dataset.xlsx", sheet = "Table4", skip = 0) %>% select(GrantNumber, T4_L1_Ca, T4_L2_Ca, T4_L3_Ca, T4_L4_Ca, T4_L5_Ca, T4_L16_Ca, T4_L23_Ca, T4_L12_Ca, T4_L12_Cb)

namevec <- c("Year", "Grant Number", "<100%", "101-150%", "151-200%", ">200%", "Unknown", "FarmWorker", "Homeless", "0-17yo", "18+yo")
uds.4 <- rbind(cbind(Year = 2014, uds.4.2014),
                  cbind(Year = 2015, uds.4.2015),
                  cbind(Year = 2016, uds.4.2016),
                  cbind(Year = 2017, uds.4.2017),

                  cbind(Year = 2018, uds.4.2018),
                  cbind(Year = 2019, uds.4.2019), 
                  cbind(Year = 2020, uds.4.2020),
                  cbind(Year = 2021, uds.4.2021)) %>%

  rename_all(vars(namevec), function(x) namevec)

# Put covariate tables together

uds.covariates <- inner_join(uds.3B, uds.4, by = c("Year","Grant Number"))

for (i in 3:ncol(uds.covariates)){
  uds.covariates[,i] <- as.numeric(gsub("-", "", uds.covariates[,i]))
}

#uds.covariates %>% filter(`Grant Number` == "H80CS24200")

uds.covariates <-  uds.covariates %>% 
    mutate(Pct.NHA = NHA / Total * 100,
         Pct.NHB = NHB / Total * 100,
         Pct.HL = HL / Total * 100,
         Pct.NHW = NHW / Total * 100,
         Pct.ESL = Eng2ndLang / Total * 100,
         Pct.LessThan100pctFPL = ifelse(Total == Unknown, 100, `<100%` / (Total- Unknown) * 100),
         Pct.LessThan200pctFPL = ifelse(Total == Unknown, 100, (`<100%` + `101-150%` + `151-200%`) / (Total - Unknown) * 100),
         Pct.FarmWorker = FarmWorker / Total * 100,
         Pct.Homeless = Homeless / Total * 100,
         Pct.Children = `0-17yo` / Total * 100) %>% 
    select(Year, `Grant Number`, Pct.NHA:Pct.Children)

```

## Read in Clinical Outcomes
```{r}
uds.7.2014 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2014-Full-Dataset.xlsx", sheet = "Table7_1", skip = 0) %>% select(GrantNumber, T7_Li_C1a, T7_Li_C1b,	T7_Li_C1c,	T7_Li_C1d, T7_Li_C2a,	T7_Li_C2b,	T7_Li_C2c) 
uds.7.2015 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2015-Full-Dataset.xlsx", sheet = "Table7_1", skip = 2) %>% select(GrantNumber, T7_Li_C1a, T7_Li_C1b,	T7_Li_C1c,	T7_Li_C1d, T7_Li_C2a,	T7_Li_C2b,	T7_Li_C2c) 
uds.7.2016 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2016-Full-Dataset.xlsx", sheet = "Table7_1", skip = 1) %>% select(GrantNumber, T7_Li_C1a, T7_Li_C1b,	T7_Li_C1c,	T7_Li_C1d, T7_Li_C2a,	T7_Li_C2b,	T7_Li_C2c)

uds.7.2017 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2017-Full-Dataset.xlsx", sheet = "Table7_1", skip = 1) %>% select(GrantNumber, T7_Li_C1a, T7_Li_C1b,	T7_Li_C1c,	T7_Li_C1d, T7_Li_C2a,	T7_Li_C2b,	T7_Li_C2c) 
uds.7.2018 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2018-Full-Dataset.xlsx", sheet = "Table7_1", skip = 1) %>% select(GrantNumber, T7_Li_C1a, T7_Li_C1b,	T7_Li_C1c,	T7_Li_C1d, T7_Li_C2a,	T7_Li_C2b,	T7_Li_C2c) 
uds.7.2019 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2019-Full-Dataset.xlsx", sheet = "Table7_1", skip = 1) %>% select(GrantNumber, T7_Li_C1a, T7_Li_C1b,	T7_Li_C1c,	T7_Li_C1d, T7_Li_C2a,	T7_Li_C2b,	T7_Li_C2c)

uds.7.2020 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2020-Full-Dataset.xlsx", sheet = "Table7_1", skip = 1) %>% select(GrantNumber, T7_Li_C1a, T7_Li_C1b,	T7_Li_C1c,	T7_Li_C1d, T7_Li_C2a,	T7_Li_C2b,	T7_Li_C2c) 
uds.7.2021 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2021-Full-Dataset.xlsx", sheet = "Table7_1", skip = 0) %>% select(GrantNumber, T7_Li_C1a, T7_Li_C1b,	T7_Li_C1c,	T7_Li_C1d, T7_Li_C2a,	T7_Li_C2b,	T7_Li_C2c)

namevec <- c("Year", "Grant Number", "TotalDeliveries", "BirthsLT1500","Births1500to2499","BirthsGTE2500","TotalHTN","Sample","Controlcount")
uds.7 <- rbind(cbind(Year = 2014, uds.7.2014),
                  cbind(Year = 2015, uds.7.2015),
                  cbind(Year = 2016, uds.7.2016),
                  cbind(Year = 2017, uds.7.2017),

                  cbind(Year = 2018, uds.7.2018),
                  cbind(Year = 2019, uds.7.2019), 
                  cbind(Year = 2020, uds.7.2020),
                  cbind(Year = 2021, uds.7.2021)) %>%

  rename_all(vars(namevec), function(x) namevec)

for (i in 3:ncol(uds.7)){
  uds.7[,i] <- as.numeric(gsub("-", "", uds.7[,i]))
}

```

## Read in my financial data
## Prep download
```{r}

cw <- read_excel("/Users/jhm65/Research/CHC Costs/UDS-990 Crosswalk.xlsx")

## Wave 1

EINs <- cw[!is.na(cw$EIN),]$EIN

urls <- paste("https://www.guidestar.org/profile/GetFTASheet?ein=", cw[!is.na(cw$EIN),]$EIN, "&guid=", sep = "")
destfiles <- paste("/Users/jhm65/Research/CHC Costs/990s/", cw[!is.na(cw$EIN),]$EIN, ".xlsx", sep = "") #990s_new or 990s

## Set Binary filter for data to include
z <- rep(NA, length(urls))
for (k in 1:length(EINs)){  #1174

 z[k] <- ifelse(is.data.frame(try(read_excel(destfiles[k], skip = 5) )) == TRUE, 1, 0)

}

## Set first rows in data
chc.990 <- read_excel(destfiles[1], skip = 5)
years <- as.numeric(colnames(chc.990)[-1])
chc.990 <- chc.990[-c(5, 6, 15, 16, 25, 26, 32, 33, 34, 35, 39, 40, 52:58),]

chc.990[,2] <- as.numeric(gsub(unlist(chc.990[,2]), pattern = "N/A", replacement = NA))
chc.990[,3] <- as.numeric(gsub(unlist(chc.990[,3]), pattern = "N/A", replacement = NA))
chc.990[,4] <- as.numeric(gsub(unlist(chc.990[,4]), pattern = "N/A", replacement = NA))
chc.990[,5] <- as.numeric(gsub(unlist(chc.990[,5]), pattern = "N/A", replacement = NA))
chc.990[,6] <- as.numeric(gsub(unlist(chc.990[,6]), pattern = "N/A", replacement = NA))

headers <- unlist(chc.990[,1])
chc.990 <- data.frame(t(chc.990[,-1]))

colnames(chc.990) <- headers
rownames(chc.990) <- NULL

chc.finances <- cbind(ID = EINs[1], year = years, chc.990)

## Repeat for the whole set
for (k in 2:1174){
  if(z[k] == 1 & k %notin% c(388, 511, 670)){ ##remove entries for now, check later if this clinic gets data
chc.990 <- read_excel(destfiles[k], skip = 5)
years <- as.numeric(colnames(chc.990)[-1])
chc.990 <- chc.990[-c(5, 6, 15, 16, 25, 26, 32, 33, 34, 35, 39, 40, 52:58),]

  for (j in 2:ncol(chc.990)){
    chc.990[,j] <- as.numeric(gsub(unlist(chc.990[,j]), pattern = "N/A", replacement = NA))
  }

headers <- unlist(chc.990[,1])
chc.990 <- data.frame(t(chc.990[,-1]))

colnames(chc.990) <- headers
rownames(chc.990) <- NULL

chc.finances <- rbind(chc.finances, cbind(ID = EINs[k], year = years, chc.990) )} else {print(k)}
}

table(chc.finances$year)

quantile(chc.finances$`Unrestricted surplus (deficit) after depreciation`, probs = c(0, 0.025, 0.1, 0.25, 0.5, 0.75, 0.9, 0.975, 1))

colnames(chc.finances) <- gsub("\\%", "pct", colnames(chc.finances))
colnames(chc.finances)[c(4, 6)] <- c("Surplus as a pct of before dep expenses","Surplus as a pct of after dep expenses")

## Wave 2

EINs <- cw[!is.na(cw$EIN),]$EIN

urls <- paste("https://www.guidestar.org/profile/GetFTASheet?ein=", cw[!is.na(cw$EIN),]$EIN, "&guid=", sep = "")
destfiles <- paste("/Users/jhm65/Research/CHC Costs/990s_new/", cw[!is.na(cw$EIN),]$EIN, ".xlsx", sep = "") #990s_new or 990s

## Set Binary filter for data to include
z <- rep(NA, length(urls))
for (k in 1:length(EINs)){  #1174

 z[k] <- ifelse(is.data.frame(try(read_excel(destfiles[k], skip = 5) )) == TRUE, 1, 0)

}

## Set first rows in data
chc.990 <- read_excel(destfiles[1], skip = 5)
years <- as.numeric(colnames(chc.990)[-1])
chc.990 <- chc.990[-c(5, 6, 15, 16, 25, 26, 32, 33, 34, 35, 39, 40, 52:58),]

chc.990[,2] <- as.numeric(gsub(unlist(chc.990[,2]), pattern = "N/A", replacement = NA))
chc.990[,3] <- as.numeric(gsub(unlist(chc.990[,3]), pattern = "N/A", replacement = NA))
chc.990[,4] <- as.numeric(gsub(unlist(chc.990[,4]), pattern = "N/A", replacement = NA))
chc.990[,5] <- as.numeric(gsub(unlist(chc.990[,5]), pattern = "N/A", replacement = NA))
chc.990[,6] <- as.numeric(gsub(unlist(chc.990[,6]), pattern = "N/A", replacement = NA))

headers <- unlist(chc.990[,1])
chc.990 <- data.frame(t(chc.990[,-1]))

colnames(chc.990) <- headers
rownames(chc.990) <- NULL

chc.fin.2 <- cbind(ID = EINs[1], year = years, chc.990)

## Repeat for the whole set
for (k in 2:1174){
  if(z[k] == 1 & k %notin% c(388, 511, 670)){ ##remove entries for now, check later if this clinic gets data
chc.990 <- read_excel(destfiles[k], skip = 5)
years <- as.numeric(colnames(chc.990)[-1])
chc.990 <- chc.990[-c(5, 6, 15, 16, 25, 26, 32, 33, 34, 35, 39, 40, 52:58),]

  for (j in 2:ncol(chc.990)){
    chc.990[,j] <- as.numeric(gsub(unlist(chc.990[,j]), pattern = "N/A", replacement = NA))
  }

headers <- unlist(chc.990[,1])
chc.990 <- data.frame(t(chc.990[,-1]))

colnames(chc.990) <- headers
rownames(chc.990) <- NULL

chc.fin.2 <- rbind(chc.fin.2, cbind(ID = EINs[k], year = years, chc.990) )} else {print(k)}
}

table(chc.fin.2$year)

quantile(chc.fin.2$`Unrestricted surplus (deficit) after depreciation`, probs = c(0, 0.025, 0.1, 0.25, 0.5, 0.75, 0.9, 0.975, 1))

colnames(chc.fin.2) <- gsub("\\%", "pct", colnames(chc.fin.2))
colnames(chc.fin.2)[c(4, 6)] <- c("Surplus as a pct of before dep expenses","Surplus as a pct of after dep expenses")

nrow(chc.finances[chc.finances$`Unrestricted surplus (deficit) after depreciation` <= 250000 & chc.finances$year == 2017,]) / nrow(chc.finances[chc.finances$year == 2017,])

```

## Read in my termination crosswalk to filter out closures
```{r}
chc.term <- read.csv("//Users/jhm65/Dropbox/Git/Competition/data_raw/CHC_termination_crosswalk.csv")
staffing <- read.csv("//Users/jhm65/Dropbox/Git/Data/FQHC_Staffing.csv")
```

## Read in census data for all zip codes
```{r}

census <- data.frame(ZipCode = 0, NAME = 0, YEAR = 0, 
                     TotalPop = 0,
                     Male = 0, Female = 0, 
                     AgeUnder18 = 0, Age18to64 = 0, Age65andOver = 0, 
                     White = 0, Black = 0, AIAN = 0, Asian = 0, MR = 0, Hispanic = 0, EducLow = 0, EducMid = 0, EducHigh = 0, 
                     PovertyStatusDet = 0, IncomeBelowPoverty = 0, Unemployed = 0, Unemp_denom = 0, 
                     MCD.denom = 0, Medicaid = 0, Unins.denom = 0, Uninsured = 0)

for (i in 1:7){
  census <- rbind(census, get_acs(geography = "zcta", survey = "acs5",
                       variables = c("B01003_001","B17001_001","B17001_002","B01001_002","B01001_026",
                                     "B02001_002","B02001_003","B02001_004","B02001_005","B02001_006","B02001_008",
                                     "B03001_003", 
                                     "B01001_003","B01001_004","B01001_005","B01001_006","B01001_007","B01001_008","B01001_009",
                                     "B01001_010","B01001_011","B01001_013","B01001_014","B01001_015","B01001_016","B01001_017",
                                     "B01001_018","B01001_019","B01001_020","B01001_021","B01001_022","B01001_023","B01001_024","B01001_025",
                                     "B01001_027","B01001_028","B01001_029","B01001_030","B01001_031","B01001_032","B01001_033","B01001_034",
                                     "B01001_035","B01001_036","B01001_037","B01001_038","B01001_039","B01001_040","B01001_041","B01001_042","B01001_043",
                                     "B01001_044","B01001_045","B01001_046","B01001_047","B01001_048","B01001_049",
                                     "B06009_002","B06009_003", "B06009_004","B06009_005","B06009_006",
                                     "B23025_002","B23025_005",
                                     "C27007_001", "C27007_004", "C27007_007", "C27007_010", "C27007_014", "C27007_017", "C27007_020",
                                     "B18135_001","B18135_007","B18135_012","B18135_018","B18135_023","B18135_029","B18135_034"), 
                       year = 2013 + i) %>% 
  pivot_wider(id_cols = c(GEOID, NAME), names_from = variable, values_from = estimate) %>% 
    mutate(YEAR = 2013 + i, TotalPop = B01003_001, PovertyStatusDet = B17001_001, IncomeBelowPoverty = B17001_002,
           AgeUnder18 = B01001_003 + B01001_004 + B01001_005 + B01001_006 + B01001_027 + B01001_028 + B01001_029 + B01001_030,
           Age18to64 = B01001_007 + B01001_008 + B01001_009 + B01001_010 + B01001_011 + B01001_013 + B01001_014 + B01001_015 + 
                       B01001_016 + B01001_017 + B01001_018 + B01001_019 + B01001_031 + B01001_032 + B01001_033 + B01001_034 +
                       B01001_035 + B01001_036 + B01001_037 + B01001_038 + B01001_039 + B01001_040 + B01001_041 + B01001_042 + B01001_043, 
           Age65andOver = B01001_020 + B01001_021 + B01001_022 + B01001_023 + B01001_024 + B01001_025 + 
                          B01001_044 + B01001_045 + B01001_046 + B01001_047 + B01001_048 + B01001_049, 
           Male = B01001_002, Female = B01001_026, 
           White = B02001_002, Black = B02001_003, AIAN = B02001_004, Asian = B02001_005 + B02001_006, MR = B02001_008, 
           Hispanic = B03001_003, 
           EducLow = B06009_002, EducMid = B06009_003 + B06009_004, EducHigh = B06009_005 + B06009_006,
           Unemployed = B23025_005, Unemp_denom = B23025_002,
           MCD.denom = C27007_001	, Medicaid = C27007_004	 + C27007_007 + C27007_010	+ C27007_014 + C27007_017 + C27007_020,
           Unins.denom = B18135_001, Uninsured = B18135_007 + B18135_012 + B18135_018 + B18135_023 + B18135_029 + B18135_034) %>% 
    select(ZipCode = GEOID, NAME, YEAR, TotalPop:Uninsured))

  print(i)
}
census <- census[-1,]

##########
census$ZipCode <- substr(census$NAME, 7, 11)
#census$ZipCode <- str_pad(census$ZipCode, 5, pad = "0")

census %>% 
  filter(YEAR == 2020) %>% 
  select(ZipCode) %>% 
  #unique() %>% 
  table() %>% 
  table()
```

# Analysis
## SEt data again

```{r}

uds.sites[!is.na(uds.sites$Zip),] %>% 
  group_by(`Grant Number`, Zip, Year) %>% 
  summarise(Counts = n()) %>% 
  select(`Grant Number`, Zip, Year, Counts) %>% 
  unique() %>% 
  arrange(Year) %>% 
  pivot_wider(id_cols = c(`Grant Number`, Zip), names_from = Year, values_from = Counts, values_fill = 0) %>% 
  arrange(Zip)

```

## Set Data
```{r}
# Identify number of clinics in each zip code-year
clinics.in.zips <- uds.sites[!is.na(uds.sites$Zip),] %>%
  group_by(Zip, Year) %>%
  summarise(Clinics = n_distinct(`Grant Number`), Sites = n()) %>%
  arrange(desc(Clinics))

# Identify clinics in competitive zip codes
clinics.comp <- uds.sites[!is.na(uds.sites$Zip),] %>% 
  select(`Grant Number`, Zip, Year) %>% 
  unique() %>% 
  left_join(clinics.in.zips, by = c("Zip", "Year")) %>% 
  group_by(`Grant Number`, Year) %>% 
  summarise(Clinics = max(Clinics)) %>% 
  pivot_wider(id_cols = `Grant Number`, names_from = Year, values_from = Clinics) %>% 
  filter(`2014` == 1 & `2015` == 1 & `2016` == 1 ) %>%   
  #filter(`2014` == 1 & `2015` == 1) %>%  
  left_join(chc.term, by = c("Grant Number" = "ClinicID")) %>%
  filter(is.na(NewClinicID)) %>%
  select(`Grant Number`:`2021`) %>% 
  pivot_longer(cols = c(`2014`:`2021`)) %>% 
  group_by(`Grant Number`) %>% 
  filter(min(value) == 1 ) %>% 
  #filter(max(value) == last(value)) %>% 
  mutate(Tx = ifelse(value > 1, 1, 0))

# Classify clinics by introduction of competition
df.comp <- left_join(clinics.comp %>% mutate(Year = as.numeric(name)), uds.zip.totals, by = c("Grant Number", "Year" = "Reporting Year")) %>% 
  left_join(uds.sites[!is.na(uds.sites$Zip),] %>% 
  group_by(`Grant Number`, Year) %>% 
  summarise(Sites = n()), by = c("Grant Number", "Year")) %>% 
  ungroup() %>% 
  group_by(`Grant Number`) %>% 
  mutate(Att = min(ifelse(Tx == 0, 2, 1)*Year, na.rm = TRUE) ) %>% 
  group_by(`Grant Number`) %>% 
  mutate(Att = ifelse(Att < 2022, Att, 0)) %>% 
  ungroup() %>% 
  mutate(ID = as.numeric(factor(`Grant Number`)))

# Identify expandors vs expandees
df.expand <- df.comp %>% 
  left_join(
      left_join(df.comp %>% select(`Grant Number`, Att), uds.sites[!is.na(uds.sites$Zip),] %>% 
        select(`Grant Number`, Zip, Year) %>% 
        group_by(`Grant Number`, Zip) %>% 
        summarise(YearFirstInZip = min(Year)) %>% 
        filter(YearFirstInZip %notin% c(2014)) %>% 
        select(`Grant Number`, YearFirstInZip) %>% 
        unique(), by = c("Grant Number")) %>% #filter(`Grant Number` == "H80CS00101") %>% 
        mutate(Expandor = ifelse(is.na(YearFirstInZip) | Att != YearFirstInZip, 0, 1)) %>%  #  is.na(YearFirstInZip) | Att != YearFirstInZip
        group_by(`Grant Number`) %>% summarise(Expandor = max(Expandor)), by = c("Grant Number")) %>% 
  left_join(uds.HIT, by = c("Grant Number", "Year"))

df.expand <- df.expand %>% 
  left_join(df.expand %>% filter(SwitchEMR == 1) %>%  
            select(`Grant Number`, Year, SwitchEMR) %>% 
            group_by(`Grant Number`) %>% 
            summarise(FirstSwitch = min(Year)), by = c("Grant Number")) #%>% mutate(SwitchEMR = ifelse(is.na(FirstSwitch), 0, ifelse(Year >= FirstSwitch & FirstSwitch >= 2014, 1, 0)))

df.qual.exp <- df.expand %>% 
  left_join(uds.6b.rates, by = c("Grant Number" = "GrantNumber", "Year")) %>% 
  left_join(uds.6b.denom, by = c("Grant Number" = "GrantNumber", "Year")) %>% 
  #filter(Expandor == 0 & Year %in% c(2015, 2016, 2017, 2018, 2019, 2020) & Att %in% c(0, 2017, 2018, 2019, 2020)) %>%  ## Regular go to analysis
  filter(Expandor == 0 & Year %in% c(2015, 2016, 2017, 2018, 2019) & Att %in% c(0, 2017, 2018, 2019)) %>%  ## Really pronounced causal responses, weak first order stuff
  #filter(Expandor == 0 & Year %in% c(2014, 2015, 2016, 2017, 2018, 2019, 2020, 2021) & Att %in% c(0, 2017, 2018, 2019)) %>%
  left_join(uds.sites %>% 
    group_by(`Grant Number`, Year) %>% 
    summarise(TotalAccess = sum(Access, na.rm = T)), by = c("Grant Number","Year")) %>%
inner_join(uds.sites %>% filter(Year == 2015) %>% select(`Grant Number`, Zip) %>% group_by(`Grant Number`) %>% summarise(Zips = n_distinct(Zip)), by = c("Grant Number")) %>%
  mutate(ZipFilter = ifelse(Zips > 2.75, 2, 1)) %>% 
#inner_join(uds.sites %>% filter(Year == 2014) %>% select(`Grant Number`, Zip) %>% group_by(`Grant Number`) %>% summarise(Zips = n_distinct(Zip)) %>% filter(Zips > 2.75), by = c("Grant Number")) %>%
#inner_join(df.expand %>% filter(Year == 2015) %>% mutate(ZipFilter = ifelse(Total > median(Total), 2, 1)) %>% select(`Grant Number`, ZipFilter), by = c("Grant Number")) %>%
  mutate(l.Sites = log(Sites),
         l.Total = log(Total))

k = 1  # 1 for larger group, 0 for smaller clinic group OOOOOOORRRRRRR 2 for large, 1 for medium, and 0 for small

# df.qual.exp <- inner_join(uds.sites %>% select(`Grant Number`, `Site State`) %>% unique(), df.qual.exp, by = c("Grant Number")) %>%
# mutate(ACA = ifelse((`Site State` %in% c('MI','NH') & Year >= 2014) |
#                     (`Site State` %in% c('PA','IN','AK') & Year >= 2015) |
#                     (`Site State` %in% c('MT','LA') & Year >= 2016) |
#                     (`Site State` %in% c('VA','ME') & Year >= 2019) |
#                     (`Site State` %in% c('ID','UT','NE') & Year >= 2020) |
#                     (`Site State` %in% c('OK','MO') & Year >= 2021) |
#                     (`Site State` %in% c('AZ','AR','CA','CO','CT','DE','DC','HI','IL','IA','KY','MD','MA','MN','NV','NJ','NM','NY','ND','OH','OR','PA','RI','VT','VA','WA') & Year >= 2014), 1, 0)) %>%
#   select(`Grant Number`, ZipFilter, value:ACA)

# %>%
#   group_by(`Grant Number`) %>%
#   filter(max(ACA) == 0 | min(ACA) == 1) %>%
#   ungroup()

df.qual.exp <- df.qual.exp %>% 
  mutate(
    None = None / Total * 100,
         MCR = MCR / Total * 100,
         COM = COM / Total * 100,
         MCD = MCD / Total * 100,

         SwitchEMR = SwitchEMR * 100, 
         eScripts = eScripts* 100, 
         eDecisionSupport = eDecisionSupport * 100, 
         InterOrgHITexchange = InterOrgHITexchange * 100, 
         PatientPortals = PatientPortals * 100, 
         UDSQualRep.highlowuse = UDSQualRep.highlowuse * 100) %>% 
  left_join(uds.covariates, by = c("Grant Number","Year")) %>% 
  left_join(uds.7, by = c("Grant Number","Year")) %>% 
  mutate(Pct.HBW = BirthsGTE2500 / TotalDeliveries * 100, 
         Pct.HTN.CNTRL = Controlcount / Sample * 100) %>%

  left_join(cw %>% select(`Grant Number`, EIN), by = c("Grant Number")) %>% 
  left_join(unique(rbind(chc.finances, chc.fin.2) %>% group_by(ID, year) %>% summarise_all(max)) %>% 
  arrange(ID, year), by = c("EIN" = "ID", "Year" = "year")) %>% 
  mutate(Liabilities = `Liabilities (as pct of assets)` * `Total net assets`, 
         Staffing = `Total expenses before depreciation` * Personnel, 
         LBE.pct = `Gross land, buildings and equipment (LBE)` / `Total net assets`, 
         `Liabilities (as pct of assets)` = `Liabilities (as pct of assets)` * 100, 
         `Surplus as a pct of before dep expenses` = `Surplus as a pct of before dep expenses` * 100, 
         Qcomp = (IMM2yo + Pap + ChildBMI + AdultBMI + Tobac + CRC + DepScrn) / 7, 
         CostOfCare = `Total expenses before depreciation` / Total) %>%

  mutate(Liquidity =  `Total net assets` / (Cash + Investments + Receivables + `Gross land, buildings and equipment (LBE)`), 
         Profitability =  `Unrestricted surplus (deficit) before depreciation` / (Cash + Investments + Receivables + `Gross land, buildings and equipment (LBE)`), 
         Efficiency = (`Total revenue (unrestricted & restricted)` - `Total expenses before depreciation`) / (Cash + Investments + Receivables + `Gross land, buildings and equipment (LBE)`), 
         NetWorth = (100 - `Liabilities (as pct of assets)`) / `Liabilities (as pct of assets)` / 100) %>% 
  mutate(Altmans.Z = 6.56 * Liquidity + 3.26 * Profitability + 6.72 * Efficiency + 1.05 * NetWorth) %>% # 6.56, 3.26, 6.72, 1.05  OR 0.18, 0.3, 0.81, 0.14
  mutate(Liquidity = ifelse(Liquidity == 'Inf', NA, Liquidity), 
         Profitability = ifelse(Profitability == 'Inf', NA, Profitability), 
         Efficiency = ifelse(Efficiency == 'Inf', NA, Efficiency), 
         NetWorth = ifelse(NetWorth == 'Inf', NA, NetWorth), 
         Altmans.Z = ifelse(Altmans.Z == 'Inf', NA, Altmans.Z)) #%>%   filter(!is.na(Altmans.Z) & Altmans.Z != 'Inf')

df.qual.exp %>%
  group_by(Att, Expandor) %>%
  summarise(n() / 6)

df.qual.exp %>% 
  group_by(Tx) %>% 
  summarise(mean(Altmans.Z, na.rm = TRUE), sd(Altmans.Z, na.rm = TRUE))

##########
## Using penetration rate to filter instead of zip codes
##########
#IncomeBelowPoverty

years <- unique(df.qual.exp$Year)

#FQHC penetration rate by zip code
temp1 <- uds.zip %>%
  filter(`Reporting Year` == 2015 ) %>%
  inner_join(uds.sites[!is.na(uds.sites$Zip),] %>%
              filter(Year %in% years), by = c("Grant Number","Zip Code" = "Zip", "Reporting Year" = "Year")) %>%
  select(`Grant Number`, `Reporting Year`, `Zip Code`, `Total Number of Patients`) %>%
  left_join(census, by = c("Zip Code" = "ZipCode","Reporting Year" = "YEAR")) %>% #,"Reporting Year" = "YEAR"
  mutate(MarketShare = `Total Number of Patients` / IncomeBelowPoverty) %>%#IncomeBelowPoverty or TotalPop
  unique()

temp1 %>% 
  filter(`Reporting Year` == 2020)

census %>% filter(YEAR == 2020)
# FQHC weighted average penetration rate conditional on site in zip

temp2 <- temp1 %>%
  mutate(Num = IncomeBelowPoverty * MarketShare) %>%
  group_by(`Grant Number`, `Reporting Year`) %>%
  summarise(Num = sum(Num, na.rm = TRUE)) %>%
  left_join(temp1 %>%
  group_by(`Grant Number`, `Reporting Year`) %>%
  summarise(YTotal = sum(IncomeBelowPoverty, na.rm = TRUE)), by = c("Grant Number","Reporting Year")) %>% #IncomeBelowPoverty
  mutate(MarketShare = Num / YTotal) %>%
  select(`Grant Number`, `Reporting Year`, MarketShare)

df.qual.exp <- df.qual.exp %>%
  select(`Grant Number`, value:Altmans.Z) %>% 
  left_join(temp2, by = c("Grant Number")) %>% #,"Year" = "Reporting Year")) %>%
  #mutate(ZipFilter = ifelse(MarketShare > median(MarketShare, na.rm = FALSE), 2,  # - 0.25
  #                   ifelse(is.na(MarketShare) == T, 1, 1))) %>%
  mutate(ZipFilter = ifelse(MarketShare >= median(MarketShare, na.rm = TRUE), 2, 1)) %>% #median(MarketShare, na.rm = TRUE)
  select(`Grant Number`, ZipFilter, MarketShare, value:Altmans.Z) %>%
  # inner_join(df.qual.exp %>% filter(Year == 2016) %>% mutate(Filter = ifelse(`Surplus as a pct of after dep expenses` < 0, 1, 
  #                                                                     ifelse(`Surplus as a pct of after dep expenses` > 0.03, 2, NA))) %>% select(`Grant Number`, Filter), by = c("Grant Number")) %>% 
  # select(`Grant Number`, ZipFilter, Filter, MarketShare, value:Altmans.Z) %>% 
  #mutate(ZipFilter = Filter) %>% 
  unique()

summary(df.qual.exp$MarketShare)
summary(df.qual.exp$Zips)
summary(df.qual.exp$`Surplus as a pct of after dep expenses`)

df.qual.exp %>% mutate(Treat = ifelse(Att > 0, 'Treat','Control')) %>% group_by(ZipFilter, Treat) %>% summarise(n(), n_distinct(ID))
df.qual.exp %>% filter(`Grant Number` == "H80CS00225")

cnames <- colnames(df.qual.exp)

compare <- read.csv("comeptition.data.final.csv")[,-1]
colnames(compare) <- cnames

# df.qual.exp <- compare
# df.qual.exp <- df.qual.exp %>%
#   #select(`Grant Number`:Year) %>%
#   inner_join(compare %>% select(`Grant Number`, Year, value), by = c("Grant Number","Year")) %>%
#   filter(value.x == value.y) %>%
#   select(`Grant Number`:Altmans.Z) %>%
#   rename(value = value.x)

df.qual.exp <- df.qual.exp %>% rbind(compare %>% filter(`Grant Number` %in% c("H80CS00765","H80CS00881","H80CS02323") & Year %in% unique(df.qual.exp$Year) )) %>% 
  filter(`Grant Number` != "H80CS00225")

uds.sites[!is.na(uds.sites$Zip),] %>% 
  #select(`Grant Number`, Zip, Year) %>% 
  unique() %>% filter(`Grant Number` %in% c("H80CS00881") & Year == 2016)

# df.qual.exp %>% 
#   select(`Grant Number`, Year, value) %>% 
#   full_join(compare %>% select(`Grant Number`, Year, value), by = c("Grant Number","Year")) %>% 
#   filter(is.na(value.x) | is.na(value.y))

#write.csv(df.qual.exp, file = "comeptition.data.final_2.csv") DO NOT UNCOMMENT THIS LINE AS IT WILL OVERWRITE MY BACKUP DATA

#df.qual.exp <- read.csv("comeptition.data.final_2.csv")

## Add in urbanicity
all.density <- uds.sites %>% 
  filter(Year == 2016) %>% 
  select(`Grant Number`, Zip) %>% 
  unique()

zipsize2 <- zctas(cb = TRUE, year = 2010)

density.calc <- all.density %>% 
  left_join(census %>% filter(YEAR == 2016), by = c("Zip" = "ZipCode")) %>% 
  left_join(zipsize2 %>% select(ZCTA5, CENSUSAREA), by = c("Zip" = "ZCTA5")) %>% 
  group_by(`Grant Number`) %>% 
  mutate(num = sum(TotalPop, na.rm = TRUE), denom = sum(CENSUSAREA, na.rm = TRUE)) %>% 
  mutate(Density = num / denom) %>% 
  mutate(Pct.Rural = ifelse(Density < 500, 1, 0), 
         Pct.Urban = ifelse(Density > 2500, 1, 0)) %>% 
  select(`Grant Number`, Density, Pct.Rural, Pct.Urban) %>% 
  unique()

df.qual.exp <- df.qual.exp %>% 
  left_join(density.calc, by = c("Grant Number")) #%>% mutate(ZipFilter = ifelse(Density >= median(Density), 2, 1))

df.expand %>% 
  filter(`Grant Number` == "H80CS00101")

uds.sites %>% filter(`Grant Number` == "H80CS00101") %>% 
  group_by(Zip) %>% 
  summarise(min(Year))
```