### 2023-08-15 code/data_cleaning_longer.Rmd
mid session updates

---
title: "data_cleaning_longer_period"
author: "Justin"
date: "2023-08-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r}
library(tigris)
library(tidygeocoder)

library(tidyverse)
library(readxl)
library(did)
library(Matching)
library(gtsummary)
library(RODBC)
library(ggpubr)
library(fixest)
library(kableExtra)
library(usmap)
library(did2s)
library(formattable)
library (readr)

`%notin%` <- Negate(`%in%`)
select <- dplyr::select

```

## Formating plots

```{r, echo = FALSE, results = "asis", message=FALSE, warning = FALSE, fig.align="center"}

is.extrafont.installed <- function(){
  if(is.element("extrafont", installed.packages()[,1])){
    library(extrafont)
    # probably need something here to run font_import()
    return(T)
  }else{
    warning("Library extrafont installed; using system sans/serif libraries as fallback fonts. 
    To enable full font support, run: 
      install.packages('extrafont') 
      font_import()")
    return(F)
  }
}

# set font
base_font_family_tufte <- function(){
  if(is.extrafont.installed()){
    library(extrafont)
    tuftefont <- choose_font(c("Gill Sans MT", "Gill Sans", "GillSans", "Verdana", "serif"), quiet = FALSE)   #
  }else{
    tuftefont <- "serif"
  }
  return(tuftefont)
}

theme_tufte_revised <- function(base_size = 11, base_family = base_font_family_tufte(), ticks = TRUE) {

  ret <- theme_bw(base_family = base_family, base_size = base_size) + 
    theme(
      axis.line = element_line(color = 'black'),
      axis.title.x = element_text(vjust = -0.3), 
      axis.title.y = element_text(vjust = 0.8),
      legend.background = element_blank(), 
      legend.key = element_blank(), 
      legend.title = element_text(face="plain"),
      panel.background = element_blank(), 
      panel.border = element_blank(),
      panel.grid = element_blank(),
      plot.background = element_blank(),
      strip.background = element_blank()
    )

  if (!ticks) {
    ret <- ret + theme(axis.ticks = element_blank())
  }

  ret
}

```

# Data Prep

## Sites and ZipCodes
```{r}

## Zips
locs <- paste("/Users/jhm65/Dropbox/Git/UDS/UDS-", c(2014:2021), "-Full-Dataset.xlsx", sep = "")
uds.zip <- read_excel(locs[1], sheet = "HealthCenterZipCodes")

new.zip.names <- c("BHCMIS ID", "Grant Number", "Reporting Year" , "Zip Code", "Zip Code Type",  "None_Uninsured Patients" ,  "Medicaid_CHIP_Other Public Patients", "Medicare Patients", "Private Patients", "Total Number of Patients" ) 
colnames(uds.zip) <-    new.zip.names

for(i in 2:length(locs)){
  temp.zip <- read_excel(locs[i], sheet = "HealthCenterZipCodes")
  colnames(temp.zip) <- new.zip.names
  uds.zip <- rbind(uds.zip, temp.zip %>% mutate(`Zip Code` = str_pad(`Zip Code`, 5, pad = "0")))
}

# ## Rename
# new.zip.names <- c("BHCMIS ID", "Grant Number", "Reporting Year" , "Zip Code", "Zip Code Type",  "None_Uninsured Patients" ,  "Medicaid_CHIP_Other Public Patients", "Medicare Patients", "Private Patients", "Total Number of Patients" ) 
# colnames(uds.zip) <-    new.zip.names
# 
# 
# ## Clean
# 
# uds.zip[uds.zip$`Total Number of Patients`== '-' | uds.zip$`Total Number of Patients`== '--',]$`Total Number of Patients` <- NA
# uds.zip$`Total Number of Patients` <- as.numeric(uds.zip$`Total Number of Patients`)
# uds.zip <- uds.zip[!is.na(uds.zip$`Total Number of Patients`),] %>% 
#     mutate(`Reporting Year` = as.numeric(`Reporting Year`))
# 
# uds.zip$`None_Uninsured Patients` <- as.numeric(uds.zip$`None_Uninsured Patients`)
# uds.zip$`Medicaid_CHIP_Other Public Patients` <- as.numeric(uds.zip$`Medicaid_CHIP_Other Public Patients`)
# uds.zip$`Medicare Patients` <- as.numeric(uds.zip$`Medicare Patients`)
# uds.zip$`Private Patients` <- as.numeric(uds.zip$`Private Patients`)
# 
# 
# uds.zip <- cbind(uds.zip, broken.total = rowSums(uds.zip[,c(6:9)], na.rm = TRUE), missings = rowSums(is.na(uds.zip[,c(6:9)]))) %>% 
#   mutate(`Total Number of Patients` = as.numeric(`Total Number of Patients`)) %>% 
#   mutate(`Total Number of Patients` = ifelse(is.na(`Total Number of Patients`), round(missings*16/4), `Total Number of Patients`)) %>% 
#   mutate(`None_Uninsured Patients` = ifelse(is.na(`None_Uninsured Patients`), (`Total Number of Patients` - broken.total) * 1/missings, `None_Uninsured Patients`), 
#          `Medicaid_CHIP_Other Public Patients` = ifelse(is.na(`Medicaid_CHIP_Other Public Patients`), (`Total Number of Patients` - broken.total) * 1/missings, `Medicaid_CHIP_Other Public Patients`), 
#          `Medicare Patients` = ifelse(is.na(`Medicare Patients`), (`Total Number of Patients` - broken.total) * 1/missings, `Medicare Patients`), 
#          `Private Patients` = ifelse(is.na(`Private Patients`), (`Total Number of Patients` - broken.total) * 1/missings, `Private Patients`))

uds.zip$`Reporting Year` <- as.numeric(uds.zip$`Reporting Year`)
uds.zip$`Total Number of Patients` <- as.numeric(uds.zip$`Total Number of Patients`)
uds.zip$`None_Uninsured Patients` <- as.numeric(uds.zip$`None_Uninsured Patients`)
uds.zip$`Medicaid_CHIP_Other Public Patients` <- as.numeric(uds.zip$`Medicaid_CHIP_Other Public Patients`)
uds.zip$`Medicare Patients` <- as.numeric(uds.zip$`Medicare Patients`)
uds.zip$`Private Patients` <- as.numeric(uds.zip$`Private Patients`)

uds.zip[is.na(uds.zip$`Total Number of Patients`),]$`Total Number of Patients` <- 0
uds.zip[grep("-", uds.zip$`None_Uninsured Patients`),]$`None_Uninsured Patients` <- 0
uds.zip[grep("-", uds.zip$`Medicaid_CHIP_Other Public Patients`),]$`Medicaid_CHIP_Other Public Patients` <- 0
uds.zip[grep("-", uds.zip$`Medicare Patients`),]$`Medicare Patients` <- 0
uds.zip[grep("-", uds.zip$`Private Patients`),]$`Private Patients` <- 0

uds.zip.totals <- uds.zip %>%
  select(`Grant Number`, `Reporting Year`, `None_Uninsured Patients`:`Total Number of Patients`) %>%
  group_by(`Grant Number`, `Reporting Year`) %>% summarise(None = sum(`None_Uninsured Patients`, na.rm = T),
                                                           MCR = sum(`Medicare Patients`, na.rm = T),
                                                           COM = sum(`Private Patients`, na.rm = T),
                                                           MCD = sum(`Medicaid_CHIP_Other Public Patients`, na.rm = T),
                                                           Total = sum(`Total Number of Patients`, na.rm = T))

uds.zip <- uds.zip[uds.zip$`Zip Code Type` == "ZipCode",]

## Sites
locs <- paste("/Users/jhm65/Dropbox/Git/UDS/UDS-", c(2012:2021), "-Full-Dataset.xlsx", sep = "")

uds.sites <- read_excel(locs[1], sheet = "HealthCenterSiteInfo") %>% 
  select(BHCMISID, GrantNumber, HealthCenterName, SiteName, SiteType, LocationType, LocationSetting, OperationalSchedule, CalendarSchedule, TotalWeeklyHoursOfOperation, SiteStreetAddress, SiteCity, SiteState, SiteZIPCode) %>% 
  mutate(Year = 2012)

colnames(uds.sites) <- c("BHCMIS ID", "Grant Number", "Health Center Name", "Site Name", "Site Type", "Location Type", "Location Setting", "Operational Schedule", "Calendar Schedule", "Total Weekly Hours Of Operation", "Site Street Address", "Site City", "Site State", "Site ZIP Code", "Year")

for(i in 2:3){
site.temp <- read_excel(locs[i], sheet = "HealthCenterSiteInfo") %>% 
                       select(BHCMISID, GrantNumber, HealthCenterName, SiteName, SiteType, LocationType, LocationSetting, OperationalSchedule, CalendarSchedule, TotalWeeklyHoursOfOperation, 
                              SiteStreetAddress, SiteCity, SiteState, SiteZIPCode) %>%
  mutate(Year = 2011 + i)

colnames(site.temp) <- c("BHCMIS ID", "Grant Number", "Health Center Name", "Site Name", "Site Type", "Location Type", "Location Setting", "Operational Schedule", "Calendar Schedule", "Total Weekly Hours Of Operation", "Site Street Address", "Site City", "Site State", "Site ZIP Code", "Year")

   uds.sites <-  rbind(uds.sites, site.temp)
}

for(i in 4:(length(locs)-1)){
  uds.sites <- rbind(uds.sites, read_excel(locs[i], sheet = "HealthCenterSiteInfo") %>% 
                       select("BHCMIS ID", "Grant Number", "Health Center Name", "Site Name", "Site Type", "Location Type", "Location Setting", "Operational Schedule", "Calendar Schedule", "Total Weekly Hours Of Operation", "Site Street Address", "Site City", "Site State", "Site ZIP Code") %>% 
  mutate(Year = 2011 + i))
}

temp.sites <- read_excel(locs[i+1], sheet = "HealthCenterSiteInfo" )%>% 
  select(BHCMISID, GrantNumber, HealthCenterName, SiteName, SiteType, LocationType, LocationSetting, OperationalSchedule, CalendarSchedule, TotalWeeklyHoursOfOperation, SiteStreetAddress, SiteCity, SiteState, SiteZIPCode) %>% 
  mutate(Year = 2021)
colnames(temp.sites) <- c("BHCMIS ID", "Grant Number", "Health Center Name", "Site Name", "Site Type", "Location Type", "Location Setting", "Operational Schedule", "Calendar Schedule", "Total Weekly Hours Of Operation", "Site Street Address", "Site City", "Site State", "Site ZIP Code", "Year")

uds.sites <- rbind(uds.sites, temp.sites)

uds.sites.old <- read.csv(file = "//Users/jhm65/Dropbox/Git/APMs/data_processed/Merged_Sites.csv", colClasses = "character") %>% 
  mutate(Year = as.numeric(Year))

colnames(uds.sites.old) <- c("BHCMIS ID", "Health Center Name", "Site Name", "Site Street Address", "Site Street Address 2", "Site State", "Site ZIP Code", "Year", "GranteeID", "Site Street Address 3","Site City","ModernGrateeID")

uds.sites.full <- bind_rows(uds.sites, uds.sites.old) %>% 
  arrange(`BHCMIS ID`, Year) %>% 
  select(`BHCMIS ID`:`Health Center Name`, Year, `Site Name`, `Site Street Address`, `Site Street Address 2`, `Site Street Address 3`, `Site City`, `Site State`, `Site ZIP Code` , `Site Type`:`Total Weekly Hours Of Operation`)

uds.sites.full$Zip <- substr(uds.sites.full$`Site ZIP Code`, 1, 5)

uds.sites.full <- uds.sites.full %>% 
  filter(`Site State` %notin% c('GU','VI','PW','MH','AS','FM','PR','MP')) #'AK','HI',

uds.sites.full[grep(pattern = "-", uds.sites.full$Zip),]$Zip <- NA

#write.csv(x = uds.sites.full, file = "//Users/jhm65/Dropbox/Git/APMs/data_processed/Sites96to21.csv")

```

## Merge with Older UDSs 1996-2011
```{r}
uds.cw <- read.csv(file = "//Users/jhm65/Dropbox/Git/APMs/data_processed/Merged_CW.csv", colClasses = "character") %>% 
  mutate(link.ID = ifelse(is.na(ModernGranteeID) & !is.na(LongGrantID) & Year != 2010, LongGrantID, 
                   ifelse(is.na(ModernGranteeID) & !is.na(GranteeID), GranteeID, BHCMIS.ID)), 
         Year = as.integer(Year))

uds.cw %>% 
  inner_join(read.csv(file = "//Users/jhm65/Dropbox/Git/APMs/data_processed/Merged_Table3B.csv"), by = c("link.ID" = "BHCMIS.ID", "Year")) %>% 
  arrange(BHCMIS.ID, Year) %>% 
  select(BHCMIS.ID:link.ID, T3B_L1_CA:T3B_L7_CC) %>% 
  filter(BHCMIS.ID == '010060' & Year %in% c(2007:2011)) %>% 
  select(Year, BHCMIS.ID, ModernGranteeID, T3B_L5A_CA, T3B_L6_CA, T3B_L1_CA, T3B_L8_CA, T3B_L4_CA, T3B_L12_CA)

namevec <- c("BHCMIS.ID", "Year", "NHA", "NHB", "HL", "NHW", "Total", "Eng2ndLang")

#fins.revenue.expenses <- read_csv("FQHC_Revenue_vs_Expenses.csv")
#cw <- read_excel("UDS-990 Crosswalk.xlsx")

```

## Clinic-level in Covariates 
```{r}

# Table 3B

uds.3B.2012 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2012-Full-Dataset.xlsx", sheet = "Table3B", skip = 0) %>% select(BHCMISID, GrantNumber, T3b_L1_Cb, T3b_L3_Cb, T3b_L8_Ca, T3b_L5_Cb, T3b_L8_Cd, T3b_L12_Ca) %>% rename(T3b_L26_Ca = T3b_L8_Cd)
uds.3B.2013 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2013-Full-Dataset.xlsx", sheet = "Table3B", skip = 0) %>% select(BHCMISID, GrantNumber, T3b_L1_Cb, T3b_L3_Cb, T3b_L8_Ca, T3b_L5_Cb, T3b_L8_Cd, T3b_L12_Ca) %>% rename(T3b_L26_Ca = T3b_L8_Cd)

uds.3B.2014 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2014-Full-Dataset.xlsx", sheet = "Table3B", skip = 0) %>% select(BHCMISID, GrantNumber, T3b_L1_Cb, T3b_L3_Cb, T3b_L8_Ca, T3b_L5_Cb, T3b_L8_Cd, T3b_L12_Ca) %>% rename(T3b_L26_Ca = T3b_L8_Cd)
uds.3B.2015 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2015-Full-Dataset.xlsx", sheet = "Table3B", skip = 2) %>% select(BHCMISID, GrantNumber, T3b_L1_Cb, T3b_L3_Cb, T3b_L8_Ca, T3b_L5_Cb, T3b_L8_Cd, T3b_L12_Ca) %>% rename(T3b_L26_Ca = T3b_L8_Cd)
uds.3B.2016 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2016-Full-Dataset.xlsx", sheet = "Table3B", skip = 1) %>% select(BHCMISID, GrantNumber, T3b_L1_Cb, T3b_L3_Cb, T3b_L8_Ca, T3b_L5_Cb, T3b_L26_Ca, T3b_L12_Ca)

uds.3B.2017 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2017-Full-Dataset.xlsx", sheet = "Table3B", skip = 1) %>% select(BHCMISID, GrantNumber, T3b_L1_Cb, T3b_L3_Cb, T3b_L8_Ca, T3b_L5_Cb, T3b_L26_Ca, T3b_L12_Ca)
uds.3B.2018 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2018-Full-Dataset.xlsx", sheet = "Table3B", skip = 1) %>% select(BHCMISID, GrantNumber, T3b_L1_Cb, T3b_L3_Cb, T3b_L8_Ca, T3b_L5_Cb, T3b_L26_Ca, T3b_L12_Ca)
uds.3B.2019 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2019-Full-Dataset.xlsx", sheet = "Table3B", skip = 1) %>% select(BHCMISID, GrantNumber, T3b_L1_Cb, T3b_L3_Cb, T3b_L8_Ca, T3b_L5_Cb, T3b_L26_Ca, T3b_L12_Ca)

uds.3B.2020 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2020-Full-Dataset.xlsx", sheet = "Table3B", skip = 1) %>% select(BHCMISID, GrantNumber, T3b_L1_Cb, T3b_L3_Cb, T3b_L8_Ca, T3b_L5_Cb, T3b_L26_Ca, T3b_L12_Ca)
uds.3B.2021 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2021-Full-Dataset.xlsx", sheet = "Table3B", skip = 0) %>% select(BHCMISID, GrantNumber, T3b_L1_Cb, T3b_L3_Cb, T3b_L8_Ca, T3b_L5_Cb, T3b_L26_Ca, T3b_L12_Ca)

namevec <- c("Year", "BHCMISID", "ModernGranteeID", "NHA", "NHB", "HL", "NHW", "Total", "Eng2ndLang")
uds.3B <- rbind(cbind(Year = 2012, uds.3B.2012),
                cbind(Year = 2013, uds.3B.2013),
                cbind(Year = 2014, uds.3B.2014),
                  cbind(Year = 2015, uds.3B.2015),
                  cbind(Year = 2016, uds.3B.2016),
                  cbind(Year = 2017, uds.3B.2017),

                  cbind(Year = 2018, uds.3B.2018),
                  cbind(Year = 2019, uds.3B.2019), 
                  cbind(Year = 2020, uds.3B.2020),
                  cbind(Year = 2021, uds.3B.2021)) %>%

  rename_all(vars(namevec), function(x) namevec)

uds.3B <- rbind(
  uds.cw %>% 
  inner_join(read.csv(file = "//Users/jhm65/Dropbox/Git/APMs/data_processed/Merged_Table3B.csv"), by = c("link.ID" = "BHCMIS.ID", "Year")) %>% 
  arrange(BHCMIS.ID, Year) %>% 
  select(BHCMIS.ID:link.ID, T3B_L1_CA:T3B_L7_CC) %>% 
  filter(Year %in% c(2007)) %>% 
  select(Year, BHCMIS.ID, ModernGranteeID, T3B_L5A_CA, T3B_L6_CA, T3B_L1_CA, T3B_L8_CA, T3B_L4_CA, T3B_L12_CA) %>% 
    setNames(namevec), 
  uds.cw %>% 
  inner_join(read.csv(file = "//Users/jhm65/Dropbox/Git/APMs/data_processed/Merged_Table3B.csv"), by = c("link.ID" = "BHCMIS.ID", "Year")) %>% 
  arrange(BHCMIS.ID, Year) %>% 
  select(BHCMIS.ID:link.ID, T3B_L1_CA:T3B_L7_CC) %>% 
  filter(Year %in% c(2008)) %>% 
  select(Year, BHCMIS.ID, ModernGranteeID, T3B_L5A_CA, T3B_L6_CA, T3B_L1_CA, T3B_L8_CA, T3B_L4_CA, T3B_L12_CA) %>% 
    setNames(namevec),
    uds.cw %>% 
  inner_join(read.csv(file = "//Users/jhm65/Dropbox/Git/APMs/data_processed/Merged_Table3B.csv"), by = c("link.ID" = "BHCMIS.ID", "Year")) %>% 
  arrange(BHCMIS.ID, Year) %>% 
  select(BHCMIS.ID:link.ID, T3B_L1_CA:T3B_L7_CC) %>% 
  filter(Year %in% c(2009)) %>% 
  select(Year, BHCMIS.ID, ModernGranteeID, T3B_L1_CB, T3B_L3_CB, T3B_L8_CA, T3B_L5_CB, T3B_L8_CD, T3B_L12_CA) %>% 
    setNames(namevec), 
  uds.cw %>% 
  inner_join(read.csv(file = "//Users/jhm65/Dropbox/Git/APMs/data_processed/Merged_Table3B.csv"), by = c("link.ID" = "BHCMIS.ID", "Year")) %>% 
  arrange(BHCMIS.ID, Year) %>% 
  select(BHCMIS.ID:link.ID, T3B_L1_CA:T3B_L7_CC) %>% 
  filter(Year %in% c(2010)) %>% 
  select(Year, BHCMIS.ID, ModernGranteeID, T3B_L1_CB, T3B_L3_CB, T3B_L8_CA, T3B_L5_CB, T3B_L8_CD, T3B_L12_CA) %>% 
    setNames(namevec), 
  uds.cw %>% 
  inner_join(read.csv(file = "//Users/jhm65/Dropbox/Git/APMs/data_processed/Merged_Table3B.csv"), by = c("link.ID" = "BHCMIS.ID", "Year")) %>% 
  arrange(BHCMIS.ID, Year) %>% 
  select(BHCMIS.ID:link.ID, T3B_L1_CA:T3B_L7_CC) %>% 
  filter(Year %in% c(2011)) %>% 
  select(Year, BHCMIS.ID, ModernGranteeID, T3B_L1_CB, T3B_L3_CB, T3B_L8_CA, T3B_L5_CB, T3B_L8_CD, T3B_L12_CA) %>% 
    setNames(namevec), 
  uds.3B)
uds.3B %>% 
  arrange(BHCMISID, Year)

# Table 4
uds.4.2012 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2012-Full-Dataset.xlsx", sheet = "Table4", skip = 0) %>% select(BHCMISID, GrantNumber, T4_L1_Ca, T4_L2_Ca, T4_L3_Ca, T4_L4_Ca, T4_L5_Ca, T4_L16_Ca, T4_L23_Ca, T4_L7_Ca, T4_L7_Cb, T4_L8_Ca, T4_L8_Cb, T4_L9_Ca, T4_L9_Cb, T4_L10_Ca, T4_L10_Cb, T4_L11_Ca, T4_L11_Cb) 
uds.4.2013 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2013-Full-Dataset.xlsx", sheet = "Table4", skip = 0) %>% select(BHCMISID, GrantNumber, T4_L1_Ca, T4_L2_Ca, T4_L3_Ca, T4_L4_Ca, T4_L5_Ca, T4_L16_Ca, T4_L23_Ca, T4_L7_Ca, T4_L7_Cb, T4_L8_Ca, T4_L8_Cb, T4_L9_Ca, T4_L9_Cb, T4_L10_Ca, T4_L10_Cb, T4_L11_Ca, T4_L11_Cb)

uds.4.2014 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2014-Full-Dataset.xlsx", sheet = "Table4", skip = 0) %>% select(BHCMISID, GrantNumber, T4_L1_Ca, T4_L2_Ca, T4_L3_Ca, T4_L4_Ca, T4_L5_Ca, T4_L16_Ca, T4_L23_Ca, T4_L7_Ca, T4_L7_Cb, T4_L8_Ca, T4_L8_Cb, T4_L9_Ca, T4_L9_Cb, T4_L10_Ca, T4_L10_Cb, T4_L11_Ca, T4_L11_Cb) 
uds.4.2015 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2015-Full-Dataset.xlsx", sheet = "Table4", skip = 2) %>% select(BHCMISID, GrantNumber, T4_L1_Ca, T4_L2_Ca, T4_L3_Ca, T4_L4_Ca, T4_L5_Ca, T4_L16_Ca, T4_L23_Ca, T4_L7_Ca, T4_L7_Cb, T4_L8_Ca, T4_L8_Cb, T4_L9_Ca, T4_L9_Cb, T4_L10_Ca, T4_L10_Cb, T4_L11_Ca, T4_L11_Cb) 
uds.4.2016 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2016-Full-Dataset.xlsx", sheet = "Table4", skip = 1) %>% select(BHCMISID, GrantNumber, T4_L1_Ca, T4_L2_Ca, T4_L3_Ca, T4_L4_Ca, T4_L5_Ca, T4_L16_Ca, T4_L23_Ca, T4_L7_Ca, T4_L7_Cb, T4_L8_Ca, T4_L8_Cb, T4_L9_Ca, T4_L9_Cb, T4_L10_Ca, T4_L10_Cb, T4_L11_Ca, T4_L11_Cb)

uds.4.2017 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2017-Full-Dataset.xlsx", sheet = "Table4", skip = 1) %>% select(BHCMISID, GrantNumber, T4_L1_Ca, T4_L2_Ca, T4_L3_Ca, T4_L4_Ca, T4_L5_Ca, T4_L16_Ca, T4_L23_Ca, T4_L7_Ca, T4_L7_Cb, T4_L8_Ca, T4_L8_Cb, T4_L9_Ca, T4_L9_Cb, T4_L10_Ca, T4_L10_Cb, T4_L11_Ca, T4_L11_Cb) 
uds.4.2018 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2018-Full-Dataset.xlsx", sheet = "Table4", skip = 1) %>% select(BHCMISID, GrantNumber, T4_L1_Ca, T4_L2_Ca, T4_L3_Ca, T4_L4_Ca, T4_L5_Ca, T4_L16_Ca, T4_L23_Ca, T4_L7_Ca, T4_L7_Cb, T4_L8_Ca, T4_L8_Cb, T4_L9_Ca, T4_L9_Cb, T4_L10_Ca, T4_L10_Cb, T4_L11_Ca, T4_L11_Cb) 
uds.4.2019 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2019-Full-Dataset.xlsx", sheet = "Table4", skip = 1) %>% select(BHCMISID, GrantNumber, T4_L1_Ca, T4_L2_Ca, T4_L3_Ca, T4_L4_Ca, T4_L5_Ca, T4_L16_Ca, T4_L23_Ca, T4_L7_Ca, T4_L7_Cb, T4_L8_Ca, T4_L8_Cb, T4_L9_Ca, T4_L9_Cb, T4_L10_Ca, T4_L10_Cb, T4_L11_Ca, T4_L11_Cb)

uds.4.2020 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2020-Full-Dataset.xlsx", sheet = "Table4", skip = 1) %>% select(BHCMISID, GrantNumber, T4_L1_Ca, T4_L2_Ca, T4_L3_Ca, T4_L4_Ca, T4_L5_Ca, T4_L16_Ca, T4_L23_Ca, T4_L7_Ca, T4_L7_Cb, T4_L8_Ca, T4_L8_Cb, T4_L9_Ca, T4_L9_Cb, T4_L10_Ca, T4_L10_Cb, T4_L11_Ca, T4_L11_Cb) 
uds.4.2021 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2021-Full-Dataset.xlsx", sheet = "Table4", skip = 0) %>% select(BHCMISID, GrantNumber, T4_L1_Ca, T4_L2_Ca, T4_L3_Ca, T4_L4_Ca, T4_L5_Ca, T4_L16_Ca, T4_L23_Ca, T4_L7_Ca, T4_L7_Cb, T4_L8_Ca, T4_L8_Cb, T4_L9_Ca, T4_L9_Cb, T4_L10_Ca, T4_L10_Cb, T4_L11_Ca, T4_L11_Cb)

namevec <- c("Year", "BHCMISID", "ModernGranteeID", "<100%", "101-150%", "151-200%", ">200%", "Unknown", "FarmWorker", "Homeless", 
             "Uninsured.young", "Uninsured.old", "MCD.young", "MCD.old", "MCR.young", "MCR.old", "OtherPublic.young", "OtherPublic.old", "COM.young", "COM.old")
uds.4 <- rbind(cbind(Year = 2012, uds.4.2012),
               cbind(Year = 2013, uds.4.2013),
               cbind(Year = 2014, uds.4.2014),
                  cbind(Year = 2015, uds.4.2015),
                  cbind(Year = 2016, uds.4.2016),
                  cbind(Year = 2017, uds.4.2017),

                  cbind(Year = 2018, uds.4.2018),
                  cbind(Year = 2019, uds.4.2019), 
                  cbind(Year = 2020, uds.4.2020),
                  cbind(Year = 2021, uds.4.2021)) %>%

  rename_all(vars(namevec), function(x) namevec)

uds.4 <- rbind(
uds.cw %>% 
  inner_join(read.csv(file = "//Users/jhm65/Dropbox/Git/APMs/data_processed/Merged_Table4.csv"), by = c("link.ID" = "BHCMIS.ID", "Year")) %>% 
  filter(Year %in% c(2007, 2011)) %>% 
  select(Year, BHCMIS.ID, ModernGranteeID, T4_L1_CA, T4_L2_CA, T4_L3_CA, T4_L4_CA, T4_L5_CA, 
         T4_L16_CA, T4_L23_CA, 
         T4_L7_CA, T4_L7_CB, T4_L8_CA, T4_L8_CB, T4_L9_CA, T4_L9_CB, T4_L10_CA, T4_L10_CB, T4_L11_CA, T4_L11_CB) %>% 
  setNames(namevec), 
uds.cw %>% 
  inner_join(read.csv(file = "//Users/jhm65/Dropbox/Git/APMs/data_processed/Merged_Table4.csv"), by = c("link.ID" = "BHCMIS.ID", "Year")) %>% 
  filter(Year %in% c(2008:2010)) %>% 
  select(Year, BHCMIS.ID, ModernGranteeID, T4_L1_CA, T4_L2_CA, T4_L3_CA, T4_L4_CA, T4_L5_CA, 
         T4_L16_CA, T4_L23_CA, 
         T4_L7_CMIA, T4_L7_CMIB, T4_L8_CMIA, T4_L8_CMIB, T4_L9_CMIA, T4_L9_CMIB, T4_L10_CMIA, T4_L10_CMIB, T4_L11_CMIA, T4_L11_CMIB) %>% 
  setNames(namevec), 
uds.4)

uds.4 %>% 
  arrange(BHCMISID, Year)

# Put covariate tables together

uds.covariates <- inner_join(uds.3B, uds.4, by = c("Year","BHCMISID","ModernGranteeID"))

for (i in 4:ncol(uds.covariates)){
  uds.covariates[,i] <- as.numeric(gsub("-", "", uds.covariates[,i]))
}

#uds.covariates %>% filter(`Grant Number` == "H80CS24200")

for (i in 1:nrow(uds.covariates)){
  if (!is.na(uds.covariates[i,]$Total)) {
uds.covariates[i,is.na(uds.covariates[i,])] <- 0
} 
}

uds.covariates <-  uds.covariates %>% 
    mutate(Pct.NHA = NHA / Total * 100,
         Pct.NHB = NHB / Total * 100,
         Pct.HL = HL / Total * 100,
         Pct.NHW = NHW / Total * 100,
         Pct.ESL = Eng2ndLang / Total * 100,
         Pct.LessThan100pctFPL = ifelse(Total == Unknown, 100, `<100%` / (Total- Unknown) * 100),
         Pct.LessThan200pctFPL = ifelse(Total == Unknown, 100, (`<100%` + `101-150%` + `151-200%`) / (Total - Unknown) * 100),
         Pct.FarmWorker = FarmWorker / Total * 100,
         Pct.Homeless = Homeless / Total * 100,
         Pct.None = (Uninsured.young + Uninsured.old) / Total * 100, 
         Pct.MCD = (MCD.young + MCD.old) / Total * 100, 
         Pct.MCR = (MCR.young + MCR.old) / Total * 100, 
         Pct.COM = (COM.young + COM.old) / Total * 100) %>% 
    select(Year:ModernGranteeID, Total, Pct.NHA:Pct.COM)

uds.covariates %>% 
  arrange(BHCMISID, Year)

```

## Table 6A
```{r}
uds.6a.2011andBefore <- uds.cw %>% 
  inner_join(read.csv(file = "//Users/jhm65/Dropbox/Git/APMs/data_processed/Merged_Table6A.csv"), by = c("link.ID" = "BHCMIS.ID", "Year")) %>% 
  select(BHCMIS.ID, Year, T6A_L10_CA:T6A_L26D_CB) %>% 
  filter(Year >= 2010) %>% 
  rename(BHCMISID = BHCMIS.ID) %>% 
  rename_all(.funs = toupper)

uds.6a.2012 <- cbind(Year = 2012, read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2012-Full-Dataset.xlsx", sheet = "Table6A", skip = 0)) %>% rename_all(.funs = toupper)
uds.6a.2013 <- cbind(Year = 2013, read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2013-Full-Dataset.xlsx", sheet = "Table6A", skip = 0)) %>% rename_all(.funs = toupper)

uds.6a.2014 <- cbind(Year = 2014, read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2014-Full-Dataset.xlsx", sheet = "Table6A", skip = 0)) %>% rename_all(.funs = toupper)
uds.6a.2015 <- cbind(Year = 2015, read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2015-Full-Dataset.xlsx", sheet = "Table6A", skip = 2)) %>% rename_all(.funs = toupper)
uds.6a.2016 <- cbind(Year = 2016, read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2016-Full-Dataset.xlsx", sheet = "Table6A", skip = 1)) %>% rename_all(.funs = toupper)

uds.6a.2017 <- cbind(Year = 2017, read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2017-Full-Dataset.xlsx", sheet = "Table6A", skip = 1)) %>% rename_all(.funs = toupper)
uds.6a.2018 <- cbind(Year = 2018, read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2018-Full-Dataset.xlsx", sheet = "Table6A", skip = 1)) %>% rename_all(.funs = toupper)
uds.6a.2019 <- cbind(Year = 2019, read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2019-Full-Dataset.xlsx", sheet = "Table6A", skip = 1)) %>% rename_all(.funs = toupper)

uds.6a.2020 <- cbind(Year = 2020, read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2020-Full-Dataset.xlsx", sheet = "Table6A", skip = 1)) %>% rename_all(.funs = toupper)
uds.6a.2021 <- cbind(Year = 2021, read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2021-Full-Dataset.xlsx", sheet = "Table6A", skip = 0)) %>% rename_all(.funs = toupper) %>% filter(!is.na(BHCMISID))

uds.6a.2012[,-c(1:3)] <- data.frame(sapply(uds.6a.2012[,-c(1:3)], as.numeric))
uds.6a.2013[,-c(1:3)] <- data.frame(sapply(uds.6a.2013[,-c(1:3)], as.numeric))

uds.6a.2014[,-c(1:3)] <- data.frame(sapply(uds.6a.2014[,-c(1:3)], as.numeric))
uds.6a.2015[,-c(1:3)] <- data.frame(sapply(uds.6a.2015[,-c(1:3)], as.numeric))
uds.6a.2016[,-c(1:3)] <- data.frame(sapply(uds.6a.2016[,-c(1:3)], as.numeric))

uds.6a.2017[,-c(1:3)] <- data.frame(sapply(uds.6a.2017[,-c(1:3)], as.numeric))
uds.6a.2018[,-c(1:3)] <- data.frame(sapply(uds.6a.2018[,-c(1:3)], as.numeric))
uds.6a.2019[,-c(1:3)] <- data.frame(sapply(uds.6a.2019[,-c(1:3)], as.numeric))

uds.6a.2020[,-c(1:3)] <- data.frame(sapply(uds.6a.2020[,-c(1:3)], as.numeric))
uds.6a.2021[,-c(1:3)] <- data.frame(sapply(uds.6a.2021[,-c(1:3)], as.numeric))

uds.6a <- bind_rows(uds.6a.2012, uds.6a.2013, uds.6a.2014,
      uds.6a.2015, uds.6a.2016, uds.6a.2017, uds.6a.2018, uds.6a.2019, uds.6a.2020, uds.6a.2021, uds.6a.2011andBefore) %>% 
  arrange(BHCMISID, YEAR) %>% 
  select(YEAR:T6A_L34_CB)

uds.6a
```

## Table 6B Measures
```{r}

uds.6b.2012 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2012-Full-Dataset.xlsx", sheet = "Table6BClinicalmeasures", skip = 0) 
uds.6b.2013 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2013-Full-Dataset.xlsx", sheet = "Table6BClinicalmeasures", skip = 0)

uds.6b.2014 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2014-Full-Dataset.xlsx", sheet = "Table6BClinicalmeasures", skip = 0) 
uds.6b.2015 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2015-Full-Dataset.xlsx", sheet = "Table6BClinicalmeasures", skip = 1) 
uds.6b.2016 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2016-Full-Dataset.xlsx", sheet = "Table6BClinicalmeasures", skip = 0)

uds.6b.2017 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2017-Full-Dataset.xlsx", sheet = "Table6BClinicalmeasures", skip = 0) 
uds.6b.2018 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2018-Full-Dataset.xlsx", sheet = "Table6BClinicalmeasures", skip = 0) 
uds.6b.2019 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2019-Full-Dataset.xlsx", sheet = "Table6BClinicalmeasures", skip = 0)

uds.6b.2020 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2020-Full-Dataset.xlsx", sheet = "Table6BClinicalmeasures", skip = 0) 
uds.6b.2021 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2021-Full-Dataset.xlsx", sheet = "Table6BClinicalmeasures", skip = 0)

uds.6b.names <- c("BHCMISID","GrantNumber","Imm.Num","Imm.Rate","imm.denom","Pap.Rate","Pap.Num","pap.denom",
                  "crc.rate","crc.num","crc.denom", 
                  "ChildBMI.rate","ChildBMI.num","ChildBMI.denom","AdultBMI.rate","AdultBMI.num","AdultBMI.denom", "asthma.rate","asthma.num","asthma.denom", 
                  "cad.rate","cad.num","cad.denom", "ivd.rate","ivd.num","ivd.denom")

uds.quality.6b <- rbind(
  cbind(Year = 2012, uds.6b.2012[,c(1:8, 30:32, 9:14, 21:29)] %>% setNames(uds.6b.names)),
  cbind(Year = 2013, uds.6b.2013[,c(1:8, 30:32, 9:14, 21:29)] %>% setNames(uds.6b.names)),

  cbind(Year = 2014, uds.6b.2014[,c(1:8, 27:29, 9:14, 18:26)] %>% setNames(uds.6b.names)),
  cbind(Year = 2015, uds.6b.2015[,c(1:8, 27:29, 9:14, 18:26)] %>% setNames(uds.6b.names)),
  cbind(Year = 2016, uds.6b.2016[,c(1:8, 27:29, 9:14, 18:26)] %>% setNames(uds.6b.names)),

  cbind(Year = 2017, uds.6b.2017[,c(1:8, 27:29, 9:14, 18:26)] %>% setNames(uds.6b.names)),
  cbind(Year = 2018, uds.6b.2018[,c(1:8, 27:29, 9:14, 18:26)] %>% setNames(uds.6b.names)),
  cbind(Year = 2019, uds.6b.2019[,c(1:8, 27:29, 9:14, 18:26)] %>% setNames(uds.6b.names)),

  cbind(Year = 2020, cbind(uds.6b.2020[,c(1:8, 27:29, 12:17)], NA, NA, NA, uds.6b.2020[,c(21:26)]) %>% setNames(uds.6b.names)),
  cbind(Year = 2021, cbind(uds.6b.2021[,c(1:8, 27:29, 12:17)], NA, NA, NA, uds.6b.2021[,c(21:26)]) %>% setNames(uds.6b.names))) %>% 
  mutate(imm.rate = as.numeric(Imm.Num) / as.numeric(imm.denom), 
         pap.rate = as.numeric(Pap.Num) / as.numeric(pap.denom), 
         imm.denom = as.numeric(imm.denom), 
         pap.denom = as.numeric(pap.denom), 
         crc.rate = as.numeric(crc.num) / as.numeric(crc.denom), 
         crc.denom = as.numeric(crc.denom), 
         childbmi.rate = as.numeric(as.numeric(ChildBMI.num) / as.numeric(ChildBMI.denom)), 
         childbmi.denom = as.numeric(ChildBMI.denom),
         adultbmi.rate = as.numeric(as.numeric(AdultBMI.num) / as.numeric(AdultBMI.denom)), 
         adultbmi.denom = as.numeric(AdultBMI.denom),
         asthma.rate = as.numeric(as.numeric(asthma.num) / as.numeric(asthma.denom)), 
         cad.rate = as.numeric(as.numeric(cad.num) / as.numeric(cad.denom)), 
         ivd.rate = as.numeric(as.numeric(ivd.num) / as.numeric(ivd.denom)), 
         ) %>% 
  select(Year, BHCMISID, GrantNumber, imm.rate, imm.denom, pap.rate, pap.denom, crc.rate, crc.denom, asthma.rate, cad.rate, ivd.rate, childbmi.rate, childbmi.denom, adultbmi.rate, adultbmi.denom) %>%

  bind_rows(uds.cw %>% 
  inner_join(read.csv(file = "//Users/jhm65/Dropbox/Git/APMs/data_processed/Merged_Table6B.csv"), by = c("link.ID" = "BHCMIS.ID", "Year")) %>% 
  filter(Year != 2008) %>% 
    arrange(BHCMIS.ID, Year) %>% 
  #select(BHCMIS.ID:link.ID, T3B_L1_CA:T3B_L7_CC) %>% 
  #filter(BHCMIS.ID == '010060') %>% 
  select(BHCMIS.ID, Year, T6B_L10_N_CC:T6B_L11_D_CC, T6B_L10_CA:T6B_L11_CC, T6B_L12_CA:T6B_L13_CC) %>% 
  mutate(imm.rate = ifelse(Year == 2009, T6B_L10_N_CC / T6B_L10_D_CC, 
                    ifelse(Year %in% c(2010:2011), T6B_L10_CC / T6B_L10_CB, NA)), 
         imm.denom = ifelse(Year == 2009, T6B_L10_D_CC, 
                     ifelse(Year %in% c(2010:2011), T6B_L10_CA, NA)),
         pap.rate = ifelse(Year == 2009, T6B_L11_N_CC / T6B_L11_D_CC, 
                    ifelse(Year %in% c(2010:2011), T6B_L11_CC / T6B_L11_CB, NA)), 
         pap.denom = ifelse(Year == 2009, T6B_L11_D_CC, 
                     ifelse(Year %in% c(2010:2011), T6B_L11_CA, NA)), 
         childbmi.rate = T6B_L12_CC / T6B_L12_CB, 
         childbmi.denom = T6B_L12_CA, 
         adultbmi.rate = T6B_L13_CC / T6B_L13_CB, 
         adultbmi.denom = T6B_L13_CA ) %>% 
  select(BHCMIS.ID, Year, imm.rate:adultbmi.denom) %>% 
  rename(BHCMISID = BHCMIS.ID)) %>% 
    arrange(BHCMISID, Year)

## Removes FQHCs that were not open during those first 3 years of the study period
uds.quality.6b <- uds.quality.6b %>% 
  inner_join(uds.cw %>% mutate(STATE = toupper(STATE)) %>% unique() %>% mutate(STATE = ifelse(BHCMIS.ID == '010530', 'ME', 
                                                                                       ifelse(BHCMIS.ID == '090090', 'AZ', STATE))) %>% 
               select(BHCMIS.ID, STATE) %>% unique(), by = c("BHCMISID" = "BHCMIS.ID"))

# uds.cw %>% mutate(STATE = toupper(STATE)) %>% 
#   select(BHCMIS.ID, STATE)  %>% unique() %>% 
#   group_by(BHCMIS.ID) %>% 
#   mutate(row = row_number()) %>% 
#   group_by(BHCMIS.ID) %>% 
#   mutate(max.row = max(row)) %>% 
#   filter(max.row > 1)
#
# uds.cw %>% 
#   mutate(STATE = toupper(STATE)) %>% 
#   select(BHCMIS.ID, STATE) %>% 
#   group_by(BHCMIS.ID, STATE) %>% 
#   summarise(n()) %>% 
#   filter(BHCMIS.ID %in% c('010530','090090'))
# 
# uds.cw %>% filter(BHCMIS.ID == '010530')

```

## Table 6B - EHR or Sampling Used
```{r}
vec.years <- 2012:2021
vec.skips <- rep(0, length(vec.years)) 
vec.skips[c(4)] <- 2
vec.skips[c(5,6,7,8,9)] <- 1
vec.files <- paste("//Users/jhm65/Dropbox/Git/UDS/UDS-",vec.years, "-Full-Dataset.xlsx", sep = "")

uds.quality.sampling.v.ehr <- uds.cw %>% 
  inner_join(read.csv(file = "//Users/jhm65/Dropbox/Git/APMs/data_processed/Merged_Table6B.csv"), by = c("link.ID" = "BHCMIS.ID", "Year")) %>% 
  filter(Year != 2008) %>% 
    arrange(BHCMIS.ID, Year) %>% 
  #select(BHCMIS.ID:link.ID, T3B_L1_CA:T3B_L7_CC) %>% 
  #filter(BHCMIS.ID == '010060') %>% 
  #select(BHCMIS.ID, Year, T6B_L10_N_CC:T6B_L11_D_CC, T6B_L10_CA:T6B_L11_CC) %>% 
  select(BHCMIS.ID, Year, T6B_L10_CA, T6B_L10_CB, T6B_L11_CA, T6B_L11_CB, T6B_L12_CA,	T6B_L12_CB, T6B_L13_CA, T6B_L13_CB) %>% 
  cbind(T6B_L19_CA = NA,	T6B_L19_CB = NA) %>% 
  rename(YEAR = Year)

for (i in 1:length(vec.years)){
read_excel(vec.files[i], sheet = "Table6B", skip = vec.skips[i])
temp <- read_excel(vec.files[i], sheet = "Table6B", skip = vec.skips[i])%>% 
  select(BHCMISID, T6b_L10_Ca, T6b_L10_Cb, T6b_L11_Ca, T6b_L11_Cb, T6B_L19_Ca,	T6B_L19_Cb, T6b_L12_Ca,	T6b_L12_Cb, T6b_L13_Ca, T6b_L13_Cb) %>% 
    mutate(Year = vec.years[i])
colnames(temp) <- toupper(colnames(temp))
colnames(temp)[1] <- "BHCMIS.ID"

  uds.quality.sampling.v.ehr <- uds.quality.sampling.v.ehr %>% 
    rbind(temp) %>% 
    arrange(BHCMIS.ID, YEAR)
}

uds.quality.sampling.v.ehr <- uds.quality.sampling.v.ehr %>% 
  mutate(imm.ehr = ifelse(T6B_L10_CA == T6B_L10_CB, 1, 0), 
         pap.ehr = ifelse(T6B_L11_CA == T6B_L11_CB, 1, 0), 
         crc.ehr = ifelse(T6B_L19_CA == T6B_L19_CB, 1, 0),
         childbmi.ehr = ifelse(T6B_L12_CA == T6B_L12_CB, 1, 0), 
         adultbmi.ehr = ifelse(T6B_L13_CA == T6B_L13_CB, 1, 0)) %>% 
  mutate(ehr = imm.ehr + pap.ehr + crc.ehr + childbmi.ehr + adultbmi.ehr)

```

## Table 7 Clinical Measures

```{r}

uds.7.2012 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2012-Full-Dataset.xlsx", sheet = "Table7Clinicalmeasures", skip = 0) 
uds.7.2013 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2013-Full-Dataset.xlsx", sheet = "Table7Clinicalmeasures", skip = 0)

uds.7.2014 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2014-Full-Dataset.xlsx", sheet = "Table7Clinicalmeasures", skip = 0) 
uds.7.2015 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2015-Full-Dataset.xlsx", sheet = "Table7Clinicalmeasures", skip = 1) 
uds.7.2016 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2016-Full-Dataset.xlsx", sheet = "Table7Clinicalmeasures", skip = 0)

uds.7.2017 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2017-Full-Dataset.xlsx", sheet = "Table7Clinicalmeasures", skip = 0) 
uds.7.2018 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2018-Full-Dataset.xlsx", sheet = "Table7Clinicalmeasures", skip = 0) 
uds.7.2019 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2019-Full-Dataset.xlsx", sheet = "Table7Clinicalmeasures", skip = 0)

uds.7.2020 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2020-Full-Dataset.xlsx", sheet = "Table7Clinicalmeasures", skip = 0) 
uds.7.2021 <- read_excel("//Users/jhm65/Dropbox/Git/UDS/UDS-2021-Full-Dataset.xlsx", sheet = "Table7Clinicalmeasures", skip = 1)

uds.7.names <- c("BHCMISID","GrantNumber","Race","Ethnicity","htn.total","htn.control","dm.total","dm.controlled")

uds.quality.7 <- rbind(
  cbind(Year = 2012, uds.7.2012[,c(1:6, 8:9)] %>% setNames(uds.7.names) %>% mutate(dm.controlled = as.numeric(dm.total) - as.numeric(dm.controlled))),
  cbind(Year = 2013, uds.7.2013[,c(1:6, 8:9)] %>% setNames(uds.7.names) %>% mutate(dm.controlled = as.numeric(dm.total) - as.numeric(dm.controlled))),

  cbind(Year = 2014, uds.7.2014[,c(1:6, 8:9)] %>% setNames(uds.7.names)),
  cbind(Year = 2015, uds.7.2015[,c(1:4, 8:9, 11, 14)] %>% setNames(uds.7.names) %>% mutate(dm.controlled = as.numeric(dm.total) - as.numeric(dm.controlled))),
  cbind(Year = 2016, uds.7.2016[,c(1:4, 8:9, 11, 14)] %>% setNames(uds.7.names) %>% mutate(dm.controlled = as.numeric(dm.total) - as.numeric(dm.controlled))),

  cbind(Year = 2017, uds.7.2017[,c(1:4, 8:9, 11, 14)] %>% setNames(uds.7.names) %>% mutate(dm.controlled = as.numeric(dm.total) - as.numeric(dm.controlled))),
  cbind(Year = 2018, uds.7.2018[,c(1:4, 8:9, 11:12)] %>% setNames(uds.7.names) %>% mutate(dm.controlled = as.numeric(dm.total) - as.numeric(dm.controlled))),
  cbind(Year = 2019, uds.7.2019[,c(1:4, 8:9, 11:12)] %>% setNames(uds.7.names) %>% mutate(dm.controlled = as.numeric(dm.total) - as.numeric(dm.controlled))),

  cbind(Year = 2020, uds.7.2020[,c(1:4, 8:9, 11:12)] %>% setNames(uds.7.names) %>% mutate(dm.controlled = as.numeric(dm.total) - as.numeric(dm.controlled))),
  cbind(Year = 2021, uds.7.2021[,c(1:4, 8:9, 11:12)] %>% setNames(uds.7.names) %>% mutate(dm.controlled = as.numeric(dm.total) - as.numeric(dm.controlled)))) %>% 
  mutate(htn.rate = as.numeric(htn.control) / as.numeric(htn.total), 
         dm.rate = as.numeric(dm.controlled) / as.numeric(dm.total), 
         htn.total = as.numeric(htn.total), 
         dm.total = as.numeric(dm.total)) %>% 
  select(Year:Ethnicity, htn.total, htn.rate, dm.total, dm.rate) %>%

  bind_rows(read.csv(file = "//Users/jhm65/Dropbox/Git/APMs/data_processed/Merged_Table7clin.csv") %>% 
  select(Year, ModernGranteeID:dm.controlled) %>% 
  rename(GrantNumber = ModernGranteeID) %>% 
    mutate(htn.rate = as.numeric(htn.control) / as.numeric(htn.total), 
         dm.rate = as.numeric(dm.controlled) / as.numeric(dm.total), 
         htn.total = as.numeric(htn.total), 
         dm.total = as.numeric(dm.total)) %>% 
    select(Year:Ethnicity, htn.total, htn.rate, dm.total, dm.rate)) %>% 
    arrange(GrantNumber, Year, Ethnicity, Race)

## Removes FQHCs that were not open during those first 3 years of the study period
uds.quality.7 <- uds.quality.7 %>% 
  inner_join(uds.cw %>% mutate(STATE = toupper(STATE)) %>% unique() %>% mutate(STATE = ifelse(BHCMIS.ID == '010530', 'ME', 
                                                                                       ifelse(BHCMIS.ID == '090090', 'AZ', STATE))) %>% 
               select(ModernGranteeID, STATE) %>% unique(), by = c("GrantNumber" = "ModernGranteeID")) %>% 
  mutate(Race = ifelse(Race == '-', NA, Race),
         Ethnicity = ifelse(Ethnicity == '-', NA, Ethnicity))

uds.quality.7[is.na(uds.quality.7$htn.total),]$htn.total <- 0
uds.quality.7[is.na(uds.quality.7$htn.rate),]$htn.rate <- 0
uds.quality.7[is.na(uds.quality.7$dm.total),]$dm.total <- 0
uds.quality.7[is.na(uds.quality.7$dm.rate),]$dm.rate <- 0

uds.quality.7 <- uds.quality.7 %>% 
   inner_join(uds.cw %>% filter(Year == 2011) %>% unique() %>% select(ModernGranteeID, BHCMIS.ID), by = c("GrantNumber" = "ModernGranteeID")) %>% 
  arrange(Year, BHCMIS.ID, GrantNumber) %>% 
  mutate(BHCMISID = ifelse(is.na(BHCMISID), BHCMIS.ID, BHCMISID))

uds.quality.7 %>% 
  filter(Race == " Total" | is.na(Race))%>% 
  arrange(BHCMISID, Year) %>% 
  group_by(Year, BHCMISID) %>% 
  summarise(htn.rate = sum(htn.rate * htn.total) / sum(htn.total), 
            htn.total = sum(htn.total), 
            dm.rate = sum(dm.rate * dm.total) / sum(dm.total), 
            dm.total = sum(dm.total)) %>% 
  arrange(BHCMISID, Year)
```

```{r}
# df.comp %>% 
#   ggplot(aes(x = Year, y = dm.rate)) + geom_point() + geom_smooth()

```

## Financials
```{r}
cw <- read_excel("/Users/jhm65/Dropbox/Git/Data/UDS-990 Crosswalk.xlsx")

fins.revenue.expenses <- read.csv("//Users/jhm65/Dropbox/Git/Data/FQHC_Revenue_vs_Expenses.csv", header = TRUE)
names(fins.revenue.expenses) <- gsub("^X\\.", "", names(fins.revenue.expenses))
names(fins.revenue.expenses) <- gsub("\\.\\.", "", names(fins.revenue.expenses))

staffing <- read.csv("/Users/jhm65/Dropbox/Git/Data/FQHC_Staffing.csv")

fins.revenue.expenses <- uds.cw %>% filter(Year == 2011) %>% unique() %>% select(ModernGranteeID, BHCMIS.ID) %>% 
  left_join(cw, by = c("ModernGranteeID" = "Grant Number")) %>% 
  left_join(fins.revenue.expenses, by = c("EIN" = "EIN")) %>% 
  select(ModernGranteeID:EIN, form_year:fringe_rate) %>% 
  left_join(staffing %>% select(EIN:Staff), by = c("EIN","form_year" = "Year"))
```

## APM History

```{r}

apm.history <- read_excel("/Users/jhm65/Dropbox/Git/APMs/data_raw/APM_History.xlsx") %>%
 mutate(Class = ifelse(is.na(Year), "0 - Control", Classification))

```

## Full dataset

```{r}

included.years <- 2010:2021
included.treat <- 2013:2021

## Broader definitions 5 / 7 quality measures significant improvements overall
# included.years <- 2009:2021
# included.treat <- 2013:2021

## Original good filters
# included.years <- 2009:2019
# included.treat <- 2013:2019

df.comp <- uds.covariates %>% 
  select(Year:BHCMISID, Total:Pct.COM) %>% 
  inner_join(uds.cw %>% select(BHCMIS.ID, STATE) %>% 
               mutate(STATE = toupper(STATE)) %>% 
               unique() %>% mutate(STATE = ifelse(BHCMIS.ID == '010530', 'ME', ifelse(BHCMIS.ID == '090090', 'AZ', STATE))) %>% unique(), by = c("BHCMISID" = "BHCMIS.ID")) %>% 
  select(BHCMISID, Year, STATE, Total:Pct.COM) %>% 
  arrange(BHCMISID, Year) %>% 
  left_join(uds.6a, by = c("BHCMISID","Year" = "YEAR")) %>% 
  left_join(uds.quality.6b %>% select(Year, BHCMISID, imm.rate:adultbmi.denom), by = c("BHCMISID","Year")) %>% 
  left_join(uds.quality.7 %>% 
  filter(Race == " Total" | is.na(Race)) %>% 
  group_by(Year, BHCMISID) %>% 
  summarise(htn.rate = sum(htn.rate * htn.total) / sum(htn.total), 
            htn.total = sum(htn.total), 
            dm.rate = sum(dm.rate * dm.total) / sum(dm.total), 
            dm.total = sum(dm.total)) %>% 
  arrange(BHCMISID, Year), by = c("BHCMISID","Year")) %>% 
    left_join(fins.revenue.expenses, by = c("BHCMISID" = "BHCMIS.ID", "Year" = "form_year")) %>% 
  inner_join(apm.history %>% rename(att = Year), by = c("STATE" = "Postal Code")) %>% 
  ungroup() %>% 
  mutate(att = ifelse(is.na(att), 0, att)) %>% 
  mutate(ID = as.numeric(as.factor(BHCMISID))) %>% 
  arrange(BHCMISID) %>% 
  mutate(ACA = ifelse((STATE %in% c('MI','NH') & Year >= 2014) |
                    (STATE %in% c('PA','IN','AK') & Year >= 2015) |
                    (STATE %in% c('MT','LA') & Year >= 2016) |
                    (STATE %in% c('VA','ME') & Year >= 2019) |
                    (STATE %in% c('ID','UT','NE') & Year >= 2020) |
                    (STATE %in% c('OK','MO') & Year >= 2021) |
                    (STATE %in% c('AZ','AR','CA','CO','CT','DE','DC','HI','IL','IA','KY','MD','MA','MN','NV','NJ','NM','NY','ND','OH','OR','PA','RI','VT','VA','WA') & Year >= 2014), 1, 0))

class.vec <- rep(sort(unique(df.comp$Classification)), table(df.comp$Classification))

set.seed(123) #432

df.comp <- df.comp %>% 
  group_by(ID) %>% 
  mutate(Class.random = ifelse(Class == "0 - Control", sample(class.vec, size = 1, replace = T), Class), 
         att.random = ifelse(Class == "0 - Control", sample(included.treat, size = 1, replace = T), att)) %>% 
                                 mutate(Tx = ifelse(Year >= att & att > 0, 1, 0)) %>% 
  filter(Year %in% included.years) %>% 
  mutate(year.rel = as.factor(ifelse(att.random > 0, Year - att.random, NA)), 
         Treatment = ifelse(att > 0, 1, 0)) %>% 
  select(BHCMISID:STATE, Name:ID, ACA:Treatment, Total:dm.total, fs_contributions:Staff)

df.comp <- df.comp %>% 
  left_join(df.comp %>% 
  group_by(BHCMISID) %>% 
  summarise(counts = n()) %>% filter(counts == length(included.years)), by = c("BHCMISID")) %>% 
  group_by(BHCMISID) %>% 
  mutate(closed = ifelse(is.na(counts) & Year == max(Year), 1, 0)) #%>%  filter(APM %in% c('Yes','No'))

df.comp$year.rel <- relevel(as.factor(df.comp$year.rel), ref = "-1")

## Filter for balanced panel, remove closures and late openings
df.comp <- df.comp %>%
  left_join(df.comp %>% summarise(Counts = n(), MinYear = min(Year)), by = c("BHCMISID")) %>%
  filter(Counts == length(included.years)) #  filter(Counts < 13 & MinYear == 2009)    filter(Counts == length(included.years))

## Remove some weird shit in CT
df.comp <- df.comp %>% 
  filter(BHCMISID != "01E00010") %>%  #%>%  filter(STATE != 'CT')
select(BHCMISID:FIPS, ID, Total:Pct.COM, imm.rate:MinYear)

```

## Clean up the missing addresses, geocoding

```{r}

tbcoded <- uds.sites.full[is.na(uds.sites.full$Zip) & uds.sites.full$Year >= included.years[1],]
coded <- geo(street = tbcoded[c(1:1000),]$`Site Street Address`, 
             city = tbcoded[c(1:1000),]$`Site City`, 
             state = tbcoded[c(1:1000),]$`Site State`, 
             method = "census", 
             full_results = TRUE)

for (g in 1:13){
  g
coded <- rbind(coded, 
               geo(street = tbcoded[c((1 + g*1000):((g+1)*1000)),]$`Site Street Address`, 
             city = tbcoded[c((1 + g*1000):((g+1)*1000)),]$`Site City`, 
             state = tbcoded[c((1 + g*1000):((g+1)*1000)),]$`Site State`, 
             method = "census", 
             full_results = TRUE) )
}

coded <- rbind(coded, 
               geo(street = tbcoded[c(14001:14253),]$`Site Street Address`, 
             city = tbcoded[c(14001:14253),]$`Site City`, 
             state = tbcoded[c(14001:14253),]$`Site State`, 
             method = "census", 
             full_results = TRUE) )

coded[grep(pattern = "[[:digit:]]{5}$", coded$matched_address),]

uds.sites.full[is.na(uds.sites.full$Zip) & uds.sites.full$Year >= included.years[1],]$Zip <- str_extract(coded$matched_address, pattern = "[[:digit:]]{5}$")
uds.sites.full[is.na(uds.sites.full$Zip) & uds.sites.full$Year >= included.years[1],]

uds.sites.full.temp <- uds.sites.full

```

```{r}

uds.sites.full <- uds.sites.full.temp

uds.sites.full.temp %>% filter(`Grant Number` %in% c("H80CS00881") & Year == 2016)

chc <- read.csv("//Users/jhm65/Dropbox/Git/Competition/data_raw/chc.csv")

chc %>% filter(Health.Center.Number %in% c("H80CS00881"))

uds.sites.full[is.na(uds.sites.full$Zip) & uds.sites.full$Year >= included.years[1],] <- uds.sites.full[is.na(uds.sites.full$Zip) & uds.sites.full$Year >= included.years[1],] %>%
  left_join(chc, by = c("Grant Number" = "Health.Center.Number", "Site Street Address" = "Site.Address")) %>%
  mutate(Zip = ifelse(is.na(Zip), substr(Health.Center.Organization.ZIP.Code, 1, 5), Zip)) %>%
  select(`BHCMIS ID`:Zip) %>%
  unique()

uds.sites.full[is.na(uds.sites.full$Zip) & uds.sites.full$Year >= included.years[1],] <- uds.sites.full[is.na(uds.sites.full$Zip) & uds.sites.full$Year >= included.years[1],] %>%
  left_join(chc, by = c("Grant Number" = "Health.Center.Number", "Site Name" = "Site.Name")) %>%
  mutate(Zip = ifelse(is.na(Zip), substr(Health.Center.Organization.ZIP.Code, 1, 5), Zip)) %>%
  select(`BHCMIS ID`:Zip) %>%
  unique()

uds.sites.full[is.na(uds.sites.full$Zip) & uds.sites.full$Year >= included.years[1],]
```

```{r}

table(uds.sites.full.temp$`Location Setting`)

uds.sites.full <- uds.sites.full %>% #filter(`Location Type` == "Permanent" & `Location Setting` %in% c("All Other Clinic Types")) %>% 
  mutate(Access = as.numeric(ifelse(`Total Weekly Hours Of Operation` == "-" & `Operational Schedule` == "Full-Time", 40, 
                  ifelse(`Total Weekly Hours Of Operation` == "-" & `Operational Schedule` == "Part-Time", 10, `Total Weekly Hours Of Operation`))))

uds.sites.full[is.na(uds.sites.full$Zip),]

uds.sites.full %>% 
  arrange(`Total Weekly Hours Of Operation`, `Operational Schedule` )
```

## Classify FQHCs into Treated, control based on zip codes

```{r}

sites.in.zips <- uds.sites.full %>% 
  filter(Year >= included.years[1]) %>% 
  group_by(`BHCMIS ID`, Zip, Year) %>% 
  summarise(Counts = n()) %>% 
  pivot_wider(id_cols = c(Zip, `BHCMIS ID`), names_from = Year, values_from = Counts) %>% 
  pivot_longer(cols = `2010`:`2021`) %>% 
  mutate(value = ifelse(is.na(value), 0, value)) %>% 
  arrange(Zip, name)

clinics.in.zips <- uds.sites.full %>% 
  filter(Year >= included.years[1]) %>%
  select(`BHCMIS ID`, Zip, Year) %>% 
  unique() %>% 
  group_by(Zip, Year) %>% 
  summarise(Counts = n()) %>% 
  pivot_wider(id_cols = c(Zip), names_from = Year, values_from = Counts) %>% 
  pivot_longer(cols = `2011`:`2017`) %>% 
  mutate(value = ifelse(is.na(value), 0, value)) %>% 
  arrange(Zip, name)

sites.in.zips %>% 
  left_join(clinics.in.zips, by = c("Zip","name"))%>% 
  group_by(`Grant Number`, Year) %>% 
  summarise(Clinics = max(Clinics)) 
```

## Write the file
```{r}
#write_csv(df.comp, file = "/Users/jhm65/Dropbox/Projects/Competition/data_processed/analyticalfile.csv")

```